### 2017-08-19 paper.tex
Initial commit - Wrote the abstract - Did basic survey of literature - Not even a skeleton for the prose yet - Empty figures

\documentclass[twocolumn,longbibliography]{quantumarticle-customized}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[pdfpagelabels,pdftex,bookmarks,breaklinks]{hyperref}
\usepackage[all]{hypcap}
\hypersetup{colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue}

\title{Saving $T$ gates by reusing $|A\rangle$ states}
\author{Craig Gidney}
\affiliation{Google, Santa Barbara, CA 93117, USA}
\email{craiggidney@google.com}

\def\sectionautorefname{Section}

\begin{document}
\maketitle

\begin{abstract}
A temporary Toffoli gate (sometimes called ``phase-congruent" or ``matched") must apply the correct state permutation, but may incorrectly phase or entangle the involved qubits until a later Toffoli gate fixes these errors while uncomputing the permutation.
We improve the number of T gates needed to perform temporary Toffoli gates from 4 to 3 via non-destructive use of an ancilla in the state $|A\rangle = \frac{1}{\sqrt{2}} |0\rangle + \frac{1+i}{2} |1\rangle$.

Many common circuits use temporary Toffoli gates, and $T$ gates dominate the cost of error corrected quantum computing, so this optimization is significant and widely applicable.
For example, we reduce the T-cost of $n$-bit addition from $8n + O(1)$ to $6n + O(1)$ which in turn reduces the overall cost of Shor's factoring algorithm by nearly 25\%.
\end{abstract}

\section{Introduction} \label{sec:introduction}

When no ancillae are available, a Toffoli gate requires seven T gates to perform [[[cite]]].
If ancillae are available, that cost can be reduced to four [[[[cite]]] by doing a measurement teleportation thing.
When a Toffoli gate is applied temporarily, in a fashion where phase errors are acceptable until the Toffoli is uncomputed, a simpler inline trick suffices for achieving 4 T gates per Toffoli.

See \autoref{fig:inline-toffoli}.

\begin{figure}
  \centering
  \makebox[\linewidth]{
  }
  \caption{
	Simple Toffoli construction with 7 T gates.
	Doesn't require any ancillae or a later paired Toffoli.
  }
  \label{fig:inline-toffoli}
\end{figure}

\begin{figure}
  \centering
  \makebox[\linewidth]{
  }
  \caption{
	Toffoli construction with 4 T gates based on measurement teleportation.
  }
  \label{fig:ancilla-toffoli}
\end{figure}

\begin{figure}
  \centering
  \makebox[\linewidth]{
  }
  \caption{
	Existing temporary Toffoli construction with 4 T gates.
  }
  \label{fig:inline-temporary-toffoli}
\end{figure}

\begin{figure}
  \centering
  \makebox[\linewidth]{
  }
  \caption{
	Our temporary Toffoli construction, which uses an $|A\rangle$ ancilla and 3 T gates.
  }
  \label{fig:ancilla-temporary-toffoli}
\end{figure}

\begin{figure}
  \centering
  \makebox[\linewidth]{
  }
  \caption{
	$n$-bit inline addition with T-cost of $8n + O(1)$ based on [[[cite]]].
  }
  \label{fig:inline-addition}
\end{figure}

\begin{figure}
  \centering
  \makebox[\linewidth]{
  }
  \caption{
	$n$-bit addition with T-cost of $6n + O(1)$.
  }
  \label{fig:ancilla-addition}
\end{figure}

\begin{figure}
  \centering
  \makebox[\linewidth]{
  }
  \caption{
	Temporary addition construction.
  }
  \label{fig:temporary-ancilla-addition}
\end{figure}

Mention binary to unary as an example?

\section{Acknowledgements}

Austin.

\bibliographystyle{plain}
\bibliography{citations}

\cite{barenco1995}
https://arxiv.org/pdf/quant-ph/9503016.pdf   ->   the paired Toffoli construction that uses 4T

https://arxiv.org/pdf/cond-mat/9409111.pdf   -> shows CCNOT can be done with three 2-qubit interactions, ignores single-qubit???

https://arxiv.org/pdf/1308.4134.pdf -> proof of 7T gate minimum with no ancilla

https://arxiv.org/pdf/1212.5069.pdf  ->   does a single Toffoli with four T gates by turning one of the controls classical

https://arxiv.org/pdf/1210.0974.pdf  ->   a phase-neglected 4T gate construction that only uses a single layer of T gates

Quantum Computation and Quantum Information -> shows standard decomposition into 7T gates. But do they say it's optimal?

https://arxiv.org/pdf/1206.0758.pdf -> just an example of the 7T construction;;; probablynot relevant

\end{document}

### 2017-08-19 paper.tex
Making some of the figures in Qcircuit

@@
 \usepackage[pdfpagelabels,pdftex,bookmarks,breaklinks]{hyperref}
+\usepackage{tikz}
 \usepackage[all]{hypcap}
+\input{Qcircuit}
 \hypersetup{colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue}
@@
 \begin{abstract}
-A temporary Toffoli gate (sometimes called ``phase-congruent" or ``matched") must apply the correct state permutation, but may incorrectly phase or entangle the involved qubits until a later Toffoli gate fixes these errors while uncomputing the permutation.
+A temporary Toffoli gate (sometimes called ``phase-congruent" or ``matched") performs the same operation as a normal Toffoli gate, but may incorrectly phase or entangle the involved qubits until a later Toffoli gate fixes these errors while uncomputing the permutation.
 We improve the number of T gates needed to perform temporary Toffoli gates from 4 to 3 via non-destructive use of an ancilla in the state $|A\rangle = \frac{1}{\sqrt{2}} |0\rangle + \frac{1+i}{2} |1\rangle$.
 
-Many common circuits use temporary Toffoli gates, and $T$ gates dominate the cost of error corrected quantum computing, so this optimization is significant and widely applicable.
+Many quantum circuits use temporary Toffoli gates, and $T$ gates dominate the cost of error corrected quantum computing, so this optimization is significant and widely applicable.
 For example, we reduce the T-cost of $n$-bit addition from $8n + O(1)$ to $6n + O(1)$ which in turn reduces the overall cost of Shor's factoring algorithm by nearly 25\%.
@@
 
+An operation $\tilde{U}$ is a temporary variant of $U$ if it meets the following criteria:
+
+$$\forall b \in \{0, 1\}^{\otimes n}, \; \tilde{U} \cdot \left( |b\rangle \otimes | \psi_0 \rangle \right) = \left( U |b \rangle \right) \otimes | \psi_{1,b} \rangle$$
+
+where $n$ is the number of qubits that $U$ acts on, $b$ is a computational basis state (a bitstring), $\psi_0$ is an input ancilla state, and $\psi_{1,c}$ is the various output ancilla states for each input $b$.
+If we start with a computational basis state, apply the operation, and trace out the ancilla, then the output state will be correct up to a global phase (that may differ depending on the input state).
+The phase and value of the traced-out ancilla may differ for different inputs, so $\tilde{U}$ will not perform the correct operation when its input is in superposition: it causes phase errors and produces entangled garbage.
+Informally speaking, $\tilde{U}$ does the right thing classically but causes quantum-specific problems (i.e. incoherent phases) that must be fixed by a later operation.
+
 See \autoref{fig:inline-toffoli}.
 
+\newcommand{\qH}{\gate{H}}
+\newcommand{\qT}{\gate{T}}
+\newcommand{\qTi}{\gate{T^\dagger}}
+\newcommand{\qS}{\gate{S}}
+\newcommand{\qSi}{\gate{S^\dagger}}
+\newcommand{\qA}{\lstick{|A\rangle}}
+\newcommand{\qO}{\lstick{|0\rangle}}
+
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &   & & &\qw &\qT &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw  &\ctrl{1} &\qw \\
+      &\ctrl{1} &\qw & & = & & &\qw &\qT &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\targ    &\qTi &\targ    &\qw \\
+      &\targ    &\qw & &   & & &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH  &\qw      &\qw \\
+    }
   }
   \caption{
-	Simple Toffoli construction with 7 T gates.
+	Simple Toffoli construction with seven T gates.
 	Doesn't require any ancillae or a later paired Toffoli.
@@
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &       & & &\ctrl{1} &\ctrl{1}     &\qw & &   & & &\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\qw &\qw \\
+      &\ctrl{1} &\qw & & \cong & & &\ctrl{1} &\qSi         &\qw & & = & & &\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw &\qw \\
+      &\targ    &\qw & &       & & &\targ    &\gate{Z}\qwx &\qw & &   & & &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qw \\
+    }
+  }
+  \caption{
+	Existing temporary Toffoli construction with 4 T gates.
+  }
+  \label{fig:inline-temporary-toffoli}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw  &\qw &\qw    &\ctrl{1}   &\qw \\
+      &\ctrl{1} &\qw & & = & & &     &\qw &\qw &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw  &\qw &\qw    &\gate{Z}   &\qw \\
+      &\targ    &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw  &\qw &\qw    &\qw\cwx    &\qw \\
+      &         &    & &   & & & \qO &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qSi &\qH &\meter &\cw\cwx    & \\
+    }
   }
@@
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &    &\ctrl{1} &\qw & &       &&&     &\ctrl{1} &\ctrl{1} &\qw  &&&   & & & \lstick{x} &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\qw &\qw &\rstick{x} \\
+      &    &\ctrl{1} &\qw & & \cong &&&     &\ctrl{1} &\qS      &\qw  &&& = & & & \lstick{y} &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw &\qw &\rstick{y} \\
+      &\qO &\targ    &\qw & &       &&& \qO &\targ    &\qw      &\qw  &&&   & & & \qA        &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qw &\rstick{x \land y} \\
+    }
   }
   \caption{
-	Existing temporary Toffoli construction with 4 T gates.
+	Control merging with three T gates.
+	The $|A\rangle$ state is recovered when uncomputing the merged control.
   }
-  \label{fig:inline-temporary-toffoli}
+  \label{fig:ancilla-temporary-toffoli}
 \end{figure}
@@
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &       &&&     &\ctrl{1} &\ctrl{1} &\qw  &&&   & & & \lstick{x} &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\rstick{x} \\
+      &\ctrl{1} &\qw & & \cong &&&     &\ctrl{2} &\qS      &\qw  &&& = & & & \lstick{y} &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\rstick{y} \\
+      &\targ    &\qw & &       &&&     &\targ    &\qw      &\qw  &&&   & & & \lstick{t} &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\rstick{t \oplus (x \land y)} \\
+      &         &    & &       &&& \qO &\targ    &\qw      &\qw  &&&   & & & \qA        &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qw &\rstick{x \land y} \\
+    }
   }
@@
 	Our temporary Toffoli construction, which uses an $|A\rangle$ ancilla and 3 T gates.
+	The $|A\rangle$ state is recovered when uncomputing the temporary Toffoli.
   }
@@
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &    &\ctrl{2}          &\qw               &\qw &&     &&&     &\ctrl{2}          &\qw               &\qw &\ctrl{1} &\qw      &\qw &&& &&& \lstick{x} &\ctrl{2} &\qw      &\qw &\ctrl{3} &\qw  &\qw      &\qw &\qw      &\qw  &\qw      &\qw &\rstick{x}         \\
+      &    &\qw               &\ctrl{1}          &\qw &&\cong&&&     &\qw               &\ctrl{1}          &\qw &\qS      &\ctrl{1} &\qw &&&=&&& \lstick{y} &\qw      &\qw      &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{1} &\qw &\rstick{y}         \\
+      &    &\multigate{1}{+1} &\multigate{1}{+1} &\qw &&     &&&     &\multigate{1}{+1} &\multigate{1}{+1} &\qw &\qS\qwx  &\qS      &\qw &&& &&& \lstick{z} &\targ    &\ctrl{1} &\qw &\qw      &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\targ    &\qw &\rstick{(x+y+z)_0} \\
+      &\qO &\ghost{+1}        &\ghost{+1}        &\qw &&     &&& \qO &\ghost{+1}        &\ghost{+1}        &\qw &\qw      &\qw      &\qw &&& &&& \qA        &\qw      &\targ    &\qT &\targ    &\qSi &\targ    &\qT &\targ    &\qTi &\qH      &\qw &\rstick{(x+y+z)_1} \\
+    }
   }
   \caption{
-	Temporary addition construction.
+	Temporary full adder.
   }
@@
 Mention binary to unary as an example?
+
+Show example doing an $n$-control NOT with $6n + O(1)$ or $3n + O(1)$ if temporary.
+
+Mention that this trick works more generally. Anytime you have an ancilla that starts HT and ends Tdag H.
 

### 2017-08-20 paper.tex
Oh. That 4T Toffoli circuit can be cut into merge/unmerge pieces with total cost 4 instead of 6.

@@
       &\ctrl{1} &\qw & & = & & &     &\qw &\qw &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw  &\qw       &\qw &\qw    &\gate{Z}       &\qw \\
-      &\targ    &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw  &\qw &\qw    &\qw\cwx    &\qw \\
-      &         &    & &   & & & \qO &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qSi &\qH &\meter &\cw\cwx    & \\
+      &\targ    &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\qw  &\targ     &\qw &\qw    &\qw\cwx        &\qw \\
+      &         &    & &   & & & \qO &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qSi &\ctrl{-1} &\qH &\meter &\cw\cwx\bullet & \\
     }
@@
 	The $|A\rangle$ state is recovered when uncomputing the merged control.
+  }
+  \label{fig:ancilla-temporary-toffoli}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x &&\ctrl{1} &\qw & x  && &&          &&\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw  &\qw &\\
+      &y &&\ctrl{1} &\qw & y  &&=&&          &&\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\qw  &\qw &\\
+      &  &&         &\qw & xy && &&|0\rangle &&\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH &\qSi &\qw &\\
+    }
+  }
+  \caption{
+	Control merging with T-cost of 4.
+	The leading $H$ and $T$ gates can be cut by directly feeding an $|A\rangle$ state into the circuit, instead of using one to perform the $T$ gate.
+  }
+  \label{fig:ancilla-temporary-toffoli}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x  &&\ctrl{1} &\qw &x && &&\qw &\qw    &\ctrl{1} &\qw \\
+      &y  &&\ctrl{1} &\qw &y &&=&&\qw &\qw    &\gate{Z} &\qw \\
+      &xy &&\qw      &    &  && &&\qH &\meter &\cw \cwx \bullet &    \\
+    }
+  }
+  \caption{
+	Control un-merging with T-cost of 0.
+  }
+  \label{fig:ancilla-temporary-toffoli}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw  &\qw      &\qw &\qw    &\ctrl{1}         &\qw \\
+      &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\qw  &\qw      &\qw &\qw    &\gate{Z}         &\qw \\
+      &         &    &&=&&         &\ctrl{1} &\qw      &    &&=&&|0\rangle &&\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH &\qSi &\ctrl{1} &\qH &\meter &\cw \cwx \bullet &    \\
+      &\targ    &\qw && &&\qw      &\targ    &\qw      &\qw && &&          &&\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw      &\qw &\qw  &\targ    &\qw &\qw    &\qw              &\qw \\
+    }
+  }
+  \caption{
+	Toffoli gate from control merging and un-merging with T-cost 4.
+	Equivalent to the construction from [[[[cite]]]
   }
@@
 
+https://arxiv.org/pdf/1706.05113.pdf   has a ctrl-add with 21N + O(1). This technique+commutators achieves 8N.
+
 \end{document}

### 2017-08-22 paper.tex
Tweaking focus towards temporary-and

--- 2017-08-19 paper.tex
@@
 \usepackage[pdfpagelabels,pdftex,bookmarks,breaklinks]{hyperref}
+\usepackage{tikz}
 \usepackage[all]{hypcap}
@@
 
-\title{Saving $T$ gates by reusing $|A\rangle$ states}
+\input{Qcircuit}
+\newcommand{\qH}{\gate{H}}
+\newcommand{\qT}{\gate{T}}
+\newcommand{\qTi}{\gate{T^\dagger}}
+\newcommand{\qS}{\gate{S}}
+\newcommand{\qSi}{\gate{S^\dagger}}
+\newcommand{\qA}{\lstick{|A\rangle}}
+\newcommand{\qO}{\lstick{|0\rangle}}
+
+\title{T gates, temporary Toffolis, and halving the cost of quantum computation}
 \author{Craig Gidney}
@@
 \begin{abstract}
-A temporary Toffoli gate (sometimes called ``phase-congruent" or ``matched") must apply the correct state permutation, but may incorrectly phase or entangle the involved qubits until a later Toffoli gate fixes these errors while uncomputing the permutation.
-We improve the number of T gates needed to perform temporary Toffoli gates from 4 to 3 via non-destructive use of an ancilla in the state $|A\rangle = \frac{1}{\sqrt{2}} |0\rangle + \frac{1+i}{2} |1\rangle$.
+We show how to compute the logical-and of two controls using four T gates, and uncompute it using no T gates.
+This construction halves the average T-cost of temporary Toffoli gates.
+A temporary Toffoli gate is one which will be uncomputed later (typically by a second Toffoli gate).
 
-Many common circuits use temporary Toffoli gates, and $T$ gates dominate the cost of error corrected quantum computing, so this optimization is significant and widely applicable.
-For example, we reduce the T-cost of $n$-bit addition from $8n + O(1)$ to $6n + O(1)$ which in turn reduces the overall cost of Shor's factoring algorithm by nearly 25\%.
+We use our temporary logical-and construction to produce quantum circuits with T gate counts lower than previously reported.
+T gates dominate the cost of surface-code-based error corrected quantum computation.
+Temporary Toffolis are used in many common quantum circuits, such as addition.
+As such, the optimization we present in this paper are widely applicable and represent significant savings in the projected costs of quantum computation.
+For example, by applying previously known optimizations and the ones presented in this paper, we improve on the projected T-cost of Shor's algorithm from [[[[cite]]] by 70\% and the projected minimum time down by 66%.
+
+We demonstrate an $n$-bit adder with T-cost $4n + O(1)$, a controlled $n$-bit adder with T-cost $8n + O(1)$, an $n$-controlled NOT with T-cost $4n + O(1)$, and a temporary pop-count with T-cost $6n + O(1)$.
+
+We discuss investing $|A\rangle$ states, a fallback optimization technique that may reduce the number of T gates needed to compute quantum circuits not amenable to temporary ANDs.
 \end{abstract}
@@
 
-When no ancillae are available, a Toffoli gate requires seven T gates to perform [[[cite]]].
-If ancillae are available, that cost can be reduced to four [[[[cite]]] by doing a measurement teleportation thing.
-When a Toffoli gate is applied temporarily, in a fashion where phase errors are acceptable until the Toffoli is uncomputed, a simpler inline trick suffices for achieving 4 T gates per Toffoli.
+The textbook construction of a Toffoli gate uses seven T gates [[[cite textbook]]] \autoref{fig:textbook-toffoli} and eight Clifford gates.
+On an error corrected quantum computer that uses the surface code, T gates are XXX orders of magnitude more expensive than Clifford gates [[[cite]]].
+Because of this disparity, optimizing error corrected quantum circuits largely consists of reducing their T-cost (the number of T gates used, or rather the number of $|A\rangle = \frac{1}{\sqrt{2}} |0\rangle + \frac{1+i}{2} |1\rangle$ states which must be distilled in order to perform those T gates).
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1}&\qw & &   & & &\ctrl{1}&\qw  &\ctrl{1}&\qT &\qw     &\qw  &\ctrl{2}&\qw &\qw     &\qw  &\ctrl{2}&\qw &\qw \\
+      &\ctrl{1}&\qw & & = & & &\targ   &\qTi &\targ   &\qT &\ctrl{1}&\qw  &\qw     &\qw &\ctrl{1}&\qw  &\qw     &\qw &\qw \\
+      &\targ   &\qw & &   & & &\qw     &\qw  &\qH     &\qT &\targ   &\qTi &\targ   &\qT &\targ   &\qTi &\targ   &\qH &\qw \\
+    }
+  }
+  \caption{
+	Textbook Toffoli construction [[[[cite]]]] from eight Clifford gates and seven T gates.
+  }
+  \label{fig:textbook-toffoli}
+\end{figure}
+
+The textbook Toffoli construction is optimal in some highly constrained situations (e.g.
+not allowed to use any ancilla qubits and not allowed to merge T gates across operations [[[cite]]]), but is not optimal in general.
+For example, when two Toffoli gates share a control, applying the textbook construction to each places two $T$ gates on the shared control.
+These two $T$ gates can be merged into a single $S$ gate, thereby reducing the T-cost of each Toffoli from 7 to 6.
+
+More commonly, Toffoli gates tend to come in matched pairs where one is uncomputing the effect of the other.
+When this occurs, the three $T$ gates on the control qubits of the textbook construction are no longer necessary \autoref{fig:bad-phase-toffoli}.
+Removing those $T$ gates introduces phase errors, but the second Toffoli gate will uncompute those errors while uncomputing the state permutation.
+This technique reduces the T-cost of paired Toffolis from 7 to 4.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &       & & &\ctrl{1}     &\ctrl{1} &\qw & &   & & &\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\qw &\qw \\
+      &\ctrl{1} &\qw & & \cong & & &\qS          &\ctrl{1} &\qw & & = & & &\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw &\qw \\
+      &\targ    &\qw & &       & & &\gate{Z}\qwx &\targ    &\qw & &   & & &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qw \\
+    }
+  }
+  \caption{
+	Existing temporary Toffoli construction with 4 T gates.
+  }
+  \label{fig:bad-phase-toffoli}
+\end{figure}
+
+In [[[cite]]], Jones improves on the paired construction by demonstrating how to reduce the T-cost of any Toffoli to 4.
+Jones does this by using a temporary ancilla and applying an erasure technique. [[[figures]]]
+
+We improve on Jones by separating their construction into three pieces (merge, use, erase), and showing how to apply these pieces to circuit construction tasks.
+In particular, we reduce the average T-cost of paired Toffolis from 4 to 2.
+Many common circuits (e.g. addition) have paired Toffolis, and so this optimization can halve their T-cost.
+
+\section{Temporary ANDs}
 
@@
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw  &\qw       &\qw &\qw    &\ctrl{1}       &\qw \\
+      &\ctrl{1} &\qw & & = & & &     &\qw &\qw &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw  &\qw       &\qw &\qw    &\gate{Z}       &\qw \\
+      &\targ    &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\qw  &\targ     &\qw &\qw    &\qw\cwx        &\qw \\
+      &         &    & &   & & & \qO &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qSi &\ctrl{-1} &\qH &\meter &\cw\cwx\bullet & \\
     }
-  \caption{
-	Simple Toffoli construction with 7 T gates.
-	Doesn't require any ancillae or a later paired Toffoli.
-  }
-  \label{fig:inline-toffoli}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \makebox[\linewidth]{
   }
@@
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x &&\ctrl{1} &\qw & x  && &&          &&\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw &\qw &\\
+      &y &&\ctrl{1} &\qw & y  &&=&&          &&\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\qw &\qw &\\
+      &  &&         &\qw & xy && &&|A\rangle &&\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH &\qS &\qw &\\
+    }
   }
   \caption{
-	Existing temporary Toffoli construction with 4 T gates.
+	Control merging with T-cost of 4 (three T gates, one $|A\rangle$ state).
+	The $|A\rangle$ state is fed directly into the circuit, instead of using an $H$ and $T$ gate on a $|0\rangle$ state (which would consume an $|A\rangle$ state anyways).
+	The input $|A\rangle$ state can be recovered by using the un-merging construction in \autoref{fig:unmerge-recover-a}, reducing the net T cost of this circuit to 3.
+	However, the un-merging construction in \autoref{fig:unmerge-free} is a better choice, despite not recovering the $|A\rangle$ state, because the total T-cost of merge+unmerge using \autoref{fig:unmerge-free} is 4+0=4 instead of 4+3-1=6.
   }
-  \label{fig:inline-temporary-toffoli}
+  \label{fig:ancilla-temporary-toffoli}
 \end{figure}
@@
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x_k &&\targ    &\ctrl{1}&\qw      &\qw &\qw &\qw &\qw     &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw     &\qw &\qw &\qw      &\ctrl{1}&\targ    &\ctrl{1}&\qw &x_k     &&&&& \\
+      &y_k &&\targ    &\ctrl{1}&\qw      &\qw &\qw &\qw &\qw     &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw     &\qw &\qw &\qw      &\ctrl{1}&\qw      &\targ   &\qw &&&(x+y)_k &&& \\
+      &    &&         &        &\ctrl{1} &\qw &\qw &\qw &\qw     &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw     &\qw &\qw &\ctrl{1} &\qw     &         &        &    &&&        &&& \\
+      &c_k &&\ctrl{-3}&\qw     &\targ    &\qw &\qw &    &c_{k+1} &    &    &    &\ldots &    &    &    &c_{k+1} &    &    &\targ    &\qw     &\ctrl{-3}&\qw     &\qw &c_k     &&&&& \\
+    }
+  }
+  \caption{
+	Inline full adder block with a T-cost of 4.
+  }
+  \label{fig:addition}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x  &&\ctrl{1} &\qw &x && &&\qw &\qw    &\ctrl{1} &\qw \\
+      &y  &&\ctrl{1} &\qw &y &&=&&\qw &\qw    &\gate{Z} &\qw \\
+      &xy &&\qw      &    &  && &&\qH &\meter &\cw \cwx \bullet &    \\
+    }
+  }
+  \caption{
+	Control un-merging with T-cost of 0.
+  }
+  \label{fig:unmerge-free}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw  &\qw      &\qw &\qw    &\ctrl{1}         &\qw \\
+      &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\qw  &\qw      &\qw &\qw    &\gate{Z}         &\qw \\
+      &         &    &&=&&         &\ctrl{1} &\qw      &    &&=&&|0\rangle &&\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH &\qSi &\ctrl{1} &\qH &\meter &\cw \cwx \bullet &    \\
+      &\targ    &\qw && &&\qw      &\targ    &\qw      &\qw && &&          &&\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw      &\qw &\qw  &\targ    &\qw &\qw    &\qw              &\qw \\
+    }
+  }
+  \caption{
+	Toffoli gate from control merging and un-merging with T-cost 4.
+	Equivalent to the construction from [[[[cite]]]
+  }
+  \label{fig:ancilla-temporary-toffoli}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
+      &\ctrl{2} &\qw &\qw &\qw    &\qw &\qw &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
+      &         &    &    &       &    &    &         &    &&=&&         &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw      &    & \\
+      &\targ    &\qw &    &\ldots &    &    &\targ    &\qw && &&\qw      &\targ    &\qw &    &\ldots &    &    &\targ    &\qw      &\qw & \\
+    }
+  }
+  \caption{
+	Toffoli gate pair from control merging and un-merging with T-cost 4.
+	Equivalent to the construction from [[[[cite]]].
+  }
+  \label{fig:ancilla-temporary-toffoli}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &       &&&     &\ctrl{1} &\ctrl{1} &\qw  &&&   & & & \lstick{x} &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\rstick{x} \\
+      &\ctrl{1} &\qw & & \cong &&&     &\ctrl{2} &\qS      &\qw  &&& = & & & \lstick{y} &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\rstick{y} \\
+      &\targ    &\qw & &       &&&     &\targ    &\qw      &\qw  &&&   & & & \lstick{t} &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\rstick{t \oplus (x \land y)} \\
+      &         &    & &       &&& \qO &\targ    &\qw      &\qw  &&&   & & & \qA        &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qw &\rstick{x \land y} \\
+    }
   }
@@
 \begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x  &&\ctrl{1} &\qw &x && && \qw  &\qw &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw &           \\
+      &y  &&\ctrl{1} &\qw &y &&=&& \qw  &\qw &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw &           \\
+      &xy &&\qw      &    &  && && \qSi &\qH &\targ    &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qw &|A \rangle \\
+    }
+  }
+  \caption{
+	A non-optimal unmerging circuit that uses three $T$ gates, but recovers an $|A\rangle$ state.
+	Avoids measurement, but has a net T-cost of two (which is worse than \autoref{fig:unmerge-free}'s T-cost of zero).
+  }
+  \label{fig:unmerge-recover-a}
+\end{figure}
+
+\begin{figure}
   \centering
@@
 \begin{figure}
-  \centering
-  \makebox[\linewidth]{
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &    &\ctrl{2}          &\qw               &\qw &&     &&&     &\qS      &\qS      &\gate{Z}     &\ctrl{2}          &\qw               &\qw &&& &&& \lstick{x} &\ctrl{2} &\qw      &\qw &\ctrl{3} &\qw  &\qw      &\qw &\qw      &\qw  &\qw      &\qw &\rstick{x}         \\
+      &    &\qw               &\ctrl{1}          &\qw &&\cong&&&     &\ctrl{-1}&\qS \qwx &\gate{Z}\qwx &\qw               &\ctrl{1}          &\qw &&&=&&& \lstick{y} &\qw      &\qw      &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{1} &\qw &\rstick{y}         \\
+      &    &\multigate{1}{+1} &\multigate{1}{+1} &\qw &&     &&&     &\qw      &\ctrl{-1}&\qw          &\multigate{1}{+1} &\multigate{1}{+1} &\qw &&& &&& \lstick{z} &\targ    &\ctrl{1} &\qw &\qw      &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\targ    &\qw &\rstick{(x+y+z)_0} \\
+      &\qO &\ghost{+1}        &\ghost{+1}        &\qw &&     &&& \qO &\qw      &\qw      &\ctrl{-2}    &\ghost{+1}        &\ghost{+1}        &\qw &&& &&& \qA        &\qw      &\targ    &\qT &\targ    &\qSi &\targ    &\qT &\targ    &\qTi &\qH      &\qw &\rstick{(x+y+z)_1} \\
+    }
   }
   \caption{
-	Temporary addition construction.
+	Temporary full adder.
   }
@@
 
+https://arxiv.org/pdf/1706.05113.pdf   has a ctrl-add with 21N + O(1). This technique+commutators achieves 8N.
+
+https://pdfs.semanticscholar.org/8b84/d1bc2928922937a16205f9d8c925ff76689b.pdf     has a pretty good ripple carry
+
 \end{document}

### 2017-08-23 paper.tex
Hints of structure appearing

@@
 
-\section{Introduction} \label{sec:introduction}
+\section{Introduction}
+\label{sec:introduction}
 
-The textbook construction of a Toffoli gate uses seven T gates [[[cite textbook]]] \autoref{fig:textbook-toffoli} and eight Clifford gates.
-On an error corrected quantum computer that uses the surface code, T gates are XXX orders of magnitude more expensive than Clifford gates [[[cite]]].
+A common construction technique used by quantum circuits is {\em uncomputation}.
+For example, suppose we wish to phase the amplitude of each computational basis state $|k\rangle$ of some register $R_{\text{in}}$ by a number of radians computed by some function $f(k)$.
+A naive method to perform this task is to 1) construct a classical reversible circuit $C$ for approximating $f$, 2) decompose the gates used by $C$ into a quantum gate set so that the circuit can be run on a quantum computer, 3) prepare a fresh register $R_\text{out}$ and use the circuit $C$ on it and $R_\text{in}$ to produce a binary fixed-point 2s-complement output stored in $R_{\text{out}}$, and then 4) phase by an amount proportional to $R_\text{out}$ by applying a $Z^{\theta 2^r}$ operation with appropriate scaling factor $\theta$ to each bit position $r$ within $R_\text{out}$.
+
+The problem with this naive method is that it produces garbage.
+The register $R_\text{out}$ ends up entangled with $R_\text{in}$.
+If the garbage isn't cleaned up, this decoheres $R_\text{in}$; defeating the purpose of using a quantum computer.
+The method we use to clean up the garbage is to uncompute $R_\text{out}$ by running $C$ in reverse.
+This destroys the garbage, without undoing the phasing effects.
+
+Because uncomputation is so common, it is important to understand how to perform it efficiently.
+In this paper, our main focus will be to optimize the uncomputation of Toffoli gates.
+Even more specifically, we focus on surface-code-based quantum computation, where the dominant cost of any circuit is the number of T gates.
+As such, we focus on optimizing the number of T gates needed to compute and uncompute Toffoli gates and then on applying these optimizations elsewhere.
+
+The paper is divided into sections.
+In \autoref{sec:introduction}, we introduce.
+In \autoref{sec:review}, we discuss previous work on optimizing T gate counts and note that they use 8 T gates to compute and uncompute a Toffoli gate.
+In ????, we show how to improve the T-count from 8 to 6 by investing A states.
+In ????, we show how erasing further reduces T-cost from 6 to 4.
+In ????, we review our constructions and demonstrate how they significantly improve the T-cost of several basic arithmetic tasks.
+Finally, ???? concludes and talks about the future.
+
+On a surface-code-based quantum computer, T gates are XXX orders of magnitude more expensive than Clifford gates [[[cite]]].
 Because of this disparity, optimizing error corrected quantum circuits largely consists of reducing their T-cost (the number of T gates used, or rather the number of $|A\rangle = \frac{1}{\sqrt{2}} |0\rangle + \frac{1+i}{2} |1\rangle$ states which must be distilled in order to perform those T gates).
+
+\section{Previous Work}
+\label{sec:review}
+
+The textbook construction of a Toffoli gate uses seven T gates \cite{Nielsen2009} (see \autoref{fig:textbook-toffoli}).
 
@@
   \caption{
-	Textbook Toffoli construction [[[[cite]]]] from eight Clifford gates and seven T gates.
+	Textbook Toffoli construction from \cite{Nielsen2009}.
+	Uses eight Clifford gates and seven T gates.
   }
@@
 
-The textbook Toffoli construction is optimal in some highly constrained situations (e.g.
-not allowed to use any ancilla qubits and not allowed to merge T gates across operations [[[cite]]]), but is not optimal in general.
-For example, when two Toffoli gates share a control, applying the textbook construction to each places two $T$ gates on the shared control.
-These two $T$ gates can be merged into a single $S$ gate, thereby reducing the T-cost of each Toffoli from 7 to 6.
+We can see why this construction works by considering which computation basis states are phased when applying a $Z$ rotation to a qubit temporarily storing a parity such as $x \oplus y$.
+We start with an 8-dimensional vector $[0, 0, 0, 0, 0, 0, 0, 0]$ representing the amount of phase added into each basis state.
+Applying a $Z^t$ rotation to a qubit will add $t$ into half of the vector entries (which four entries are increased depents on the value the qubit is temporarily storing).
+W.l.o.g.
+we order the vector so that applying $Z^{t}$ to qubit \#1 adds $t \cdot [0, 1, 0, 1, 0, 1, 0, 1]$ into the phase-accumulation vector, applying $Z^{t}$ to qubit \#2 adds $t \cdot [0, 0, 1, 1, 0, 0, 1, 1]$, and applying $Z^{t}$ to qubit \#3 adds $t \cdot [0, 0, 0, 0, 1, 1, 1, 1]$.
 
-More commonly, Toffoli gates tend to come in matched pairs where one is uncomputing the effect of the other.
-When this occurs, the three $T$ gates on the control qubits of the textbook construction are no longer necessary \autoref{fig:bad-phase-toffoli}.
-Removing those $T$ gates introduces phase errors, but the second Toffoli gate will uncompute those errors while uncomputing the state permutation.
+If we temporarily CNOT qubit \#1 into qubit \#2 and phase qubit \#2, the corresponding vector is $[0, 1, 1, 0, 0, 1, 1, 0]$.
+Given this ordering and the parity rule, we can prepare a table of vectors corresponding to phasing qubits storing various additions:
+
+$$
+\begin{aligned}
+\text{target}             &\rightarrow & \text{phase vector} \\
+\hline \\
+q_1                       &\rightarrow &[0, 1, 0, 1, 0, 1, 0, 1] \\
+q_2                       &\rightarrow &[0, 0, 1, 1, 0, 0, 1, 1] \\
+q_3                       &\rightarrow &[0, 0, 0, 0, 1, 1, 1, 1] \\
+q_1 \oplus q_2            &\rightarrow &[0, 1, 1, 0, 0, 1, 1, 0] \\
+q_1 \oplus q_3            &\rightarrow &[0, 1, 0, 1, 1, 0, 1, 0] \\
+q_2 \oplus q_3            &\rightarrow &[0, 0, 1, 1, 1, 1, 0, 0] \\
+q_1 \oplus q_2 \oplus q_3 &\rightarrow &[0, 1, 1, 0, 1, 0, 0, 1] \\
+\text{(global)}           &\rightarrow &[1, 1, 1, 1, 1, 1, 1, 1] \\
+\end{aligned}
+$$
+
+Note that this set of vectors spans the full eight dimensional space.
+Any desired phase function can be implemented by rotating parities.
+(And in general this continues to be true for any number of qubits.)
+For the specific case of the CCZ gate, we wish to perform $[0, 0, 0, 0, 0, 0, 0, 1]$.
+We do so by adding the odd-number-of-qubit parities and subtracting the even-number-of-qubit parities, all times a quarter.
+This is what the circuit construction is doing, but after framing the target qubit with Hadamards to switch the axis of interaction from Z to X so that a conditional NOT is performed instead of a conditional phase-flip.
+
+Assuming we aren't permitted to move work to other qubits or to share work with other operations, the textbook Toffoli construction is optimal \cite{Gosset2014}.
+Of course, in practice, we can do both.
+For example, when several adjacent Toffoli gates share the same controls, all but one can be replaced by toggle-propagating CNOT operations (see \autoref{fig:shared-controls}).
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1}&\ctrl{1}&\ctrl{1}&\ctrl{1}&\qw & &   & & &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\\
+      &\ctrl{1}&\ctrl{2}&\ctrl{3}&\ctrl{4}&\qw & & = & & &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\\
+      &\targ   &\qw     &\qw     &\qw     &\qw & &   & & &\ctrl{3}&\ctrl{2}&\ctrl{1}&\targ   &\ctrl{1}&\ctrl{2}&\ctrl{3}&\\
+      &\qw     &\targ   &\qw     &\qw     &\qw & &   & & &\qw     &\qw     &\targ   &\qw     &\targ   &\qw     &\qw     &\\
+      &\qw     &\qw     &\targ   &\qw     &\qw & &   & & &\qw     &\targ   &\qw     &\qw     &\qw     &\targ   &\qw     &\\
+      &\qw     &\qw     &\qw     &\targ   &\qw & &   & & &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\\
+    }
+  }
+  \caption{
+	The T-cost of $N$ adjacent Toffolis sharing the same controls is $0 \cdot N + O(1)$.
+	Each additional Toffoli can be replaced with CNOTs framing a root Toffoli.
+  }
+  \label{fig:shared-controls}
+\end{figure}
+
+It is not common for adjacent Toffolis to have the same controls, but it is common for a Toffoli to later be uncomputed by a second matching Toffoli (i.e. for the Toffoli to be {\em temporary}).
+When this occurs, the three $T$ gates on the control qubits of the textbook construction can be omitted.
+This introduces phase errors (see \autoref{fig:bad-phase-toffoli}), but the second Toffoli gate will uncompute those errors while uncomputing the state permutation (see \autoref{fig:cancelled-bad-phase-toffoli}) \cite{Barenco1995}.
 This technique reduces the T-cost of paired Toffolis from 7 to 4.
@@
   \caption{
-	Existing temporary Toffoli construction with 4 T gates.
+	Starting with \autoref{fig:textbook-toffoli} then dropping T gates on the controls produces an operation with a T-cost of 4 that performs the correct permutation.
+	However, the operation introduces phase errors.
   }
@@
 
-In [[[cite]]], Jones improves on the paired construction by demonstrating how to reduce the T-cost of any Toffoli to 4.
-Jones does this by using a temporary ancilla and applying an erasure technique. [[[figures]]]
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw && &&&\ctrl{1}     &\ctrl{1}  &\qw &\qw &\qw    &\qw &\qw     &\ctrl{1}&\ctrl{1}     &\qw & \\
+      &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw &&=&&&\qS          &\ctrl{1}  &\qw &\qw &\qw    &\qw &\qw     &\ctrl{1}&\qSi         &\qw & \\
+      &\targ   &\qw &    &\ldots &    &    &\targ   &\qw && &&&\gate{Z}\qwx &\targ     &\qw &    &\ldots &    &        &\targ   &\gate{Z}\qwx &\qw & \\
+    }
+  }
+  \caption{
+	When two Toffolis form a compute/uncompute pair, they cancel phase errors in underlying constructions.
+	Fixes the problem with the construction in \autoref{fig:bad-phase-toffoli}, achieving a per-Toffoli T-cost of 4 for paired Toffolis \cite{Barenco1995}.
+  }
+  \label{fig:cancelled-bad-phase-toffoli}
+\end{figure}
 
-We improve on Jones by separating their construction into three pieces (merge, use, erase), and showing how to apply these pieces to circuit construction tasks.
-In particular, we reduce the average T-cost of paired Toffolis from 4 to 2.
-Many common circuits (e.g. addition) have paired Toffolis, and so this optimization can halve their T-cost.
-
-\section{Temporary ANDs}
-
-See \autoref{fig:inline-toffoli}.
+In \cite{Jones2013}, Jones shows how to perform an unpaired Toffoli with a T-cost of 4 by using an ancilla and an erasure technique that replaces one of the quantum controls on a Toffoli with a classical control.
+See \autoref{fig:jones-toffoli}.
+Jones also shows how to rearrange all of the T gates into the same column, so that the Toffoli construction has a T-depth of 1 (see figure?).
 
@@
   }
-  \label{fig:ancilla-toffoli}
+  \label{fig:jones-toffoli}
 \end{figure}
+
+We improve on Jones by noting that their construction separates cleanly into three pieces (merge, use, erase), using these pieces to reduce the average T-cost of paired Toffolis from 4 to 2, and then applying this improvement to common circuits.
+
+\section{AND-ing and Erasing}
+\label{sec:temporary-and}
+
+When the target of a Toffoli gate is a qubit in the $|0\rangle$ state, and the controls are in a computation basis state, the target qubit ends up storing the logical-and of the two controls.
+We continue to say that the target stores the logical-and of the two controls even when they are under superposition, with the understanding that we mean ``within each computation basis state making up the superposition of the overall system".
+
+In the textbook construction of the Toffoli gate, the three T gates on the control wires are implementing a controlled-S.
+A controlled-S phases states by $i$ where both the control and the target are ON.
+If the target of the Toffoli gate starts in the $|0\rangle$ state, then the target ends up ON when both controls are satisfied.
+Therefore the target is ON in exactly the state where a phase factor of $i$ should be applied, and the controlled-S on the controls can be replaced by an unconditional S gate on the target after the Toffoli is finished.
+
+A minor optimization which we apply is to note that the first two operations on the target are transitioning from the $|0\rangle$ state to the $|A\rangle$ state.
+Instead of consuming an $|A\rangle$ state to perform the $T$ gate (and potentially performing a fixup S gate), we simply feed the $|A\rangle$ state we would have consumed directly into the circuit.
+
+This construction defines our logical AND operation, which we draw as a wire emerging from two controls as shown in the left hand side of \autoref{fig:logical-and} (which also shows our construction for the operation).
 
@@
   }
-  \label{fig:ancilla-temporary-toffoli}
+  \label{fig:logical-and}
 \end{figure}
+
+The logical AND construction consumes four $|A\rangle$ states, so its T-cost is 4.
+
+To erase the ancilla produced by the logical AND, we use a technique we refer to as ``erasure".
+We note that a Toffoli gate would obviously clear the value.
+Then, since the target qubit can be discarded after clearing it, we are free to introduce a Hadamard gate and a measurement before discarding it.
+We then hop the Hadamard over the Toffoli, transforming it into a CCZ operation.
+The controls and targets of a CCZ are equivalent, so we rewrite the circuit into a CZC.
+We then hop the Measurement over the CZC, turning a quantum control into a classical control.
+These circuit moves turned the Toffoli gate into a CZ gate that we apply or not based on the outcome of the measreument
+We ``Clifford-ized" the circuit by adding a Hadamard gate and a measurement.
+
+We represent erasing a logical AND with a wire terminating into two controls, as shown in \autoref{fig:erase-and} which also shows the underlying construction.
+The T-cost of erasure is 0.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x  &&\ctrl{1} &\qw &x && &&\qw &\qw    &\ctrl{1} &\qw \\
+      &y  &&\ctrl{1} &\qw &y &&=&&\qw &\qw    &\gate{Z} &\qw \\
+      &xy &&\qw      &    &  && &&\qH &\meter &\cw \cwx \bullet &    \\
+    }
+  }
+  \caption{
+	Control un-merging with T-cost of 0.
+  }
+  \label{fig:erase-and}
+\end{figure}
+
+The benefit of separating the logical-and-creating operation from the logical-and-erasing operation is that, as long as we have the ancilla storing the logical AND around, we can Cliffordize any Toffoli gate whose controls would be the two qubits we merged by replacing the Toffoli with a CNOT with the ancilla qubit as a control.
+When the parts are simply placed next to each other, we end up with a construction exactly equivalent to the ones presented by Jones \cite{Jones2013}.
+See \autoref{fig:merge-use-erase-toffoli}.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw  &\qw      &\qw &\qw    &\ctrl{1}         &\qw \\
+      &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\qw  &\qw      &\qw &\qw    &\gate{Z}         &\qw \\
+      &         &    &&=&&         &\ctrl{1} &\qw      &    &&=&&|0\rangle &&\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH &\qSi &\ctrl{1} &\qH &\meter &\cw \cwx \bullet &    \\
+      &\targ    &\qw && &&\qw      &\targ    &\qw      &\qw && &&          &&\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw      &\qw &\qw  &\targ    &\qw &\qw    &\qw              &\qw \\
+    }
+  }
+  \caption{
+	Performing a Toffoli gate by computing, using, and erasing a logical-AND.
+	Has a T-cost of 4, and is exactly equivalent to the construction from \cite{Jones2013}.
+  }
+  \label{fig:merge-use-erase-toffoli}
+\end{figure}
+
+In order to make savings, we need to keep the ancilla around for longer.
+We need a situation like the one shown in \autoref{fig:paired-toffoli-to-logical-and}, where a single logical-AND computation replaces two Toffolis.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
+      &\ctrl{2} &\qw &\qw &\qw    &\qw &\qw &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
+      &         &    &    &       &    &    &         &    &&=&&         &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw      &    & \\
+      &\targ    &\qw &    &\ldots &    &    &\targ    &\qw && &&\qw      &\targ    &\qw &    &\ldots &    &    &\targ    &\qw      &\qw & \\
+    }
+  }
+  \caption{
+	When distant Toffolis share the same controls, a single logical-AND can Cliffordize both.
+    The T-cost of computing and uncomputing the logical-AND is 4, so in effect this reduces the average T-cost of paired Toffolis to 2.
+  }
+  \label{fig:paired-toffoli-to-logical-and}
+\end{figure}
+
+Fortunately, this pattern of distant Toffolis sharing controls is very common in quantum circuits.
+It occurs during uncomputation, and also in circuits that sweep back and forth across qubits like addition.
+
+\section{T-optimized Circuit Constructions}
+\label{sec:circuit-constructions}
 
@@
 \begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &x  &&\ctrl{1} &\qw &x && &&\qw &\qw    &\ctrl{1} &\qw \\
-      &y  &&\ctrl{1} &\qw &y &&=&&\qw &\qw    &\gate{Z} &\qw \\
-      &xy &&\qw      &    &  && &&\qH &\meter &\cw \cwx \bullet &    \\
-    }
-  }
-  \caption{
-	Control un-merging with T-cost of 0.
-  }
-  \label{fig:unmerge-free}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw  &\qw      &\qw &\qw    &\ctrl{1}         &\qw \\
-      &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\qw  &\qw      &\qw &\qw    &\gate{Z}         &\qw \\
-      &         &    &&=&&         &\ctrl{1} &\qw      &    &&=&&|0\rangle &&\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH &\qSi &\ctrl{1} &\qH &\meter &\cw \cwx \bullet &    \\
-      &\targ    &\qw && &&\qw      &\targ    &\qw      &\qw && &&          &&\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw      &\qw &\qw  &\targ    &\qw &\qw    &\qw              &\qw \\
-    }
-  }
-  \caption{
-	Toffoli gate from control merging and un-merging with T-cost 4.
-	Equivalent to the construction from [[[[cite]]]
-  }
-  \label{fig:ancilla-temporary-toffoli}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
-      &\ctrl{2} &\qw &\qw &\qw    &\qw &\qw &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
-      &         &    &    &       &    &    &         &    &&=&&         &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw      &    & \\
-      &\targ    &\qw &    &\ldots &    &    &\targ    &\qw && &&\qw      &\targ    &\qw &    &\ldots &    &    &\targ    &\qw      &\qw & \\
-    }
-  }
-  \caption{
-	Toffoli gate pair from control merging and un-merging with T-cost 4.
-	Equivalent to the construction from [[[[cite]]].
-  }
-  \label{fig:ancilla-temporary-toffoli}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &\ctrl{1} &\qw & &       &&&     &\ctrl{1} &\ctrl{1} &\qw  &&&   & & & \lstick{x} &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\rstick{x} \\
-      &\ctrl{1} &\qw & & \cong &&&     &\ctrl{2} &\qS      &\qw  &&& = & & & \lstick{y} &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\rstick{y} \\
-      &\targ    &\qw & &       &&&     &\targ    &\qw      &\qw  &&&   & & & \lstick{t} &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\rstick{t \oplus (x \land y)} \\
-      &         &    & &       &&& \qO &\targ    &\qw      &\qw  &&&   & & & \qA        &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qw &\rstick{x \land y} \\
-    }
-  }
-  \caption{
-	Our temporary Toffoli construction, which uses an $|A\rangle$ ancilla and 3 T gates.
-	The $|A\rangle$ state is recovered when uncomputing the temporary Toffoli.
-  }
-  \label{fig:ancilla-temporary-toffoli}
-\end{figure}
-
-\begin{figure}
   \centering
@@
   \label{fig:inline-addition}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &x  &&\ctrl{1} &\qw &x && && \qw  &\qw &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw &           \\
-      &y  &&\ctrl{1} &\qw &y &&=&& \qw  &\qw &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw &           \\
-      &xy &&\qw      &    &  && && \qSi &\qH &\targ    &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qw &|A \rangle \\
-    }
-  }
-  \caption{
-	A non-optimal unmerging circuit that uses three $T$ gates, but recovers an $|A\rangle$ state.
-	Avoids measurement, but has a net T-cost of two (which is worse than \autoref{fig:unmerge-free}'s T-cost of zero).
-  }
-  \label{fig:unmerge-recover-a}
 \end{figure}
@@
 
+\section{Investing $|A\rangle$ states}
+\label{sec:invest}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &       &&&     &\ctrl{1} &\ctrl{1} &\qw  &&&   & & & \lstick{x} &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\rstick{x} \\
+      &\ctrl{1} &\qw & & \cong &&&     &\ctrl{2} &\qS      &\qw  &&& = & & & \lstick{y} &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\rstick{y} \\
+      &\targ    &\qw & &       &&&     &\targ    &\qw      &\qw  &&&   & & & \lstick{t} &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\rstick{t \oplus (x \land y)} \\
+      &         &    & &       &&& \qO &\targ    &\qw      &\qw  &&&   & & & \qA        &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qw &\rstick{x \land y} \\
+    }
+  }
+  \caption{
+	Our temporary Toffoli construction, which uses an $|A\rangle$ ancilla and 3 T gates.
+	The $|A\rangle$ state is recovered when uncomputing the temporary Toffoli.
+  }
+  \label{fig:ancilla-temporary-toffoli}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x  &&\ctrl{1} &\qw &x && && \qw  &\qw &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw &           \\
+      &y  &&\ctrl{1} &\qw &y &&=&& \qw  &\qw &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw &           \\
+      &xy &&\qw      &    &  && && \qSi &\qH &\targ    &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qw &|A \rangle \\
+    }
+  }
+  \caption{
+	A non-optimal unmerging circuit that uses three $T$ gates, but recovers an $|A\rangle$ state.
+	Avoids measurement, but has a net T-cost of two (which is worse than \autoref{fig:unmerge-free}'s T-cost of zero).
+  }
+  \label{fig:unmerge-recover-a}
+\end{figure}
+
 \section{Acknowledgements}
@@
 
-\cite{barenco1995}
 https://arxiv.org/pdf/quant-ph/9503016.pdf   ->   the paired Toffoli construction that uses 4T

### 2017-08-23 paper.tex
Almost starting to feel coherent

@@
 \begin{abstract}
-We show how to compute the logical-and of two controls using four T gates, and uncompute it using no T gates.
-This construction halves the average T-cost of temporary Toffoli gates.
-A temporary Toffoli gate is one which will be uncomputed later (typically by a second Toffoli gate).
+A ``temporary Toffoli" is a doubly-controlled NOT operation that will later be uncomputed by a second double-controlled NOT.
+We construct temporary Toffoli gates using four T gates; half as many as previously reported.
+We demonstrate how to apply our construction (and other optimizations) to common operations and algorithms, and show significant improvement over previously reported T-costs.
+For example we cut the reported T-cost of addition by 50\%, the reported T-cost of controlled-addition by 70\%, and the reported T-cost of Shor's algorithm by ?????\%.
 
-We use our temporary logical-and construction to produce quantum circuits with T gate counts lower than previously reported.
-T gates dominate the cost of surface-code-based error corrected quantum computation.
-Temporary Toffolis are used in many common quantum circuits, such as addition.
-As such, the optimization we present in this paper are widely applicable and represent significant savings in the projected costs of quantum computation.
-For example, by applying previously known optimizations and the ones presented in this paper, we improve on the projected T-cost of Shor's algorithm from [[[[cite]]] by 70\% and the projected minimum time down by 66%.
-
-We demonstrate an $n$-bit adder with T-cost $4n + O(1)$, a controlled $n$-bit adder with T-cost $8n + O(1)$, an $n$-controlled NOT with T-cost $4n + O(1)$, and a temporary pop-count with T-cost $6n + O(1)$.
-
-We discuss investing $|A\rangle$ states, a fallback optimization technique that may reduce the number of T gates needed to compute quantum circuits not amenable to temporary ANDs.
+T gates dominate the cost of surface-code-based quantum computation.
+Temporary Toffolis are a basic building block that applies to many quantum circuits.
+As such, a temporary Toffoli gate that uses fewer T gates represents a widely applicable and significant improvement to the projected costs of quantum computation.
 \end{abstract}
@@
 
-A common construction technique used by quantum circuits is {\em uncomputation}.
-For example, suppose we wish to phase the amplitude of each computational basis state $|k\rangle$ of some register $R_{\text{in}}$ by a number of radians computed by some function $f(k)$.
-A naive method to perform this task is to 1) construct a classical reversible circuit $C$ for approximating $f$, 2) decompose the gates used by $C$ into a quantum gate set so that the circuit can be run on a quantum computer, 3) prepare a fresh register $R_\text{out}$ and use the circuit $C$ on it and $R_\text{in}$ to produce a binary fixed-point 2s-complement output stored in $R_{\text{out}}$, and then 4) phase by an amount proportional to $R_\text{out}$ by applying a $Z^{\theta 2^r}$ operation with appropriate scaling factor $\theta$ to each bit position $r$ within $R_\text{out}$.
+The surface code is a quantum error correcting code that works on a 2D nearest-neighbour array of qubits and achieves a threshold error rate of approximately 1\% [[[[cite so much]]]].
+This makes the surface code a likely component in the architecture of large error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies [[[[cite so much]]]] and other known error correcting codes have lower thresholds or require higher connectivity [[[[cite so much]]]].
 
-The problem with this naive method is that it produces garbage.
-The register $R_\text{out}$ ends up entangled with $R_\text{in}$.
-If the garbage isn't cleaned up, this decoheres $R_\text{in}$; defeating the purpose of using a quantum computer.
-The method we use to clean up the garbage is to uncompute $R_\text{out}$ by running $C$ in reverse.
-This destroys the garbage, without undoing the phasing effects.
+One of the downsides of the surface code is that it has no cheap mechanism to apply non-Clifford operations such as $T$ gates.
+Instead, $T$ gates are performed by distilling and consuming $|A\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle$ states.
+Consuming an $|A\rangle$ state to perform a T gate is simple (see \autoref{fig:t-equals-a}), but distilling $|A\rangle$ states has significant cost.
 
-Because uncomputation is so common, it is important to understand how to perform it efficiently.
-In this paper, our main focus will be to optimize the uncomputation of Toffoli gates.
-Even more specifically, we focus on surface-code-based quantum computation, where the dominant cost of any circuit is the number of T gates.
-As such, we focus on optimizing the number of T gates needed to compute and uncompute Toffoli gates and then on applying these optimizations elsewhere.
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &\gate{T}&\qw &&=&&&          &&\qw &\ctrl{1}&\qw    &\qS            &\qw &\\
+      &        &    && &&&|A\rangle &&\qw &\targ   &\meter &\cw\cwx\bullet &    &\\
+    }
+  }
+  \caption{
+	Performing a T gate with Clifford operations by consuming an $|A\rangle$ state, where $|A\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle$.
+  }
+  \label{fig:t-equals-a}
+\end{figure}
 
-The paper is divided into sections.
-In \autoref{sec:introduction}, we introduce.
-In \autoref{sec:review}, we discuss previous work on optimizing T gate counts and note that they use 8 T gates to compute and uncompute a Toffoli gate.
-In ????, we show how to improve the T-count from 8 to 6 by investing A states.
-In ????, we show how erasing further reduces T-cost from 6 to 4.
-In ????, we review our constructions and demonstrate how they significantly improve the T-cost of several basic arithmetic tasks.
-Finally, ???? concludes and talks about the future.
+In [[[[cite]]]] the $|A\rangle$ state distillation is done by preparing an EPR pair $(P, Q)$, moving $P$ out of the surface code and into a different error correcting code that can perform $T$ gates, and using a $T$ gate within that code to rotate and measure $P$ along an axis correlated with the diagonal $X+Y$ axis of $Q$.
+If this process succeeds without an error being detected, it steers \cite{Wiseman2007} $Q$ into the $|A\rangle$ state or else into the complementary state $Z|A\rangle$.
+The $Z$ rotation is easily corrected, and the result of measuring $P$ indicates whether the correction is needed or not.
 
-On a surface-code-based quantum computer, T gates are XXX orders of magnitude more expensive than Clifford gates [[[cite]]].
-Because of this disparity, optimizing error corrected quantum circuits largely consists of reducing their T-cost (the number of T gates used, or rather the number of $|A\rangle = \frac{1}{\sqrt{2}} |0\rangle + \frac{1+i}{2} |1\rangle$ states which must be distilled in order to perform those T gates).
+The resulting `foothold' $|A\rangle$ states have a non-negligible amount of noise, but can be used to power T gates for rounds of distillation occurring entirely within the surface code.
+Each round of distillation reduces the amount of noise in the final states, but at significant cost.
+After accounting for distilling the $|A\rangle$ state, the spacetime volume cost of a T gate is one to two orders of magnitude higher than Clifford operations native to the surface code (such as the CNOT gate) [[[[cite]]]].
+
+Because T gates are so expensive for the surface code, and the surface code is a likely component of future quantum computers, it is important to consider and optimize the number of T gates used by quantum circuits.
+In this paper, we focus on optimizing the number of T gates needed to perform Toffoli gates that will later be uncomputed by a second Toffoli gate (``temporary Toffoli gates").
+Temporary Toffolis show up in many circuits, especially ones that involve computing and uncomputing classical functions, so these optimizations are widely applicable.
+
+Our paper is divided into six sections.
+In \autoref{sec:introduction}, we motivate the problem and insert self-referential descriptions.
+\autoref{sec:review} discusses how previous work on optimizing the T gate counts of Toffolis improved the cost of temporary Toffolis from 14 to 8.
+In \autoref{sec:invest}, we show how to improve the cost of temporary Toffolis from 8 to 6 by investing and later recovering an $|A\rangle$ state instead of consuming it to perform T gates.
+\autoref{sec:temporary-and} explains an ancilla erasing trick that improves the cost of temporary Toffolis further, from 6 to 4.
+This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to significantly improve the T-cost of several basic arithmetic tasks (e.g.
+halving the T-cost of addition).
+Finally, \autoref{sec:conclusion} concludes and discusses future work.
 
@@
 The textbook construction of a Toffoli gate uses seven T gates \cite{Nielsen2009} (see \autoref{fig:textbook-toffoli}).
+Assuming we aren't permitted to involve other qubits or to share work with other operations, this construction is optimal \cite{Gosset2014}.
+Of course, in practice, we can do both.
+For example, when several adjacent Toffoli gates share the same controls, all but one can be replaced by CNOT operations (see \autoref{fig:shared-controls}).
 
@@
 
-We can see why this construction works by considering which computation basis states are phased when applying a $Z$ rotation to a qubit temporarily storing a parity such as $x \oplus y$.
-We start with an 8-dimensional vector $[0, 0, 0, 0, 0, 0, 0, 0]$ representing the amount of phase added into each basis state.
-Applying a $Z^t$ rotation to a qubit will add $t$ into half of the vector entries (which four entries are increased depents on the value the qubit is temporarily storing).
-W.l.o.g.
-we order the vector so that applying $Z^{t}$ to qubit \#1 adds $t \cdot [0, 1, 0, 1, 0, 1, 0, 1]$ into the phase-accumulation vector, applying $Z^{t}$ to qubit \#2 adds $t \cdot [0, 0, 1, 1, 0, 0, 1, 1]$, and applying $Z^{t}$ to qubit \#3 adds $t \cdot [0, 0, 0, 0, 1, 1, 1, 1]$.
-
-If we temporarily CNOT qubit \#1 into qubit \#2 and phase qubit \#2, the corresponding vector is $[0, 1, 1, 0, 0, 1, 1, 0]$.
-Given this ordering and the parity rule, we can prepare a table of vectors corresponding to phasing qubits storing various additions:
-
-$$
-\begin{aligned}
-\text{target}             &\rightarrow & \text{phase vector} \\
-\hline \\
-q_1                       &\rightarrow &[0, 1, 0, 1, 0, 1, 0, 1] \\
-q_2                       &\rightarrow &[0, 0, 1, 1, 0, 0, 1, 1] \\
-q_3                       &\rightarrow &[0, 0, 0, 0, 1, 1, 1, 1] \\
-q_1 \oplus q_2            &\rightarrow &[0, 1, 1, 0, 0, 1, 1, 0] \\
-q_1 \oplus q_3            &\rightarrow &[0, 1, 0, 1, 1, 0, 1, 0] \\
-q_2 \oplus q_3            &\rightarrow &[0, 0, 1, 1, 1, 1, 0, 0] \\
-q_1 \oplus q_2 \oplus q_3 &\rightarrow &[0, 1, 1, 0, 1, 0, 0, 1] \\
-\text{(global)}           &\rightarrow &[1, 1, 1, 1, 1, 1, 1, 1] \\
-\end{aligned}
-$$
-
-Note that this set of vectors spans the full eight dimensional space.
-Any desired phase function can be implemented by rotating parities.
-(And in general this continues to be true for any number of qubits.)
-For the specific case of the CCZ gate, we wish to perform $[0, 0, 0, 0, 0, 0, 0, 1]$.
-We do so by adding the odd-number-of-qubit parities and subtracting the even-number-of-qubit parities, all times a quarter.
-This is what the circuit construction is doing, but after framing the target qubit with Hadamards to switch the axis of interaction from Z to X so that a conditional NOT is performed instead of a conditional phase-flip.
-
-Assuming we aren't permitted to move work to other qubits or to share work with other operations, the textbook Toffoli construction is optimal \cite{Gosset2014}.
-Of course, in practice, we can do both.
-For example, when several adjacent Toffoli gates share the same controls, all but one can be replaced by toggle-propagating CNOT operations (see \autoref{fig:shared-controls}).
-
 \begin{figure}
@@
     \Qcircuit @R=1.5em @C=0.7em {
-      &\ctrl{1}&\ctrl{1}&\ctrl{1}&\ctrl{1}&\qw & &   & & &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\\
-      &\ctrl{1}&\ctrl{2}&\ctrl{3}&\ctrl{4}&\qw & & = & & &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\\
-      &\targ   &\qw     &\qw     &\qw     &\qw & &   & & &\ctrl{3}&\ctrl{2}&\ctrl{1}&\targ   &\ctrl{1}&\ctrl{2}&\ctrl{3}&\\
-      &\qw     &\targ   &\qw     &\qw     &\qw & &   & & &\qw     &\qw     &\targ   &\qw     &\targ   &\qw     &\qw     &\\
-      &\qw     &\qw     &\targ   &\qw     &\qw & &   & & &\qw     &\targ   &\qw     &\qw     &\qw     &\targ   &\qw     &\\
-      &\qw     &\qw     &\qw     &\targ   &\qw & &   & & &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\\
+      &\ctrl{1}&\ctrl{1}&\ctrl{1}&\ctrl{1}&\qw && &&&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw &\\
+      &\ctrl{1}&\ctrl{2}&\ctrl{3}&\ctrl{4}&\qw &&=&&&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw &\\
+      &\targ   &\qw     &\qw     &\qw     &\qw && &&&\ctrl{3}&\ctrl{2}&\ctrl{1}&\targ   &\ctrl{1}&\ctrl{2}&\ctrl{3}&\qw &\\
+      &\qw     &\targ   &\qw     &\qw     &\qw && &&&\qw     &\qw     &\targ   &\qw     &\targ   &\qw     &\qw     &\qw &\\
+      &\qw     &\qw     &\targ   &\qw     &\qw && &&&\qw     &\targ   &\qw     &\qw     &\qw     &\targ   &\qw     &\qw &\\
+      &\qw     &\qw     &\qw     &\targ   &\qw && &&&\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &\\
     }
@@
 	The T-cost of $N$ adjacent Toffolis sharing the same controls is $0 \cdot N + O(1)$.
-	Each additional Toffoli can be replaced with CNOTs framing a root Toffoli.
+	The marginal T-cost is 0 because each additional Toffoli can be replaced with CNOTs framing a root Toffoli.
   }
@@
 
-It is not common for adjacent Toffolis to have the same controls, but it is common for a Toffoli to later be uncomputed by a second matching Toffoli (i.e. for the Toffoli to be {\em temporary}).
+It is not common for adjacent Toffolis to have the same controls, but it is common for a Toffoli to later be uncomputed by a second matching Toffoli (i.e. for the Toffoli's effect to be temporary).
 When this occurs, the three $T$ gates on the control qubits of the textbook construction can be omitted.
@@
 
-In \cite{Jones2013}, Jones shows how to perform an unpaired Toffoli with a T-cost of 4 by using an ancilla and an erasure technique that replaces one of the quantum controls on a Toffoli with a classical control.
-See \autoref{fig:jones-toffoli}.
+In \cite{Jones2013}, Jones shows how to perform an Toffoli with a T-cost of 4 even when not pairs.
+They do so by involving an ancilla qubit and an erasure technique that turns what would have been a quantum control into a classical control.
+We show this construction in \autoref{fig:jones-toffoli}.
 Jones also shows how to rearrange all of the T gates into the same column, so that the Toffoli construction has a T-depth of 1 (see figure?).
@@
 
-We improve on Jones by noting that their construction separates cleanly into three pieces (merge, use, erase), using these pieces to reduce the average T-cost of paired Toffolis from 4 to 2, and then applying this improvement to common circuits.
+By using Jones' construction twice (or the phase-correct-by-pairing construction), a temporary Toffoli can be computed with eight T gates.
+Four T gates for the initial Toffoli, and four T gates to uncompute its effect later with another Toffoli.
+
+\section{Investing $|A\rangle$ states}
+\label{sec:invest}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw & &       &&&     &\ctrl{1} &\ctrl{1} &\qw  &&&   & & & \lstick{x} &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\rstick{x} \\
+      &\ctrl{1} &\qw & & \cong &&&     &\ctrl{2} &\qS      &\qw  &&& = & & & \lstick{y} &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\rstick{y} \\
+      &\targ    &\qw & &       &&&     &\targ    &\qw      &\qw  &&&   & & & \lstick{t} &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\rstick{t \oplus (x \land y)} \\
+      &         &    & &       &&& \qO &\targ    &\qw      &\qw  &&&   & & & \qA        &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qw &\rstick{x \land y} \\
+    }
+  }
+  \caption{
+	Our temporary Toffoli construction, which uses an $|A\rangle$ ancilla and 3 T gates.
+	The $|A\rangle$ state is recovered when uncomputing the temporary Toffoli.
+  }
+  \label{fig:ancilla-temporary-toffoli}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x  &&\ctrl{1} &\qw &x && && \qw  &\qw &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw &           \\
+      &y  &&\ctrl{1} &\qw &y &&=&& \qw  &\qw &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw &           \\
+      &xy &&\qw      &    &  && && \qSi &\qH &\targ    &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qw &|A \rangle \\
+    }
+  }
+  \caption{
+	A non-optimal unmerging circuit that uses three $T$ gates, but recovers an $|A\rangle$ state.
+	Avoids measurement, but has a net T-cost of two (which is worse than \autoref{fig:unmerge-free}'s T-cost of zero).
+  }
+  \label{fig:unmerge-recover-a}
+\end{figure}
 
@@
 
-\section{Investing $|A\rangle$ states}
-\label{sec:invest}
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &\ctrl{1} &\qw & &       &&&     &\ctrl{1} &\ctrl{1} &\qw  &&&   & & & \lstick{x} &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\rstick{x} \\
-      &\ctrl{1} &\qw & & \cong &&&     &\ctrl{2} &\qS      &\qw  &&& = & & & \lstick{y} &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\rstick{y} \\
-      &\targ    &\qw & &       &&&     &\targ    &\qw      &\qw  &&&   & & & \lstick{t} &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\rstick{t \oplus (x \land y)} \\
-      &         &    & &       &&& \qO &\targ    &\qw      &\qw  &&&   & & & \qA        &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qw &\rstick{x \land y} \\
-    }
-  }
-  \caption{
-	Our temporary Toffoli construction, which uses an $|A\rangle$ ancilla and 3 T gates.
-	The $|A\rangle$ state is recovered when uncomputing the temporary Toffoli.
-  }
-  \label{fig:ancilla-temporary-toffoli}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &x  &&\ctrl{1} &\qw &x && && \qw  &\qw &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw &           \\
-      &y  &&\ctrl{1} &\qw &y &&=&& \qw  &\qw &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw &           \\
-      &xy &&\qw      &    &  && && \qSi &\qH &\targ    &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qw &|A \rangle \\
-    }
-  }
-  \caption{
-	A non-optimal unmerging circuit that uses three $T$ gates, but recovers an $|A\rangle$ state.
-	Avoids measurement, but has a net T-cost of two (which is worse than \autoref{fig:unmerge-free}'s T-cost of zero).
-  }
-  \label{fig:unmerge-recover-a}
-\end{figure}
-
 \section{Acknowledgements}
@@
 
+As an example, suppose we wish to phase the amplitude of each computational basis state $|k\rangle$ of some register $R_{\text{in}}$ by a number of radians computed by some function $f(k)$.
+We want to apply the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$.
+A naive method to perform this task is to 1) construct a classical reversible circuit $C$ for approximating $f$, 2) decompose the gates used by $C$ into a quantum gate set so that the circuit can be run on a quantum computer, 3) prepare a fresh register $R_\text{out}$ and use the circuit $C$ on it and $R_\text{in}$ to produce a binary fixed-point 2s-complement output stored in $R_{\text{out}}$, and then 4) phase by an amount proportional to $R_\text{out}$ by applying a $Z^{\theta 2^r}$ operation with appropriate scaling factor $\theta$ to each bit position $r$ within $R_\text{out}$.
+
+The problem with this naive method is that it produces garbage.
+The register $R_\text{out}$ ends up entangled with $R_\text{in}$.
+If the garbage isn't cleaned up, this decoheres $R_\text{in}$; defeating the purpose of using a quantum computer.
+The method we use to clean up the garbage is to uncompute $R_\text{out}$ by running $C$ in reverse.
+This destroys the garbage, without undoing the phasing effects.
+See \autoref{fig:phase-by-function}.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=2.0em @C=0.7em {
+      &\gate{\exp \big(i \sum_k f(k)|k\rangle\langle k|\big)}&\qw &&=&&&                      &&&\qw &\gate{\text{input k}}         &\qw               &\gate{\text{input k}}         &\qw &\\
+      &                                                      &    && &&&|0\rangle^{\otimes p} &&&\qw &\gate{0 \rightarrow \left\lfloor \frac{f(k)}{2^{p+1} \pi} \right\rfloor \pmod{2^p}}\qwx &\gate{\text{Grad}^{2^{-p}}}&\gate{\left\lfloor \frac{f(k)}{2^{p+1} \pi} \right\rfloor \pmod{2^p} \rightarrow 0}\qwx &\qw &\\
+    }
+  }
+  \caption{
+	Example use of uncomputation: phasing by any computable function $f$, up to a precision of $\epsilon = 2 \pi / 2^p$ radians, by computing an approximation of $f$ under superposition then applying a phase gradient to the output before uncomputing it.
+    \\
+	The gradient operation is implemented by a column of Z-axis rotations, one for each of the $p$ qubits in the temporary output register (the Z rotation applied to the $b$'th qubit is the gate $Z^{2^{-b}}$).
+  }
+  \label{fig:phase-by-function}
+\end{figure}
+
 \end{document}

### 2017-08-23 paper.tex
Investing A states

@@
 \begin{abstract}
-A ``temporary Toffoli" is a doubly-controlled NOT operation that will later be uncomputed by a second double-controlled NOT.
+A ``temporary Toffoli" is a doubly-controlled-not operation that will later be uncomputed by a second double-controlled-not.
 We construct temporary Toffoli gates using four T gates; half as many as previously reported.
 We demonstrate how to apply our construction (and other optimizations) to common operations and algorithms, and show significant improvement over previously reported T-costs.
-For example we cut the reported T-cost of addition by 50\%, the reported T-cost of controlled-addition by 70\%, and the reported T-cost of Shor's algorithm by ?????\%.
+We reduce the reported T-cost of addition by 50\% and the reported T-cost of Shor's algorithm by ?????\%.
 
@@
 One of the downsides of the surface code is that it has no cheap mechanism to apply non-Clifford operations such as $T$ gates.
-Instead, $T$ gates are performed by distilling and consuming $|A\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle$ states.
+Instead, $T$ gates are performed by distilling and consuming $|A\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle)$ states.
 Consuming an $|A\rangle$ state to perform a T gate is simple (see \autoref{fig:t-equals-a}), but distilling $|A\rangle$ states has significant cost.
@@
 
-In [[[[cite]]]] the $|A\rangle$ state distillation is done by preparing an EPR pair $(P, Q)$, moving $P$ out of the surface code and into a different error correcting code that can perform $T$ gates, and using a $T$ gate within that code to rotate and measure $P$ along an axis correlated with the diagonal $X+Y$ axis of $Q$.
+In [[[[cite]]]] the $|A\rangle$ state distillation is done by preparing an EPR pair $(P, Q)$, translating $P$ out of the surface code and into a different error correcting code that can perform $T$ gates, and using a $T$ gate within that code to rotate and measure $P$ along an axis correlated with the diagonal $X+Y$ axis of $Q$.
 If this process succeeds without an error being detected, it steers \cite{Wiseman2007} $Q$ into the $|A\rangle$ state or else into the complementary state $Z|A\rangle$.
@@
 
-The resulting `foothold' $|A\rangle$ states have a non-negligible amount of noise, but can be used to power T gates for rounds of distillation occurring entirely within the surface code.
-Each round of distillation reduces the amount of noise in the final states, but at significant cost.
-After accounting for distilling the $|A\rangle$ state, the spacetime volume cost of a T gate is one to two orders of magnitude higher than Clifford operations native to the surface code (such as the CNOT gate) [[[[cite]]]].
+The resulting `foothold' $|A\rangle$ states have a non-negligible amount of noise, but can be used to perform approximate T gates powering additional rounds of distillation occurring entirely within the surface code.
+After accounting for the cost of distilling a sufficiently accurate $|A\rangle$ state, the spacetime volume cost of performing a T gate is one to two orders of magnitude higher than performing operations native to the surface code (such as the CNOT gate) [[[[cite]]]].
 
 Because T gates are so expensive for the surface code, and the surface code is a likely component of future quantum computers, it is important to consider and optimize the number of T gates used by quantum circuits.
+
 In this paper, we focus on optimizing the number of T gates needed to perform Toffoli gates that will later be uncomputed by a second Toffoli gate (``temporary Toffoli gates").
@@
 \autoref{sec:temporary-and} explains an ancilla erasing trick that improves the cost of temporary Toffolis further, from 6 to 4.
-This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to significantly improve the T-cost of several basic arithmetic tasks (e.g.
-halving the T-cost of addition).
+This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to improve the T-cost of several basic arithmetic tasks (e.g. halving the T-cost of addition).
 Finally, \autoref{sec:conclusion} concludes and discusses future work.
@@
 When this occurs, the three $T$ gates on the control qubits of the textbook construction can be omitted.
-This introduces phase errors (see \autoref{fig:bad-phase-toffoli}), but the second Toffoli gate will uncompute those errors while uncomputing the state permutation (see \autoref{fig:cancelled-bad-phase-toffoli}) \cite{Barenco1995}.
-This technique reduces the T-cost of paired Toffolis from 7 to 4.
+This introduces phase errors (see \autoref{fig:bad-phase-toffoli}), but the second Toffoli gate can uncompute those errors while uncomputing the state permutation \cite{Barenco1995} (see \autoref{fig:cancelled-bad-phase-toffoli}).
 
@@
   \caption{
-	When two Toffolis form a compute/uncompute pair, they cancel phase errors in underlying constructions.
-	Fixes the problem with the construction in \autoref{fig:bad-phase-toffoli}, achieving a per-Toffoli T-cost of 4 for paired Toffolis \cite{Barenco1995}.
+	When two Toffolis form a compute/uncompute pair, they can cancel each others' phase errors.
+	Fixes the problem with the construction in \autoref{fig:bad-phase-toffoli}, and achieves a per-Toffoli T-cost of 4 for paired Toffolis \cite{Barenco1995}.
   }
@@
 
-In \cite{Jones2013}, Jones shows how to perform an Toffoli with a T-cost of 4 even when not pairs.
-They do so by involving an ancilla qubit and an erasure technique that turns what would have been a quantum control into a classical control.
+The paired-phase-correction trick improves the T-cost of temporary Toffolis from 14 (seven to compute, seven to uncompute) to 8 (four to compute, four to uncompute).
+
+In \cite{Jones2013}, Jones shows how to perform an {\em unpaired} Toffoli with four T gates.
+They use an ancilla qubit and an erasure technique that turns what would have been a quantum control into a classical control.
+They also re-arrange the remaining T gates into a single column, creating a circuit with T-depth of 1.
 We show this construction in \autoref{fig:jones-toffoli}.
-Jones also shows how to rearrange all of the T gates into the same column, so that the Toffoli construction has a T-depth of 1 (see figure?).
 
@@
   \caption{
-	Toffoli construction with 4 T gates based on measurement teleportation.
+	Toffoli construction with T-cost of 4 from \cite{Jones2013}.
   }
@@
 
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &\ctrl{1} &\qw & &       &&&     &\ctrl{1} &\ctrl{1} &\qw  &&&   & & & \lstick{x} &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\rstick{x} \\
-      &\ctrl{1} &\qw & & \cong &&&     &\ctrl{2} &\qS      &\qw  &&& = & & & \lstick{y} &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\rstick{y} \\
-      &\targ    &\qw & &       &&&     &\targ    &\qw      &\qw  &&&   & & & \lstick{t} &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\rstick{t \oplus (x \land y)} \\
-      &         &    & &       &&& \qO &\targ    &\qw      &\qw  &&&   & & & \qA        &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qw &\rstick{x \land y} \\
-    }
-  }
-  \caption{
-	Our temporary Toffoli construction, which uses an $|A\rangle$ ancilla and 3 T gates.
-	The $|A\rangle$ state is recovered when uncomputing the temporary Toffoli.
-  }
-  \label{fig:ancilla-temporary-toffoli}
-\end{figure}
+The seed for this paper was an idea that accidentally improved the T-cost of a Toffoli from 7 to 6 in a surprising way.
+Instead of performing a Toffoli directly onto the target qubit, we performed the Toffoli indirectly via an ancilla.
+We started with a clean ancilla in the $|0\rangle$ state, applied a Toffoli to store the logical-and of the two controls in the ancilla, used the ancilla to control a CNOT onto the intended target, then uncomputed the ancilla.
+
+Initially, this indirect-Toffoli construction seems pointless.
+It has a T-cost of 8 even when using the phase-neglected construction, which is worse than the naive T-cost of 7.
+However, as shown in \autoref{fig:indirect-toffoli}, it turns out that the last T gate in the circuit is being used to uncompute an $|A\rangle$ state on the ancilla.
+Dropping this T gate not only reduces the T-cost from 8 to 7, it recovers an $|A\rangle$ state that can be used to power some other T gate.
+This improves the net T-cost to 6.
 
@@
     \Qcircuit @R=1.5em @C=0.7em {
-      &x  &&\ctrl{1} &\qw &x && && \qw  &\qw &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw &           \\
-      &y  &&\ctrl{1} &\qw &y &&=&& \qw  &\qw &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw &           \\
-      &xy &&\qw      &    &  && && \qSi &\qH &\targ    &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qw &|A \rangle \\
+      &\ctrl{1} &\qw && &&&    &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw &&\\
+      &\ctrl{1} &\qw &&=&&&    &\qw &\qw &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\qw &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw &&\\
+      &\targ    &\qw && &&&    &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\qw &&\\
+      &         &    && &&&\qO &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qw 
+          \gategroup{4}{6}{4}{10}{.7em}{--} \gategroup{4}{26}{4}{30}{.7em}{--} &|0\rangle & \\
+      &         &    && &&&    &    |A\rangle &  & &     &         &    &         &     &    &          &    &    &         &    &         &     &         &     & |A\rangle
     }
@@
   \caption{
-	A non-optimal unmerging circuit that uses three $T$ gates, but recovers an $|A\rangle$ state.
-	Avoids measurement, but has a net T-cost of two (which is worse than \autoref{fig:unmerge-free}'s T-cost of zero).
+	Performing a Toffoli indirectly, by applying a pair of Toffolis to an ancilla and using the intermediate value to control a CNOT onto the actual target, appears to have a T-cost of 8.
+	However, the initial T gate is being used to compute an $|A\rangle$ state and the final T gate is being used to uncompute the $|A\rangle$ state.
+    \\	
+	Instead of spending $|A\rangle$ states to perform T gates to make and unmake $|A\rangle$ states, we pass in and later recover a single $|A\rangle$ state.
+	This ``investment" reduces the net T-cost to 6.
   }
-  \label{fig:unmerge-recover-a}
+  \label{fig:indirect-toffoli}
+\end{figure}
+
+A Toffoli with T-cost of 6 isn't optimal, but the reduction from the textbook T-cost of 7 was done in an unexpected way.
+There were likely other situations where this optimization would improve on the state of the art, and we soon found one: temporary Toffolis.
+When a Toffoli is going to be uncomputed, it is not necessary to compute, uncompute, recompute, and reuncompute the ancilla qubit.
+Instead, compute the ancilla for the first Toffoli, keep the ancilla around to make performing the second Toffoli easy, then uncompute the ancilla.
+Separating the control-merging step from the control-unmerging step improves the T-cost of temporary Toffolis from 8 to 6.
+
+[[[[Keeping ancilla around is counter-intuitive, because the ancilla represents entangled garbage that can prevent interference from occurring correctly.
+It just so happens that it makes sense to do so for uncomputation tasks related to classical functions.]]]]
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw && &&&    &\qw     &\qw  &\ctrl{3}&\qw &\qw     &\qw  &\qw &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\qw &\qw &\qw     &\qw  &\ctrl{3}&\qw &\qw     &\qw &&\\
+      &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw &&=&&&    &\ctrl{2}&\qw  &\qw     &\qw &\ctrl{2}&\qw  &\qw &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\qw &\qw &\ctrl{2}&\qw  &\qw     &\qw &\ctrl{2}&\qw &&\\
+      &\targ   &\qw &    &\ldots &    &    &\targ   &\qw && &&&    &\qw     &\qw  &\qw     &\qw &\qw     &\qw  &\qw &\targ    &\qw &    &\ldots &    &    &\targ    &\qw &\qw &\qw     &\qw  &\qw     &\qw &\qw     &\qw &&\\
+      &        &    &    &       &    &    &        &    && &&&\qA &\targ   &\qTi &\targ   &\qT &\targ   &\qTi &\qH &\ctrl{-1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{-1}&\qH &\qT &\targ   &\qTi &\targ   &\qT &\targ   &\qw &|A\rangle &\\
+    }
+  }
+  \caption{
+	A temporary Toffoli construction with a net T-cost of 6.
+	Invests an $|A\rangle$ state when computing the Toffoli, holds an ancilla qubit storing the logical-and of the controls until it's time to uncompute the Toffoli, then recovers the $|A\rangle$ state.
+  }
+  \label{fig:ancilla-temporary-toffoli}
 \end{figure}
@@
     \Qcircuit @R=1.5em @C=0.7em {
-      &x &&\ctrl{1} &\qw & x  && &&          &&\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw &\qw &\\
-      &y &&\ctrl{1} &\qw & y  &&=&&          &&\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\qw &\qw &\\
-      &  &&         &\qw & xy && &&|A\rangle &&\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH &\qS &\qw &\\
+      &x &&\ctrl{1} &\qw & x  && &&          &&\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\qw &\qw &\qw &\\
+      &y &&\ctrl{1} &\qw & y  &&=&&          &&\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw &\qw &\qw &\\
+      &  &&         &\qw & xy && &&|A\rangle &&\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qS &\qw &\\
     }

### 2017-08-23 paper.tex
Fleshing out the logical-and section

@@
 
-\section{AND-ing and Erasing}
+\section{Computing and Erasing the Logical-AND}
 \label{sec:temporary-and}
 
-When the target of a Toffoli gate is a qubit in the $|0\rangle$ state, and the controls are in a computation basis state, the target qubit ends up storing the logical-and of the two controls.
-We continue to say that the target stores the logical-and of the two controls even when they are under superposition, with the understanding that we mean ``within each computation basis state making up the superposition of the overall system".
+A Toffoli gate whose target is in the $|0\rangle$ state is the quantum equivalent of the logical-AND operation.
+For each computational basis state, with the controls storing the bits $x$ and $y$, the target ends up storing $x \land y$.
 
-In the textbook construction of the Toffoli gate, the three T gates on the control wires are implementing a controlled-S.
-A controlled-S phases states by $i$ where both the control and the target are ON.
-If the target of the Toffoli gate starts in the $|0\rangle$ state, then the target ends up ON when both controls are satisfied.
-Therefore the target is ON in exactly the state where a phase factor of $i$ should be applied, and the controlled-S on the controls can be replaced by an unconditional S gate on the target after the Toffoli is finished.
+In the previous section, we showed how to compute and uncompute the logical-AND of two controls by investing an $|A\rangle$ state and performing three $T$ gates.
+This construction be improved in several ways.
 
-A minor optimization which we apply is to note that the first two operations on the target are transitioning from the $|0\rangle$ state to the $|A\rangle$ state.
-Instead of consuming an $|A\rangle$ state to perform the $T$ gate (and potentially performing a fixup S gate), we simply feed the $|A\rangle$ state we would have consumed directly into the circuit.
+First, the computation is introducing phase errors (as shown in \autoref{fig:bad-phase-toffoli}).
+Normally, correcting these phase errors would require controlled-Z on the target (applied before the Toffoli) and a controlled-S between the two controls.
+We know our target is $|0\rangle$ beforehand, so the controlled-Z has no effect and can be ignored.
+The effect of the controlled-S gate is to apply a phase factor of $i$ to the amplitudes of computational basis states where both controls on.
+In other words, the controlled-S phases the logical-and of the two controls and we just happen to have a qubit storing that value.
+We can correct computation's phase error by applying an uncontrolled S gate to the ancilla storing the logical-and.
 
-This construction defines our logical AND operation, which we draw as a wire emerging from two controls as shown in the left hand side of \autoref{fig:logical-and} (which also shows our construction for the operation).
+Second, the construction is unnecessarily deep.
+Every $T$ gate is associated with an S gate that will happen based on the result of a measurement.
+It takes time for the measurement result to become available, and it is beneficial to be waiting for several results in parallel instead of one after another.
+We can reduce our measurement depth to 1 using the same technique as in \cite{Jones2013} (except we only need to perform three T gates instead of four, thanks to having the $|A\rangle$ state as an input).
 
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &x &&\ctrl{1} &\qw & x  && &&          &&\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\qw &\qw &\qw &\\
-      &y &&\ctrl{1} &\qw & y  &&=&&          &&\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw &\qw &\qw &\\
-      &  &&         &\qw & xy && &&|A\rangle &&\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qS &\qw &\\
-    }
-  }
-  \caption{
-	Control merging with T-cost of 4 (three T gates, one $|A\rangle$ state).
-	The $|A\rangle$ state is fed directly into the circuit, instead of using an $H$ and $T$ gate on a $|0\rangle$ state (which would consume an $|A\rangle$ state anyways).
-	The input $|A\rangle$ state can be recovered by using the un-merging construction in \autoref{fig:unmerge-recover-a}, reducing the net T cost of this circuit to 3.
-	However, the un-merging construction in \autoref{fig:unmerge-free} is a better choice, despite not recovering the $|A\rangle$ state, because the total T-cost of merge+unmerge using \autoref{fig:unmerge-free} is 4+0=4 instead of 4+3-1=6.
-  }
-  \label{fig:logical-and}
-\end{figure}
-
-The logical AND construction consumes four $|A\rangle$ states, so its T-cost is 4.
-
+Third, taking another hint from \cite{Jones2013}, we can perform the uncomputation by using an erasure technique.
 To erase the ancilla produced by the logical AND, we use a technique we refer to as ``erasure".
@@
 
-We represent erasing a logical AND with a wire terminating into two controls, as shown in \autoref{fig:erase-and} which also shows the underlying construction.
-The T-cost of erasure is 0.
+With these three improvements applied, we have the three key ingredients of our temporary Toffoli with a T-cost of 4: computing the logical-and, using the logical-and, and uncomputing the logical-and.
+
+We draw the computation of the logical and as a wire emerging vertically from its two inputs (drawn as controls) and heading rightward, as shown in \autoref{fig:compute-logical-and}.
+We draw the uncomputation analogously, with a wire storing the logical-and coming in from the left then merging vertically into its two inputs (drawn as controls), as shown in \autoref{fig:erase-logical-and}.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &x &&\ctrl{1} &\qw & x  && &&          &&\targ    &\ctrl{2}&\qw     &\qTi &\qw     &\ctrl{2}&\targ     &\qw &\qw &\qw &\\
+      &y &&\ctrl{1} &\qw & y  &&=&&          &&\targ    &\qw     &\ctrl{1}&\qTi &\ctrl{1}&\qw     &\targ     &\qw &\qw &\qw &\\
+      &  &&         &\qw & xy && &&|A\rangle &&\ctrl{-2}&\targ   &\targ   &\qT  &\targ   &\targ   &\ctrl{-2} &\qH &\qS &\qw &\\
+    }
+  }
+  \caption{
+	Computing the logical-and of two qubits, with a T-cost of 4 and a measure-react depth of 1.
+  }
+  \label{fig:compute-logical-and}
+\end{figure}
 
@@
   \caption{
-	Control un-merging with T-cost of 0.
+	Uncomputing the logical-and of two qubits, with a T-cost of 0 and a measure-react depth of 1.
   }
-  \label{fig:erase-and}
+  \label{fig:erase-logical-and}
 \end{figure}
+
+During the time that the logical-and ancilla wire exists, it can be used to control operations onto other qubits.
+In effect, these operations are doubly-controlled by the inputs that produced the ancilla.
+For example, we can perform a Toffoli by computing the logical-and of the two controls, applying a CNOT from the logical-and to the actual target, then uncomputing the logical-and.
+This construction, shown in \autoref{fig:merge-use-erase-toffoli}, is equivalent to the one in \cite{Jones2013} (but broken into pieces that can be used independently).
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\targ    &\ctrl{2}&\qw     &\qTi &\qw     &\ctrl{2}&\targ     &\qw &\qw &\qw      &\qw &\qw    &\ctrl{1}         &\qw \\
+      &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\targ    &\qw     &\ctrl{1}&\qTi &\ctrl{1}&\qw     &\targ     &\qw &\qw &\qw      &\qw &\qw    &\gate{Z}         &\qw \\
+      &         &    &&=&&         &\ctrl{1} &\qw      &    &&=&&|A\rangle &&\ctrl{-2}&\targ   &\targ   &\qT  &\targ   &\targ   &\ctrl{-2} &\qH &\qS &\ctrl{1} &\qH &\meter &\cw \cwx \bullet &    \\
+      &\targ    &\qw && &&\qw      &\targ    &\qw      &\qw && &&          &&\qw      &\qw     &\qw     &\qw  &\qw     &\qw     &\qw       &\qw &\qw &\targ    &\qw &\qw    &\qw              &\qw \\
+    }
+  }
+  \caption{
+	Performing a Toffoli gate by computing the logical-and of its controls, using the result, then erasing it.
+	Has a T-cost of 4 and a measure-react depth of 2.
+	Equivalent to the construction from \cite{Jones2013} shown in \autoref{fig:jones-toffoli}.
+  }
+  \label{fig:merge-use-erase-toffoli}
+\end{figure}
+
+The logical AND construction consumes four $|A\rangle$ states, so its T-cost is 4.
+
+We represent erasing a logical AND with a wire terminating into two controls, as shown in \autoref{fig:erase-and} which also shows the underlying construction.
+The T-cost of erasure is 0.
 
@@
 See \autoref{fig:merge-use-erase-toffoli}.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw  &\qw      &\qw &\qw    &\ctrl{1}         &\qw \\
-      &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw      &\qw &\qw  &\qw      &\qw &\qw    &\gate{Z}         &\qw \\
-      &         &    &&=&&         &\ctrl{1} &\qw      &    &&=&&|0\rangle &&\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\targ    &\qH &\qSi &\ctrl{1} &\qH &\meter &\cw \cwx \bullet &    \\
-      &\targ    &\qw && &&\qw      &\targ    &\qw      &\qw && &&          &&\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw      &\qw &\qw  &\targ    &\qw &\qw    &\qw              &\qw \\
-    }
-  }
-  \caption{
-	Performing a Toffoli gate by computing, using, and erasing a logical-AND.
-	Has a T-cost of 4, and is exactly equivalent to the construction from \cite{Jones2013}.
-  }
-  \label{fig:merge-use-erase-toffoli}
-\end{figure}
 

### 2017-08-23 paper.tex
Big adder diagrams

@@
 
+In this section we will show how to apply temporary Toffolis to improve the complexity of several circuits.
+
+First, we note that existing adder constructions contain temporary Toffolis.
+The T-cost of adding two $n$-bit numbers was thought to be $8n + O(1)$.
+We replace the basic adder unit with the building block shown in \ref{fig:full-adder-block}, putting it together as shown in \ref{fig:multi-bit-adder-example} in order to create an adder with a T-cost of $4n + O(1)$.
+
 \begin{figure}
@@
     \Qcircuit @R=1.5em @C=0.7em {
-      &x_k &&\targ    &\ctrl{1}&\qw      &\qw &\qw &\qw &\qw     &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw     &\qw &\qw &\qw      &\ctrl{1}&\targ    &\ctrl{1}&\qw &x_k     &&&&& \\
-      &y_k &&\targ    &\ctrl{1}&\qw      &\qw &\qw &\qw &\qw     &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw     &\qw &\qw &\qw      &\ctrl{1}&\qw      &\targ   &\qw &&&(x+y)_k &&& \\
-      &    &&         &        &\ctrl{1} &\qw &\qw &\qw &\qw     &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw     &\qw &\qw &\ctrl{1} &\qw     &         &        &    &&&        &&& \\
-      &c_k &&\ctrl{-3}&\qw     &\targ    &\qw &\qw &    &c_{k+1} &    &    &    &\ldots &    &    &    &c_{k+1} &    &    &\targ    &\qw     &\ctrl{-3}&\qw     &\qw &c_k     &&&&& \\
+      &c_k &&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\ctrl{1}&\qw     &\qw &c_k     &&&&& \\
+      &i_k &&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\targ   &\ctrl{1}&\qw &i_k     &&&&& \\
+      &t_k &&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\qw     &\targ   &\qw &&&(t+i)_k &&& \\
+      &    &&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &    &&&        &&& \\
     }
@@
   \caption{
-	Inline full adder block with a T-cost of 4.
+	Adder building block with a T-cost of 4.
   }
-  \label{fig:addition}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \makebox[\linewidth]{
-  }
-  \caption{
-	$n$-bit inline addition with T-cost of $8n + O(1)$ based on [[[cite]]].
-  }
-  \label{fig:inline-addition}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \makebox[\linewidth]{
-  }
-  \caption{
-	$n$-bit addition with T-cost of $6n + O(1)$.
-  }
-  \label{fig:ancilla-addition}
+  \label{fig:full-adder-block}
 \end{figure}
@@
     \Qcircuit @R=1.5em @C=0.7em {
-      &    &\ctrl{2}          &\qw               &\qw &&     &&&     &\qS      &\qS      &\gate{Z}     &\ctrl{2}          &\qw               &\qw &&& &&& \lstick{x} &\ctrl{2} &\qw      &\qw &\ctrl{3} &\qw  &\qw      &\qw &\qw      &\qw  &\qw      &\qw &\rstick{x}         \\
-      &    &\qw               &\ctrl{1}          &\qw &&\cong&&&     &\ctrl{-1}&\qS \qwx &\gate{Z}\qwx &\qw               &\ctrl{1}          &\qw &&&=&&& \lstick{y} &\qw      &\qw      &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\ctrl{1} &\qw &\rstick{y}         \\
-      &    &\multigate{1}{+1} &\multigate{1}{+1} &\qw &&     &&&     &\qw      &\ctrl{-1}&\qw          &\multigate{1}{+1} &\multigate{1}{+1} &\qw &&& &&& \lstick{z} &\targ    &\ctrl{1} &\qw &\qw      &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\targ    &\qw &\rstick{(x+y+z)_0} \\
-      &\qO &\ghost{+1}        &\ghost{+1}        &\qw &&     &&& \qO &\qw      &\qw      &\ctrl{-2}    &\ghost{+1}        &\ghost{+1}        &\qw &&& &&& \qA        &\qw      &\targ    &\qT &\targ    &\qSi &\targ    &\qT &\targ    &\qTi &\qH      &\qw &\rstick{(x+y+z)_1} \\
+      &i_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\qw &i_0     &&&&&\\
+      &t_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw &&&(t+i)_0 &&&\\
+      &    &&         &\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\ctrl{1}&\qw     &        &    &&&        &&&\\
+      &i_1 &&\qw      &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\ctrl{1}&\qw &i_1     &&&&&\\
+      &t_1 &&\qw      &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\targ   &\qw &&&(t+i)_1 &&&\\
+      &    &&         &        &        &\targ   &\qw &\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\ctrl{1}&\targ   &\qw     &        &        &        &    &&&        &&&\\
+      &i_2 &&\qw      &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_2     &&&&&\\
+      &t_2 &&\qw      &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_2 &&&\\
+      &    &&         &        &        &        &    &        &        &\targ   &\qw &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\ctrl{3}&\qw     &\ctrl{1}&\targ   &\qw     &        &        &        &        &        &        &    &&&        &&&\\
+      &i_3 &&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_3     &&&&&\\
+      &t_3 &&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_3 &&&\\
+      &    &&         &        &        &        &    &        &        &        &    &        &        &\targ   &\ctrl{2}&\targ   &\qw     &        &        &        &        &        &        &        &        &        &    &&&        &&&\\
+      &i_4 &&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_4     &&&&&\\
+      &t_4 &&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_4 &&&\\
     }
@@
   \caption{
-	Temporary full adder.
+	5-bit adder with T-cost of 16.
+	General $n$-bit adder has T-cost of $4n - 4$.
   }
-  \label{fig:temporary-ancilla-addition}
+  \label{fig:multi-bit-adder-example}
 \end{figure}
 
-Mention binary to unary as an example?
+We also note how to transform our adder into a controlled adder by controlling one of the operations.
+The cost of this controlled adder is $8n + O(1)$; a significant improvement over a recently proposed construction with a T-cost of $21n + O(1)$.
 
-Show example doing an $n$-control NOT with $6n + O(1)$ or $3n + O(1)$ if temporary.
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\text{control} &&&&\qw     &\qw     &\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\qw     &\ctrl{2}&\qw     &\qw &\\
+      &&c_k            &&&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\qw     &\ctrl{2}&\qw &\\
+      &&i_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\ctrl{1}&\targ   &\qw &\\
+      &&t_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\targ   &\targ   &\qw &\\
+      &               &&&&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &    &\\
+    }
+  }
+  \caption{
+	Controlled full adder block with a T-cost of 8.
+  }
+  \label{fig:controlled-full-adder-block}
+\end{figure}
 
-Mention that this trick works more generally. Anytime you have an ancilla that starts HT and ends Tdag H.
+\begin{figure}
+  \centering
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &\text{control} &&&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\ctrl{13}&\qw     &\qw     &\qw     &\ctrl{10}&\qw     &\qw     &\ctrl{7}&\qw     &\qw     &\qw     &\ctrl{4}&\qw     &\qw     &\ctrl{1}&\qw &&&\text{control}               &&&&&&\\
+      &&i_0            &&&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\qw &i_0                          &&&&&&&&\\
+      &&t_0            &&&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw &&&&&&(t+i \cdot \text{control})_0 &&&\\
+      &&               &&&         &\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\qw     &\ctrl{2}&\qw     &        &    &&&                             &&&&&&\\
+      &&i_1            &&&\qw      &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\targ   &\qw     &\qw     &\qw &i_1                          &&&&&&&&\\
+      &&t_1            &&&\qw      &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\targ   &\qw     &\qw     &\qw &&&&&&(t+i \cdot \text{control})_1 &&&\\
+      &&               &&&         &        &        &\targ   &\qw &\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\ctrl{3}&\qw     &\qw     &\ctrl{2}&\targ   &\qw     &        &        &        &        &    &&&                             &&&&&&\\
+      &&i_2            &&&\qw      &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\ctrl{1}&\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &i_2                          &&&&&&&&\\
+      &&t_2            &&&\qw      &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\ctrl{1}&\targ   &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &&&&&&(t+i \cdot \text{control})_2 &&&\\
+      &&               &&&         &        &        &        &    &        &        &\targ   &\qw &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw      &\qw     &\ctrl{3}&\qw     &\qw      &\ctrl{2}&\qw     &        &        &        &        &        &        &        &        &    &&&                             &&&&&&\\
+      &&i_3            &&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw     &\qw      &\qw     &\qw     &\ctrl{1}&\ctrl{1} &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &i_3                          &&&&&&&&\\
+      &&t_3            &&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw     &\qw      &\qw     &\qw     &\ctrl{1}&\targ    &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &&&&&&(t+i \cdot \text{control})_3 &&&\\
+      &&               &&&         &        &        &        &    &        &        &        &    &        &        &\targ   &\ctrl{1}&\qw      &\ctrl{1}&\targ   &\qw     &         &        &        &        &        &        &        &        &        &        &        &    &&&                             &&&&&&\\
+      &&i_4            &&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\targ   &\ctrl{1} &\targ   &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &i_4                          &&&&&&&&\\
+      &&t_4            &&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\targ    &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &&&&&&(t+i \cdot \text{control})_4 &&&\\
+    }
+  }
+  \caption{
+	5-bit controlled adder with a T-cost of 36.
+	General $n$-bit controlled adder has T-cost of $8n - 4$.
+  }
+  \label{fig:ancilla-addition}
+\end{figure}
+
+Other examples of operations which benefit from this optimization include:
+
+- Integer comparisons.
+
+- Integer multiplication.
+
+- Incrementing and counting.
+
+- Modular exponentiation.
+
+- Expanding a binary register into a unary register.
+
+- Operations with a target qubit indexed by a binary qubit register.
+
+- Phasing a register by a computable function $f$ of its value (i.e. applying the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$).
+
+- Temporary permutations.
+
+- Oracles in Grover's algorithm.
+
+- Aggregate synthesis of rotations.
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+Conclude.
 
@@
 
-Austin.
+Austin [reference finding, idea bouncing, review].
 
@@
 
-As an example, suppose we wish to phase the amplitude of each computational basis state $|k\rangle$ of some register $R_{\text{in}}$ by a number of radians computed by some function $f(k)$.
-We want to apply the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$.
-A naive method to perform this task is to 1) construct a classical reversible circuit $C$ for approximating $f$, 2) decompose the gates used by $C$ into a quantum gate set so that the circuit can be run on a quantum computer, 3) prepare a fresh register $R_\text{out}$ and use the circuit $C$ on it and $R_\text{in}$ to produce a binary fixed-point 2s-complement output stored in $R_{\text{out}}$, and then 4) phase by an amount proportional to $R_\text{out}$ by applying a $Z^{\theta 2^r}$ operation with appropriate scaling factor $\theta$ to each bit position $r$ within $R_\text{out}$.
-
-The problem with this naive method is that it produces garbage.
-The register $R_\text{out}$ ends up entangled with $R_\text{in}$.
-If the garbage isn't cleaned up, this decoheres $R_\text{in}$; defeating the purpose of using a quantum computer.
-The method we use to clean up the garbage is to uncompute $R_\text{out}$ by running $C$ in reverse.
-This destroys the garbage, without undoing the phasing effects.
-See \autoref{fig:phase-by-function}.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=2.0em @C=0.7em {
-      &\gate{\exp \big(i \sum_k f(k)|k\rangle\langle k|\big)}&\qw &&=&&&                      &&&\qw &\gate{\text{input k}}         &\qw               &\gate{\text{input k}}         &\qw &\\
-      &                                                      &    && &&&|0\rangle^{\otimes p} &&&\qw &\gate{0 \rightarrow \left\lfloor \frac{f(k)}{2^{p+1} \pi} \right\rfloor \pmod{2^p}}\qwx &\gate{\text{Grad}^{2^{-p}}}&\gate{\left\lfloor \frac{f(k)}{2^{p+1} \pi} \right\rfloor \pmod{2^p} \rightarrow 0}\qwx &\qw &\\
-    }
-  }
-  \caption{
-	Example use of uncomputation: phasing by any computable function $f$, up to a precision of $\epsilon = 2 \pi / 2^p$ radians, by computing an approximation of $f$ under superposition then applying a phase gradient to the output before uncomputing it.
-    \\
-	The gradient operation is implemented by a column of Z-axis rotations, one for each of the $p$ qubits in the temporary output register (the Z rotation applied to the $b$'th qubit is the gate $Z^{2^{-b}}$).
-  }
-  \label{fig:phase-by-function}
-\end{figure}
-
 \end{document}

### 2017-08-25 paper.tex
Temporary addition
Light pass

@@
 
-\title{T gates, temporary Toffolis, and halving the cost of quantum computation}
+\title{T gates, temporary ANDs, and halving the cost of quantum addition}
 \author{Craig Gidney}
@@
 \begin{abstract}
-A ``temporary Toffoli" is a doubly-controlled-not operation that will later be uncomputed by a second double-controlled-not.
-We construct temporary Toffoli gates using four T gates; half as many as previously reported.
-We demonstrate how to apply our construction (and other optimizations) to common operations and algorithms, and show significant improvement over previously reported T-costs.
-We reduce the reported T-cost of addition by 50\% and the reported T-cost of Shor's algorithm by ?????\%.
+We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Amy2013, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
+We do so via a ``temporary AND gate" construction, which uses four T gates to store the logical-and of two qubits into an ancilla and no T gates to later erase the ancilla.
 
-T gates dominate the cost of surface-code-based quantum computation.
-Temporary Toffolis are a basic building block that applies to many quantum circuits.
-As such, a temporary Toffoli gate that uses fewer T gates represents a widely applicable and significant improvement to the projected costs of quantum computation.
+Temporary AND gates are a useful tool when creating other low T cost circuits.
+They can be applied to other integer arithmetic, modular arithmetic, Shor's algorithm, the quantum Fourier transform, Grover oracles, rotation synthesis, and many other circuits.
+Because T gates dominate the cost of quantum computation based on the surface code, and the temporary AND gate is widely applicable, our constructions represent a significant reduction in projected costs of quantum computation.
+
+In addition to our $n$-bit adder circuit with T-cost of $4n-4$, we construct an $n$-bit controlled adder circuit with T-cost of $8n-4$, a temporary adder circuit that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed at no T-cost, an approximate quantum Fourier transform circuit with T-cost of $8 n \lg b + O(b (\lg n) (\lg b))$ where $b = \lg \frac{1}{\epsilon}$ is the desired bit-precision of the operation, and a circuit to rotate $n$ qubits by any angle $\theta$ with T-cost of $4n + O(b \lg n)$.
 \end{abstract}
@@
 The surface code is a quantum error correcting code that works on a 2D nearest-neighbour array of qubits and achieves a threshold error rate of approximately 1\% [[[[cite so much]]]].
-This makes the surface code a likely component in the architecture of large error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies [[[[cite so much]]]] and other known error correcting codes have lower thresholds or require higher connectivity [[[[cite so much]]]].
+This makes the surface code a likely component in the architecture of future error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies [[[[cite so much]]]] and other known error correcting codes have lower thresholds or require stronger connectivity [[[[cite so much]]]].
 
@@
 
-In this paper, we focus on optimizing the number of T gates needed to perform Toffoli gates that will later be uncomputed by a second Toffoli gate (``temporary Toffoli gates").
-Temporary Toffolis show up in many circuits, especially ones that involve computing and uncomputing classical functions, so these optimizations are widely applicable.
+In this paper, we improve the number of T gates needed to perform Toffoli gates that will later be uncomputed by a second Toffoli gate.
+Our construction involves changing the Toffoli gates' target to a clean ancilla qubit.
+This makes the function our construction similar to classical AND gates, which produce a new output instead of toggling an existing value.
+As such, we refer to our construction as a "temporary AND gate" ( as opposed to e.g. ``temporary Toffoli gate" or ``matched Toffoli gate pair").
+
+Toffolis in compute/uncompute pairs appear in many circuits, especially ones that involve computing and uncomputing classical functions, so our temporary AND gate construction is widely applicable.
 
@@
 
-It is not common for adjacent Toffolis to have the same controls, but it is common for a Toffoli to later be uncomputed by a second matching Toffoli (i.e. for the Toffoli's effect to be temporary).
+It isn't common for adjacent Toffolis to have the same controls, but it is common for a Toffoli to later be uncomputed by a second matching Toffoli (i.e. for the Toffoli's effect to be temporary).
 When this occurs, the three $T$ gates on the control qubits of the textbook construction can be omitted.
@@
 
-The paired-phase-correction trick improves the T-cost of temporary Toffolis from 14 (seven to compute, seven to uncompute) to 8 (four to compute, four to uncompute).
-
 In \cite{Jones2013}, Jones shows how to perform an {\em unpaired} Toffoli with four T gates.
@@
 
-By using Jones' construction twice (or the phase-correct-by-pairing construction), a temporary Toffoli can be computed with eight T gates.
+By using Jones' construction twice (or the phase-correct-by-pairing construction), a temporary Toffoli can be computed with eight T gates \cite{Amy2013}.
 Four T gates for the initial Toffoli, and four T gates to uncompute its effect later with another Toffoli.
@@
 
-Initially, this indirect-Toffoli construction seems pointless.
-It has a T-cost of 8 even when using the phase-neglected construction, which is worse than the naive T-cost of 7.
-However, as shown in \autoref{fig:indirect-toffoli}, it turns out that the last T gate in the circuit is being used to uncompute an $|A\rangle$ state on the ancilla.
-Dropping this T gate not only reduces the T-cost from 8 to 7, it recovers an $|A\rangle$ state that can be used to power some other T gate.
+Initially, this indirect-Toffoli construction appears to have a T-cost of 8.
+However, as shown in \autoref{fig:indirect-toffoli}, the last T gate in the circuit is being used to uncompute an $|A\rangle$ state on the ancilla.
+Dropping this T gate not only reduces the T-cost from 8 to 7, it recovers an $|A\rangle$ state that can be consumed to perform a T gate elsewhere.
 This improves the net T-cost to 6.
@@
 	However, the initial T gate is being used to compute an $|A\rangle$ state and the final T gate is being used to uncompute the $|A\rangle$ state.
-    \\	
     Instead of spending $|A\rangle$ states to perform T gates to make and unmake $|A\rangle$ states, we pass in and later recover a single $|A\rangle$ state.
@@
 
-A Toffoli with T-cost of 6 isn't optimal, but the reduction from the textbook T-cost of 7 was done in an unexpected way.
-There were likely other situations where this optimization would improve on the state of the art, and we soon found one: temporary Toffolis.
-When a Toffoli is going to be uncomputed, it is not necessary to compute, uncompute, recompute, and reuncompute the ancilla qubit.
-Instead, compute the ancilla for the first Toffoli, keep the ancilla around to make performing the second Toffoli easy, then uncompute the ancilla.
-Separating the control-merging step from the control-unmerging step improves the T-cost of temporary Toffolis from 8 to 6.
-
-[[[[Keeping ancilla around is counter-intuitive, because the ancilla represents entangled garbage that can prevent interference from occurring correctly.
-It just so happens that it makes sense to do so for uncomputation tasks related to classical functions.]]]]
+A Toffoli with T-cost of 6 isn't optimal, but the reduction from 7 was done in an unexpected way.
+While searching for other other circuits where this optimization might apply, we realized it would be useful for the paired-Toffolis case.
+Instead of computing and uncomputing the ancilla for the first Toffoli, then recomputing and reuncomputing it for the second Toffoli, we simply skipped the first ancilla uncomputation and kept the ancilla until it was needed again.
+This separated the construction into a control-combining step and an entanglement-erasing step.
+Combining the controls had a net T-cost of 4, because it performed three $T$ gates and required investing an $|A\rangle$ state, but erasing the entanglement had a net T-cost of only 2 because although it performed three $T$ gates it recovered the invested $|A\rangle$ state.
+See \autoref{fig:ancilla-temporary-toffoli}.
 
@@
   \caption{
-	A temporary Toffoli construction with a net T-cost of 6.
+	Computing and uncomputing a Toffoli gate with a net T-cost of 6.
 	Invests an $|A\rangle$ state when computing the Toffoli, holds an ancilla qubit storing the logical-and of the controls until it's time to uncompute the Toffoli, then recovers the $|A\rangle$ state.
@@
 
-\section{Computing and Erasing the Logical-AND}
+Many circuits involve computing and later uncomputing a Toffoli (e.g. addition), and investing an $|A\rangle$ state reduced the T-cost of doing so from 8 to 6.
+We realized this represented an improvement over the state of the art but, while reviewing the literature, we found a better method for erasing the entanglement.
+We mention investing $|A\rangle$ states here only because we expect that there are circuits that will benefit from it but not our improved technique.
+
+\section{The temporary AND gate}
 \label{sec:temporary-and}
-
-A Toffoli gate whose target is in the $|0\rangle$ state is the quantum equivalent of the logical-AND operation.
-For each computational basis state, with the controls storing the bits $x$ and $y$, the target ends up storing $x \land y$.
 
@@
 The effect of the controlled-S gate is to apply a phase factor of $i$ to the amplitudes of computational basis states where both controls on.
-In other words, the controlled-S phases the logical-and of the two controls and we just happen to have a qubit storing that value.
+In other words, the controlled-S phases the logical-and of the two controls.
+We happen to have a qubit storing that value.
 We can correct computation's phase error by applying an uncontrolled S gate to the ancilla storing the logical-and.
@@
 Second, the construction is unnecessarily deep.
-Every $T$ gate is associated with an S gate that will happen based on the result of a measurement.
+Every T gate is associated with an S gate that will happen conditionally, based on the result of a measurement.
 It takes time for the measurement result to become available, and it is beneficial to be waiting for several results in parallel instead of one after another.
-We can reduce our measurement depth to 1 using the same technique as in \cite{Jones2013} (except we only need to perform three T gates instead of four, thanks to having the $|A\rangle$ state as an input).
+We can reduce our measurement depth to 1 using the same technique as in \cite{Jones2013}.
+We will continue to pass an $|A\rangle$ state into the circuit, instead of using that $|A\rangle$ state to perform a T gate, because doing so has no depth cost.
 
-Third, taking another hint from \cite{Jones2013}, we can perform the uncomputation by using an erasure technique.
+Third, taking another hint from \cite{Jones2013}, we can perform the uncomputation by using a measure-and-correct process.
 To erase the ancilla produced by the logical AND, we use a technique we refer to as ``erasure".
@@
 
-With these three improvements applied, we have the three key ingredients of our temporary Toffoli with a T-cost of 4: computing the logical-and, using the logical-and, and uncomputing the logical-and.
+With these three improvements applied, we have the three key ingredients of our temporary AND gate with a T-cost of 4: creating, using, and erasing.
 
-We draw the computation of the logical and as a wire emerging vertically from its two inputs (drawn as controls) and heading rightward, as shown in \autoref{fig:compute-logical-and}.
-We draw the uncomputation analogously, with a wire storing the logical-and coming in from the left then merging vertically into its two inputs (drawn as controls), as shown in \autoref{fig:erase-logical-and}.
+In diagrams, we draw the AND gate as a wire emerging vertically from two controls then heading rightward, as shown in \autoref{fig:compute-logical-and}.
+We draw the uncomputation analogously, with a wire storing the logical-and coming in from the left then merging vertically into the two controls that created it, as shown in \autoref{fig:erase-logical-and}.
 
@@
   \caption{
-	When distant Toffolis share the same controls, a single logical-AND can Cliffordize both.
-    The T-cost of computing and uncomputing the logical-AND is 4, so in effect this reduces the average T-cost of paired Toffolis to 2.
+	Replacing a pair of compute/uncompute Toffolis with a temporary AND gate and Clifford operations.
+	Improves the T-cost from 8 to 4.
   }
@@
 
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=1.5em @C=0.7em {
+      &c_k &&\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw &&c_k     &&& \\
+      &i_k &&\targ   &\ctrl{1}&\qw     &\ctrl{1}&\qw &&i_k     &&& \\
+      &t_k &&\targ   &\ctrl{1}&\qw     &\targ   &\qw &&&(t+i)_k && \\
+      &    &&        &        &\targ   &\qw     &\qw &&c_{k+1} &&& \\
+    }
+  }
+  \caption{
+	  Temporary addition building block with a T-cost of 4 to compute and a T-cost of 0 to uncompute (not shown).
+  }
+  \label{fig:temporary-full-adder-block}
+\end{figure}
+
 We also note how to transform our adder into a controlled adder by controlling one of the operations.
@@
 
-- Modular exponentiation.
+- Modular arithmetic.
+
+- Operations with a target qubit indexed by a binary qubit register.
 
 - Expanding a binary register into a unary register.
-
-- Operations with a target qubit indexed by a binary qubit register.
 
@@
 
-- Aggregate synthesis of rotations.
+- etc
 
@@
 
-Conclude.
+It was commonly believed that the optimal number of T gates to perform addition was 8, or 7 \cite{AustinDiscussionsAndEmails2017}.
+We showed that the number of T gates is actually 4.
+We are not aware of any method for improving the T cost further, but we do prove the temporary AND gatea lower bound of 2 In \autoref{fig:lower-bound} we show a lower bound of 2
 
@@
 
+>>>>>>>>>>>>>>>>>>>>>>> https://arxiv.org/pdf/1206.0758.pdf          paper that says addition costs 8N T
+
 https://arxiv.org/pdf/1706.05113.pdf   has a ctrl-add with 21N + O(1). This technique+commutators achieves 8N.

### 2017-08-25 paper.tex
Polishing first few seconds - Dropped description of A-state distillation - General wordery

@@
 We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Amy2013, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
-We do so via a ``temporary AND gate" construction, which uses four T gates to store the logical-and of two qubits into an ancilla and no T gates to later erase the ancilla.
+We do so via a ``temporary AND gate" construction, which uses four T gates to store the logical-and of two qubits into an ancilla and zero T gates to later erase the ancilla.
 
-Temporary AND gates are a useful tool when creating other low T cost circuits.
-They can be applied to other integer arithmetic, modular arithmetic, Shor's algorithm, the quantum Fourier transform, Grover oracles, rotation synthesis, and many other circuits.
+Temporary AND gates are a generally useful tool when optimizing T-cost.
+They can be applied to integer arithmetic, modular arithmetic, rotation synthesis, the quantum Fourier transform, Shor's algorithm, Grover oracles, and many other circuits.
 Because T gates dominate the cost of quantum computation based on the surface code, and the temporary AND gate is widely applicable, our constructions represent a significant reduction in projected costs of quantum computation.
@@
 Instead, $T$ gates are performed by distilling and consuming $|A\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle)$ states.
-Consuming an $|A\rangle$ state to perform a T gate is simple (see \autoref{fig:t-equals-a}), but distilling $|A\rangle$ states has significant cost.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\gate{T}&\qw &&=&&&          &&\qw &\ctrl{1}&\qw    &\qS            &\qw &\\
-      &        &    && &&&|A\rangle &&\qw &\targ   &\meter &\cw\cwx\bullet &    &\\
-    }
-  }
-  \caption{
-	Performing a T gate with Clifford operations by consuming an $|A\rangle$ state, where $|A\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle$.
-  }
-  \label{fig:t-equals-a}
-\end{figure}
-
-In [[[[cite]]]] the $|A\rangle$ state distillation is done by preparing an EPR pair $(P, Q)$, translating $P$ out of the surface code and into a different error correcting code that can perform $T$ gates, and using a $T$ gate within that code to rotate and measure $P$ along an axis correlated with the diagonal $X+Y$ axis of $Q$.
-If this process succeeds without an error being detected, it steers \cite{Wiseman2007} $Q$ into the $|A\rangle$ state or else into the complementary state $Z|A\rangle$.
-The $Z$ rotation is easily corrected, and the result of measuring $P$ indicates whether the correction is needed or not.
-
-The resulting `foothold' $|A\rangle$ states have a non-negligible amount of noise, but can be used to perform approximate T gates powering additional rounds of distillation occurring entirely within the surface code.
-After accounting for the cost of distilling a sufficiently accurate $|A\rangle$ state, the spacetime volume cost of performing a T gate is one to two orders of magnitude higher than performing operations native to the surface code (such as the CNOT gate) [[[[cite]]]].
-
+Consuming an $|A\rangle$ state to perform a T gate is simple, but distilling $|A\rangle$ states has significant cost.
 Because T gates are so expensive for the surface code, and the surface code is a likely component of future quantum computers, it is important to consider and optimize the number of T gates used by quantum circuits.
 
-In this paper, we improve the number of T gates needed to perform Toffoli gates that will later be uncomputed by a second Toffoli gate.
-Our construction involves changing the Toffoli gates' target to a clean ancilla qubit.
-This makes the function our construction similar to classical AND gates, which produce a new output instead of toggling an existing value.
-As such, we refer to our construction as a "temporary AND gate" ( as opposed to e.g. ``temporary Toffoli gate" or ``matched Toffoli gate pair").
-
-Toffolis in compute/uncompute pairs appear in many circuits, especially ones that involve computing and uncomputing classical functions, so our temporary AND gate construction is widely applicable.
+In this paper, we focus on improving the number of T gates needed to perform Toffoli gates that will later be uncomputed by a second Toffoli gate.
+We then apply these improvements.
+The key ideas behind our construction are 1) perform the Toffoli indirectly by targeting a clean ancilla qubit and then using the ancilla to toggle the intended target, 2) don't uncompute and recompute the ancilla if it will be needed again, 3) if a T gate is being used to compute and/or uncompute an $|A\rangle$ state then just pass in and/or recover an $|A\rangle$ state instead, and 4) uncompute the ancilla by measuring it then applying a classically controlled fixup operation.
+We refer to initializing the ancilla as ``computing the logical-AND of the controls", to uncomputing the ancilla as ``erasing the logical-AND", and to the combination of both pieces as a ``temporary AND gate".
 
 Our paper is divided into six sections.
-In \autoref{sec:introduction}, we motivate the problem and insert self-referential descriptions.
-\autoref{sec:review} discusses how previous work on optimizing the T gate counts of Toffolis improved the cost of temporary Toffolis from 14 to 8.
-In \autoref{sec:invest}, we show how to improve the cost of temporary Toffolis from 8 to 6 by investing and later recovering an $|A\rangle$ state instead of consuming it to perform T gates.
-\autoref{sec:temporary-and} explains an ancilla erasing trick that improves the cost of temporary Toffolis further, from 6 to 4.
-This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to improve the T-cost of several basic arithmetic tasks (e.g. halving the T-cost of addition).
-Finally, \autoref{sec:conclusion} concludes and discusses future work.
+In \autoref{sec:introduction}, we motivate the problem and introduce terminology.
+\autoref{sec:review} discusses how previous work on optimizing T gate counts improved the cost of matched pairs of Toffolis from 14 to 8.
+In \autoref{sec:invest}, we show how to improve the T-cost from 8 to 6 by investing and later recovering an $|A\rangle$ state instead of consuming it to perform T gates.
+\autoref{sec:temporary-and} expands on those ideas with an ancilla erasing trick that improves the T-cost from 6 to 4.
+This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to create low T-cost circuits for several basic arithmetic tasks (including cutting the T-cost of addition in half, with respect to previously reported circuits).
+Finally, \autoref{sec:conclusion} summarizes our contributions, proves that at least two T gates are needed to perform a temporary AND, and discusses other possible applications of our constructions.
 
@@
 Assuming we aren't permitted to involve other qubits or to share work with other operations, this construction is optimal \cite{Gosset2014}.
-Of course, in practice, we can do both.
+
+When optimizations can cross qubit and gate boundaries, the textbook construction isn't optimal.
 For example, when several adjacent Toffoli gates share the same controls, all but one can be replaced by CNOT operations (see \autoref{fig:shared-controls}).
@@
 
-In \cite{Jones2013}, Jones shows how to perform an {\em unpaired} Toffoli with four T gates.
-They use an ancilla qubit and an erasure technique that turns what would have been a quantum control into a classical control.
-They also re-arrange the remaining T gates into a single column, creating a circuit with T-depth of 1.
-We show this construction in \autoref{fig:jones-toffoli}.
+A typical $n$-bit quantum adder will contain $2n + O(1)$ Toffoli gates, implying a naive T-cost of $14n + O(1)$.
+However, almost all of the Toffoli gates in the first half of an adder are uncomputed by Toffoli gates in the second half.
+This allows the T gates on the controls of the Toffoli gates to be omitted, reducing their T-cost from 7 to 4 and the T-cost of addition to $8n + O(1)$.
+
+Even if a Toffoli isn't paired with a second Toffoli that uncomputes its effects, it's still possible to perform the Toffoli with T-cost of by using an ancilla, a measurement, and a conditional fixup operation \cite{Jones2013} (see \autoref{fig:jones-toffoli}).
+Our key improvement over \cite{Jones2013} is to notice that the ancilla's lifetime can be extended, so it can be used both when computing and uncomputing a Toffoli.
 
@@
 
-By using Jones' construction twice (or the phase-correct-by-pairing construction), a temporary Toffoli can be computed with eight T gates \cite{Amy2013}.
-Four T gates for the initial Toffoli, and four T gates to uncompute its effect later with another Toffoli.
-
 \section{Investing $|A\rangle$ states}
@@
 
-The seed for this paper was an idea that accidentally improved the T-cost of a Toffoli from 7 to 6 in a surprising way.
-Instead of performing a Toffoli directly onto the target qubit, we performed the Toffoli indirectly via an ancilla.
+This paper started with an idea that improved the T-cost of a Toffoli from 7 to 6 in a surprising way.
+Instead of performing a Toffoli directly onto the target qubit, we tried performing the Toffoli indirectly.
 We started with a clean ancilla in the $|0\rangle$ state, applied a Toffoli to store the logical-and of the two controls in the ancilla, used the ancilla to control a CNOT onto the intended target, then uncomputed the ancilla.
 
-Initially, this indirect-Toffoli construction appears to have a T-cost of 8.
-However, as shown in \autoref{fig:indirect-toffoli}, the last T gate in the circuit is being used to uncompute an $|A\rangle$ state on the ancilla.
-Dropping this T gate not only reduces the T-cost from 8 to 7, it recovers an $|A\rangle$ state that can be consumed to perform a T gate elsewhere.
+Initially, this indirect-Toffoli construction appeared to have a T-cost of 8.
+However, as shown in \autoref{fig:indirect-toffoli}, the last T gate in the circuit is unnecessary.
+In fact, it is actively harmful.
+Dropping this T gate and the following Hadamard not only reduces the T-cost from 8 to 7, it leaves the ancilla in an $|A\rangle$ state that can be consumed to perform a T gate elsewhere.
 This improves the net T-cost to 6.
@@
 A Toffoli with T-cost of 6 isn't optimal, but the reduction from 7 was done in an unexpected way.
-While searching for other other circuits where this optimization might apply, we realized it would be useful for the paired-Toffolis case.
-Instead of computing and uncomputing the ancilla for the first Toffoli, then recomputing and reuncomputing it for the second Toffoli, we simply skipped the first ancilla uncomputation and kept the ancilla until it was needed again.
+While searching for other circuits where this optimization might apply, we realized it would be useful in the paired-Toffolis case.
+Instead of computing and uncomputing the ancilla for the first Toffoli, then recomputing and re-uncomputing it for the second Toffoli, we simply kept the ancilla from the first Toffoli so that the second Toffoli could also use it.
 This separated the construction into a control-combining step and an entanglement-erasing step.
-Combining the controls had a net T-cost of 4, because it performed three $T$ gates and required investing an $|A\rangle$ state, but erasing the entanglement had a net T-cost of only 2 because although it performed three $T$ gates it recovered the invested $|A\rangle$ state.
-See \autoref{fig:ancilla-temporary-toffoli}.
+
+Combining the controls had a net T-cost of 4, because it involved performing three $T$ gates and investing one $|A\rangle$ state.
+But erasing the entanglement had a net T-cost of 2 because, although it performed three $T$ gates, it recovered the $|A\rangle$ state invested when combining the controls.
+\autoref{fig:ancilla-temporary-toffoli} shows how to replace a pair of Toffolis, where the second Toffoli is uncomputing the first, with this construction.
 
@@
 Many circuits involve computing and later uncomputing a Toffoli (e.g. addition), and investing an $|A\rangle$ state reduced the T-cost of doing so from 8 to 6.
-We realized this represented an improvement over the state of the art but, while reviewing the literature, we found a better method for erasing the entanglement.
-We mention investing $|A\rangle$ states here only because we expect that there are circuits that will benefit from it but not our improved technique.
+We realized this improved on existing work, and began writing this paper.
+We then found a construction that achieves the same effect with lower T-cost, which we will cover in the next section.
+However, there are likely circuits without temporary logical-ANDs that nevertheless would benefit from investing and recovering $|A\rangle$ states.
+It's an interesting fallback optimization, which we mention for completeness.
 
@@
 
-In the previous section, we showed how to compute and uncompute the logical-AND of two controls by investing an $|A\rangle$ state and performing three $T$ gates.
-This construction be improved in several ways.
+In the previous section, we showed how to compute and uncompute a Toffoli gate by investing an $|A\rangle$ state and performing three $T$ gates.
+This construction be improved in several ways, ultimately producing our temporary AND gate construction.
 
-First, the computation is introducing phase errors (as shown in \autoref{fig:bad-phase-toffoli}).
-Normally, correcting these phase errors would require controlled-Z on the target (applied before the Toffoli) and a controlled-S between the two controls.
+First, the computation step is introducing phase errors (the same as in \autoref{fig:bad-phase-toffoli}).
+Correcting these phase errors requires applying a controlled-Z gate involving the target (before the Toffoli gate), and applying a controlled-S gate to the two controls.
 We know our target is $|0\rangle$ beforehand, so the controlled-Z has no effect and can be ignored.
 The effect of the controlled-S gate is to apply a phase factor of $i$ to the amplitudes of computational basis states where both controls on.
-In other words, the controlled-S phases the logical-and of the two controls.
-We happen to have a qubit storing that value.
-We can correct computation's phase error by applying an uncontrolled S gate to the ancilla storing the logical-and.
+We happen to be computing a qubit storing whether or not both controls are on, so we can replace the controlled-S with an uncontrolled S gate on the ancilla.
+This corrects all of the phase errors that were being introduced.
 
-Second, the construction is unnecessarily deep.
-Every T gate is associated with an S gate that will happen conditionally, based on the result of a measurement.
-It takes time for the measurement result to become available, and it is beneficial to be waiting for several results in parallel instead of one after another.
-We can reduce our measurement depth to 1 using the same technique as in \cite{Jones2013}.
-We will continue to pass an $|A\rangle$ state into the circuit, instead of using that $|A\rangle$ state to perform a T gate, because doing so has no depth cost.
+Second, the construction from the previous section is unnecessarily deep.
+Each T gate is hiding a measurement and an S gate performed conditional on the outcome of that measurement.
+It is beneficial to be waiting for several measurement results in parallel, instead of one after another serially.
+As shown in \cite{Jones2013}, the T gates of a Toffoli can all be performed at the same time, and we do so.
+Note that still pass an $|A\rangle$ state into the circuit instead of performing a fourth T gate, because passing in an $|A\rangle$ state already doesn't involve measuring and reacting.
 
-Third, taking another hint from \cite{Jones2013}, we can perform the uncomputation by using a measure-and-correct process.
-To erase the ancilla produced by the logical AND, we use a technique we refer to as ``erasure".
+Third, taking another hint from \cite{Jones2013}, we can uncompute the ancilla by using a measure-and-correct process instead of with a mirror of the computation circuit.
+To uncompute the ancilla produced by the logical AND, we use a technique we refer to as ``erasure".
 We note that a Toffoli gate would obviously clear the value.
-Then, since the target qubit can be discarded after clearing it, we are free to introduce a Hadamard gate and a measurement before discarding it.
+Then, since the cleared ancilla was going to be discarded anyways, we are free to apply a Hadamard gate and a measurement to it.
 We then hop the Hadamard over the Toffoli, transforming it into a CCZ operation.
-The controls and targets of a CCZ are equivalent, so we rewrite the circuit into a CZC.
-We then hop the Measurement over the CZC, turning a quantum control into a classical control.
-These circuit moves turned the Toffoli gate into a CZ gate that we apply or not based on the outcome of the measreument
-We ``Clifford-ized" the circuit by adding a Hadamard gate and a measurement.
+Next we re-arrange the CCZ so that the ancilla is one of the controls, which is possible because the controls and targets of a CCZ are interchangeable.
+We then hop the Measurement over the CCZ, turning the quantum control into a classical control.
+Ultimately, the Toffoli gate we started with has been turned into a CZ gate that we apply or not based on the outcome of the measurement.
+All of the remaining gates are Clifford operations, which have no T-cost.
 
-With these three improvements applied, we have the three key ingredients of our temporary AND gate with a T-cost of 4: creating, using, and erasing.
+With these three improvements applied, we are left with the key ingredients of our temporary AND gate.
+We can initialize an ancilla storing the logical-AND of two qubits.
+The creation process has a T-cost of 4.
+Then we can keep the ancilla around as long as we want, using it as a control for operations that would otherwise have been conditioned on both input qubits.
+Finally, once the ancilla is no longer needed, we can erase it at no T-cost.
 
-In diagrams, we draw the AND gate as a wire emerging vertically from two controls then heading rightward, as shown in \autoref{fig:compute-logical-and}.
-We draw the uncomputation analogously, with a wire storing the logical-and coming in from the left then merging vertically into the two controls that created it, as shown in \autoref{fig:erase-logical-and}.
+In diagrams, we draw the computation of the AND gate (i.e.
+the ancilla being initialized) as an ancilla wire emerging vertically from two controls then heading rightward (see \autoref{fig:compute-logical-and}).
+We draw the uncomputation analogously, with the ancilla wire coming in from the left then merging vertically into the two controls that created it (seee \autoref{fig:erase-logical-and}).
+The initialized lifetime of the ancilla qubit corresponds to the span of the wire coming out of and later merging into the control qubits.
 
@@
 
-During the time that the logical-and ancilla wire exists, it can be used to control operations onto other qubits.
-In effect, these operations are doubly-controlled by the inputs that produced the ancilla.
-For example, we can perform a Toffoli by computing the logical-and of the two controls, applying a CNOT from the logical-and to the actual target, then uncomputing the logical-and.
-This construction, shown in \autoref{fig:merge-use-erase-toffoli}, is equivalent to the one in \cite{Jones2013} (but broken into pieces that can be used independently).
+We can perform a single Toffoli by computing the logical-and of two qubits , applying a CNOT from the logical-and onto the actual target, then uncomputing the logical-and.
+This combined construction, shown in \autoref{fig:merge-use-erase-toffoli}, is equivalent to the one in \cite{Jones2013}.
 
@@
 
-The logical AND construction consumes four $|A\rangle$ states, so its T-cost is 4.
-
-We represent erasing a logical AND with a wire terminating into two controls, as shown in \autoref{fig:erase-and} which also shows the underlying construction.
-The T-cost of erasure is 0.
-
-The benefit of separating the logical-and-creating operation from the logical-and-erasing operation is that, as long as we have the ancilla storing the logical AND around, we can Cliffordize any Toffoli gate whose controls would be the two qubits we merged by replacing the Toffoli with a CNOT with the ancilla qubit as a control.
-When the parts are simply placed next to each other, we end up with a construction exactly equivalent to the ones presented by Jones \cite{Jones2013}.
-See \autoref{fig:merge-use-erase-toffoli}.
-
-In order to make savings, we need to keep the ancilla around for longer.
-We need a situation like the one shown in \autoref{fig:paired-toffoli-to-logical-and}, where a single logical-AND computation replaces two Toffolis.
+Conceptually, our improvement on the construction in \cite{Jones2013} is merely noticing that it can be split it useful pieces that don't all need to be applied at the same time.
+In particular, when a Toffoli is later uncomputed by a second Toffoli, we simply delay erasing the ancilla wire as shown in \autoref{fig:paired-toffoli-to-logical-and}.
+This reduces the T-cost of such paired Toffolis from 8 to 4.
 
@@
 \end{figure}
-
-Fortunately, this pattern of distant Toffolis sharing controls is very common in quantum circuits.
-It occurs during uncomputation, and also in circuits that sweep back and forth across qubits like addition.
 

### 2017-08-26 paper.tex
First draft

@@
 \begin{abstract}
-We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Amy2013, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
+We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Cuccaro2004, Amy2013, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
 We do so via a ``temporary AND gate" construction, which uses four T gates to store the logical-and of two qubits into an ancilla and zero T gates to later erase the ancilla.
@@
 
-In addition to our $n$-bit adder circuit with T-cost of $4n-4$, we construct an $n$-bit controlled adder circuit with T-cost of $8n-4$, a temporary adder circuit that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed at no T-cost, an approximate quantum Fourier transform circuit with T-cost of $8 n \lg b + O(b (\lg n) (\lg b))$ where $b = \lg \frac{1}{\epsilon}$ is the desired bit-precision of the operation, and a circuit to rotate $n$ qubits by any angle $\theta$ with T-cost of $4n + O(b \lg n)$.
+In addition to our $n$-bit adder circuit with T-cost of $4n-4$, we construct an $n$-bit controlled adder circuit with T-cost of $8n-4$, a temporary adder circuit that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed at no T-cost, a phase gradient circuit with T-cost of $4n + O(b \lg n)$ where $b = \lg \frac{1}{\epsilon}$ is the desired bit-precision of the operation, a quantum Fourier transform circuit with T-cost of $8 n \lg b + O(b (\lg n) (\lg b))$, and a circuit to rotate $n$ qubits by any angle $\theta$ with T-cost of $4n + O(b \lg n)$.
 \end{abstract}
@@
 \autoref{sec:temporary-and} expands on those ideas with an ancilla erasing trick that improves the T-cost from 6 to 4.
-This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to create low T-cost circuits for several basic arithmetic tasks (including cutting the T-cost of addition in half, with respect to previously reported circuits).
-Finally, \autoref{sec:conclusion} summarizes our contributions, proves that at least two T gates are needed to perform a temporary AND, and discusses other possible applications of our constructions.
+This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to create low T-cost circuits for several addition tasks and discuss other possible applications of our constructions.
+Finally, \autoref{sec:conclusion} summarizes our contributions, and proves that at least two T gates are needed to perform a temporary AND.
 
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1}&\qw & &   & & &\ctrl{1}&\qw  &\ctrl{1}&\qT &\qw     &\qw  &\ctrl{2}&\qw &\qw     &\qw  &\ctrl{2}&\qw &\qw \\
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1}&\ctrl{1}&\ctrl{1}&\ctrl{1}&\qw && &&&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw &\\
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1} &\qw & &       & & &\ctrl{1}     &\ctrl{1} &\qw & &   & & &\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\qw &\qw \\
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw && &&&\ctrl{1}     &\ctrl{1}  &\qw &\qw &\qw    &\qw &\qw     &\ctrl{1}&\ctrl{1}     &\qw & \\
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1} &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw  &\qw       &\qw &\qw    &\ctrl{1}       &\qw \\
@@
 
-\section{Investing $|A\rangle$ states}
+\section{Warm up: Investing $|A\rangle$ states}
 \label{sec:invest}
 
-This paper started with an idea that improved the T-cost of a Toffoli from 7 to 6 in a surprising way.
-Instead of performing a Toffoli directly onto the target qubit, we tried performing the Toffoli indirectly.
+This paper started with an idea that improved the T-cost of a single Toffoli gate from 7 to 6 in a surprising way, by performing the Toffoli indirectly instead of directly.
 We started with a clean ancilla in the $|0\rangle$ state, applied a Toffoli to store the logical-and of the two controls in the ancilla, used the ancilla to control a CNOT onto the intended target, then uncomputed the ancilla.
 
-Initially, this indirect-Toffoli construction appeared to have a T-cost of 8.
+Initially, this indirect-Toffoli construction appears to have a T-cost of 8.
 However, as shown in \autoref{fig:indirect-toffoli}, the last T gate in the circuit is unnecessary.
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1} &\qw && &&&    &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw &&\\
@@
 
-A Toffoli with T-cost of 6 isn't optimal, but the reduction from 7 was done in an unexpected way.
-While searching for other circuits where this optimization might apply, we realized it would be useful in the paired-Toffolis case.
-Instead of computing and uncomputing the ancilla for the first Toffoli, then recomputing and re-uncomputing it for the second Toffoli, we simply kept the ancilla from the first Toffoli so that the second Toffoli could also use it.
-This separated the construction into a control-combining step and an entanglement-erasing step.
-
-Combining the controls had a net T-cost of 4, because it involved performing three $T$ gates and investing one $|A\rangle$ state.
-But erasing the entanglement had a net T-cost of 2 because, although it performed three $T$ gates, it recovered the $|A\rangle$ state invested when combining the controls.
-\autoref{fig:ancilla-temporary-toffoli} shows how to replace a pair of Toffolis, where the second Toffoli is uncomputing the first, with this construction.
+A Toffoli with a T-cost of 6 isn't optimal, but the reduction from 7 was done in an unexpected way.
+While searching for other circuits where this optimization might apply, we realized it would be useful when an initial Toffoli was later uncomputed by a second Toffoli.
+Instead of computing and uncomputing the ancilla for the first Toffoli, then recomputing and re-uncomputing the ancilla for the second Toffoli, we simply kept the ancilla around until it was time to uncompute the Toffoli.
+This halved the T-cost of the pair from 12 to 6.
+See \autoref{fig:ancilla-temporary-toffoli}.
 
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw && &&&    &\qw     &\qw  &\ctrl{3}&\qw &\qw     &\qw  &\qw &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\qw &\qw &\qw     &\qw  &\ctrl{3}&\qw &\qw     &\qw &&\\
@@
 
-Many circuits involve computing and later uncomputing a Toffoli (e.g. addition), and investing an $|A\rangle$ state reduced the T-cost of doing so from 8 to 6.
-We realized this improved on existing work, and began writing this paper.
-We then found a construction that achieves the same effect with lower T-cost, which we will cover in the next section.
-However, there are likely circuits without temporary logical-ANDs that nevertheless would benefit from investing and recovering $|A\rangle$ states.
-It's an interesting fallback optimization, which we mention for completeness.
+Many circuits involve computing and later uncomputing a Toffoli (e.g. addition).
+Previously, it was believed that each Toffoli had a T-cost of 4 so the pair would have a T-cost of 8.
+Our $|A\rangle$-investing construction state reduces the T-cost of the pair from 8 to 6.
 
-\section{The temporary AND gate}
+We further improve the cost of paired Toffoli gates in the next section, obsoleting the $|A\rangle$-investing construction we just explained.
+We covered it regardless, because we believe that investing and recovering $|A\rangle$ states is a useful optimization likely to apply more generally to other kinds of circuits.
+
+\section{Constructing a temporary AND}
 \label{sec:temporary-and}
 
-In the previous section, we showed how to compute and uncompute a Toffoli gate by investing an $|A\rangle$ state and performing three $T$ gates.
+In the previous section, we showed how to compute and uncompute a Toffoli gate by investing an $|A\rangle$ state and performing six $T$ gates.
 This construction be improved in several ways, ultimately producing our temporary AND gate construction.
 
-First, the computation step is introducing phase errors (the same as in \autoref{fig:bad-phase-toffoli}).
-Correcting these phase errors requires applying a controlled-Z gate involving the target (before the Toffoli gate), and applying a controlled-S gate to the two controls.
-We know our target is $|0\rangle$ beforehand, so the controlled-Z has no effect and can be ignored.
-The effect of the controlled-S gate is to apply a phase factor of $i$ to the amplitudes of computational basis states where both controls on.
-We happen to be computing a qubit storing whether or not both controls are on, so we can replace the controlled-S with an uncontrolled S gate on the ancilla.
+First, the computation step is introducing phase errors (the same ones as in \autoref{fig:bad-phase-toffoli}).
+In general, correcting these phase errors requires applying a controlled-S gate to the two controls, and pre-applying a controlled-Z gate involving the target.
+However, we know our target is $|0\rangle$ beforehand, so the pre-applied controlled-Z has no effect and can be ignored.
+Furthermore, the effect of the controlled-S gate is to apply a phase factor of $i$ to the amplitudes of computational basis states where both controls on.
+Our output happens to be a qubit storing whether or not both controls are on, so we can replace the controlled S gate on the controls with an uncontrolled S gate on the ancilla.
 This corrects all of the phase errors that were being introduced.
 
-Second, the construction from the previous section is unnecessarily deep.
-Each T gate is hiding a measurement and an S gate performed conditional on the outcome of that measurement.
+Second, the construction from the previous section is unnecessarily serial.
+In the surface code, every T gate hides a measurement and an S gate performed conditional on the outcome of that measurement.
 It is beneficial to be waiting for several measurement results in parallel, instead of one after another serially.
-As shown in \cite{Jones2013}, the T gates of a Toffoli can all be performed at the same time, and we do so.
-Note that still pass an $|A\rangle$ state into the circuit instead of performing a fourth T gate, because passing in an $|A\rangle$ state already doesn't involve measuring and reacting.
+As shown in \cite{Jones2013}, the T gates of a Toffoli can all be performed at the same time, and we follow their lead.
+Note that we still pass an $|A\rangle$ state into the circuit instead of performing a fourth T gate, because passing in an $|A\rangle$ state already doesn't involve measuring and reacting.
 
-Third, taking another hint from \cite{Jones2013}, we can uncompute the ancilla by using a measure-and-correct process instead of with a mirror of the computation circuit.
-To uncompute the ancilla produced by the logical AND, we use a technique we refer to as ``erasure".
-We note that a Toffoli gate would obviously clear the value.
-Then, since the cleared ancilla was going to be discarded anyways, we are free to apply a Hadamard gate and a measurement to it.
+Third, taking another hint from \cite{Jones2013}, we can uncompute the ancilla by using a measure-and-correct process instead of with a mirror of the circuit that computed the ancilla.
+To explain the uncomputatoin process, we start with a circuit that obviously performs the uncomputation and then perform simple circuit moves until we arrive at the intended destination.
+Our starting point for the uncomputation circuit is simply a Toffoli gate, which obviously clears the ancilla since it was computed with a Toffoli and Toffolis are their own inverse.
+Then, since the cleared ancilla is going to be discarded anyways, we are free to apply a Hadamard gate and a measurement to it after the Toffoli but before discarding it.
 We then hop the Hadamard over the Toffoli, transforming it into a CCZ operation.
 Next we re-arrange the CCZ so that the ancilla is one of the controls, which is possible because the controls and targets of a CCZ are interchangeable.
-We then hop the Measurement over the CCZ, turning the quantum control into a classical control.
-Ultimately, the Toffoli gate we started with has been turned into a CZ gate that we apply or not based on the outcome of the measurement.
-All of the remaining gates are Clifford operations, which have no T-cost.
+Finally, we invoke the deferred measurement principle \cite{Nielsen2009} to hop the Measurement over the CCZ, turning the quantum control into a classical control.
+Ultimately, the Toffoli gate we started with has been turned into a CZ gate that we apply conditional on the result of measuring the ancilla.
+The final circuit uses only Clifford operations, which have no T-cost.
 
@@
 The creation process has a T-cost of 4.
-Then we can keep the ancilla around as long as we want, using it as a control for operations that would otherwise have been conditioned on both input qubits.
+Then we can keep the ancilla around, using it as a control for operations that would otherwise have been conditioned on both input qubits.
 Finally, once the ancilla is no longer needed, we can erase it at no T-cost.
@@
 the ancilla being initialized) as an ancilla wire emerging vertically from two controls then heading rightward (see \autoref{fig:compute-logical-and}).
-We draw the uncomputation analogously, with the ancilla wire coming in from the left then merging vertically into the two controls that created it (seee \autoref{fig:erase-logical-and}).
-The initialized lifetime of the ancilla qubit corresponds to the span of the wire coming out of and later merging into the control qubits.
+We draw the uncomputation analogously, with the ancilla wire coming in from the left then merging vertically into the two controls that created it (see \autoref{fig:erase-logical-and}).
+The initialized lifetime of the ancilla qubit corresponds to the wire coming out of and later merging into the control qubits.
 
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &x &&\ctrl{1} &\qw & x  && &&          &&\targ    &\ctrl{2}&\qw     &\qTi &\qw     &\ctrl{2}&\targ     &\qw &\qw &\qw &\\
@@
   \caption{
-	Computing the logical-and of two qubits, with a T-cost of 4 and a measure-react depth of 1.
+	Computing the logical-and of two qubits, with a T-cost of 4 and a T-depth of 1.
   }
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &x  &&\ctrl{1} &\qw &x && &&\qw &\qw    &\ctrl{1} &\qw \\
@@
   \caption{
-	Uncomputing the logical-and of two qubits, with a T-cost of 0 and a measure-react depth of 1.
+	Uncomputing the logical-and of two qubits, with a T-cost of 0.
   }
@@
 
-We can perform a single Toffoli by computing the logical-and of two qubits , applying a CNOT from the logical-and onto the actual target, then uncomputing the logical-and.
-This combined construction, shown in \autoref{fig:merge-use-erase-toffoli}, is equivalent to the one in \cite{Jones2013}.
+We can perform a single Toffoli by computing the logical-and of two qubits, applying a CNOT from the logical-and onto the actual target, then uncomputing the logical-and.
+This combined construction, shown in \autoref{fig:merge-use-erase-toffoli}, is equivalent to the one from \cite{Jones2013}.
 
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\targ    &\ctrl{2}&\qw     &\qTi &\qw     &\ctrl{2}&\targ     &\qw &\qw &\qw      &\qw &\qw    &\ctrl{1}         &\qw \\
@@
 
-Conceptually, our improvement on the construction in \cite{Jones2013} is merely noticing that it can be split it useful pieces that don't all need to be applied at the same time.
-In particular, when a Toffoli is later uncomputed by a second Toffoli, we simply delay erasing the ancilla wire as shown in \autoref{fig:paired-toffoli-to-logical-and}.
-This reduces the T-cost of such paired Toffolis from 8 to 4.
+In hindsight, our improvement on the construction in \cite{Jones2013} is merely noticing that it can be split it useful pieces that don't all need to be applied at the same time.
+In particular, when a Toffoli is later uncomputed by a second Toffoli, simply delaying erasing the ancilla wire as shown in \autoref{fig:paired-toffoli-to-logical-and} reduces the T-cost of such paired Toffolis from 8 to 4.
 
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
@@
 
-\section{T-optimized Circuit Constructions}
+\section{Applications}
 \label{sec:circuit-constructions}
@@
 
-First, we note that existing adder constructions contain temporary Toffolis.
-The T-cost of adding two $n$-bit numbers was thought to be $8n + O(1)$.
-We replace the basic adder unit with the building block shown in \ref{fig:full-adder-block}, putting it together as shown in \ref{fig:multi-bit-adder-example} in order to create an adder with a T-cost of $4n + O(1)$.
+Existing adder constructions such as the Cuccaro adder \cite{Cuccaro2004} contain many Toffolis that are later uncomputed by another Toffoli.
+Our improved construction does not fundamentally change the structure of the adder.
+We simply perform the temporary Toffoli gates with our temporary AND gate construction, halving the cost.
+\autoref{fig:full-adder-block} shows the per-bit building block of our adder construction.
+\autoref{fig:multi-bit-adder-example} shows how to tile the building block into a five bit adder.
 
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &c_k &&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\ctrl{1}&\qw     &\qw &c_k     &&&&& \\
@@
 	Adder building block with a T-cost of 4.
+	Equivalent to the building block of the Cuccaro adder \cite{Cuccaro2004}, but using our temporary AND gate construction.
   }
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &i_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\qw &i_0     &&&&&\\
@@
   \caption{
-	5-bit adder with T-cost of 16.
-	General $n$-bit adder has T-cost of $4n - 4$.
+	A 5-bit adder with T-cost of 16.
+	The circuit contains Clifford operations, four logical-AND computations each with a T-cost of 4, and four logical-AND erasures with no T-cost.
+	Generalizes to an $n$-bit adder with a T-cost of $4n - 4$.
   }
@@
 
+With a slight tweak to the adder building block, we can make the sum available for use as soon as the carry signal hits the high bit (instead of needing to wait for the uncomputation sweep to finish).
+This can halve the T-cost of the addition when it is going to be uncomputed.
+Instead of paying $4n + O(1)$ T gates to compute the addition, and then $4n + O(1)$ more T gates to uncompute the addition, we simply work with the intermediate state of a single addition computation.
+\autoref{fig:temporary-full-adder-block} shows the tweaked building block, which is also a temporary three-bit adder with T-cost 4.
+
 \begin{figure}
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &c_k &&\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw &&c_k     &&& \\
-      &i_k &&\targ   &\ctrl{1}&\qw     &\ctrl{1}&\qw &&i_k     &&& \\
-      &t_k &&\targ   &\ctrl{1}&\qw     &\targ   &\qw &&&(t+i)_k && \\
-      &    &&        &        &\targ   &\qw     &\qw &&c_{k+1} &&& \\
+    \Qcircuit @R=0.7em @C=0.7em {
+      &a &&\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw &a          &&&&&&& \\
+      &b &&\targ   &\ctrl{1}&\qw     &\ctrl{1}&\qw &&b \oplus a &&&&&& \\
+      &c &&\targ   &\ctrl{1}&\qw     &\targ   &\qw &&&&(a+b+c)_0  &&&& \\
+      &  &&        &        &\targ   &\qw     &\qw &&&&(a+b+c)_1  &&&& \\
     }
@@
   \caption{
-	  Temporary addition building block with a T-cost of 4 to compute and a T-cost of 0 to uncompute (not shown).
+	  A three-bit adder with T-cost 4 that can later be uncomputed for no T-cost.
+	  A building-block for temporary addition circuits whose result will be used but then uncomputed.
   }
@@
 
-We also note how to transform our adder into a controlled adder by controlling one of the operations.
-The cost of this controlled adder is $8n + O(1)$; a significant improvement over a recently proposed construction with a T-cost of $21n + O(1)$.
+Some additions, such as the ones performed in Shor's algorithm, are conditioned on a control qubit.
+\cite{Coreas2017} derives a controlled-addition construction with T-cost of $21n + O(1)$.
+We use our temporary AND gate construction to improve this to $8n + O(1)$, and show the building block in \autoref{fig:controlled-full-adder-block}.
 
@@
   \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
+    \Qcircuit @R=0.7em @C=0.7em {
       &\text{control} &&&&\qw     &\qw     &\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\qw     &\ctrl{2}&\qw     &\qw &\\
@@
 
-\begin{figure}
-  \centering
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=1.5em @C=0.7em {
-      &\text{control} &&&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\ctrl{13}&\qw     &\qw     &\qw     &\ctrl{10}&\qw     &\qw     &\ctrl{7}&\qw     &\qw     &\qw     &\ctrl{4}&\qw     &\qw     &\ctrl{1}&\qw &&&\text{control}               &&&&&&\\
-      &&i_0            &&&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\qw &i_0                          &&&&&&&&\\
-      &&t_0            &&&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw &&&&&&(t+i \cdot \text{control})_0 &&&\\
-      &&               &&&         &\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\qw     &\ctrl{2}&\qw     &        &    &&&                             &&&&&&\\
-      &&i_1            &&&\qw      &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\targ   &\qw     &\qw     &\qw &i_1                          &&&&&&&&\\
-      &&t_1            &&&\qw      &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\targ   &\qw     &\qw     &\qw &&&&&&(t+i \cdot \text{control})_1 &&&\\
-      &&               &&&         &        &        &\targ   &\qw &\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\ctrl{3}&\qw     &\qw     &\ctrl{2}&\targ   &\qw     &        &        &        &        &    &&&                             &&&&&&\\
-      &&i_2            &&&\qw      &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\ctrl{1}&\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &i_2                          &&&&&&&&\\
-      &&t_2            &&&\qw      &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw      &\qw     &\ctrl{1}&\targ   &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &&&&&&(t+i \cdot \text{control})_2 &&&\\
-      &&               &&&         &        &        &        &    &        &        &\targ   &\qw &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw      &\qw     &\ctrl{3}&\qw     &\qw      &\ctrl{2}&\qw     &        &        &        &        &        &        &        &        &    &&&                             &&&&&&\\
-      &&i_3            &&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw     &\qw      &\qw     &\qw     &\ctrl{1}&\ctrl{1} &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &i_3                          &&&&&&&&\\
-      &&t_3            &&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw     &\qw      &\qw     &\qw     &\ctrl{1}&\targ    &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &&&&&&(t+i \cdot \text{control})_3 &&&\\
-      &&               &&&         &        &        &        &    &        &        &        &    &        &        &\targ   &\ctrl{1}&\qw      &\ctrl{1}&\targ   &\qw     &         &        &        &        &        &        &        &        &        &        &        &    &&&                             &&&&&&\\
-      &&i_4            &&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\targ   &\ctrl{1} &\targ   &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &i_4                          &&&&&&&&\\
-      &&t_4            &&&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\targ    &\qw     &\qw     &\qw     &\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw &&&&&&(t+i \cdot \text{control})_4 &&&\\
-    }
-  }
-  \caption{
-	5-bit controlled adder with a T-cost of 36.
-	General $n$-bit controlled adder has T-cost of $8n - 4$.
-  }
-  \label{fig:ancilla-addition}
-\end{figure}
+The temporary AND gate is useful for more than just arithmetic.
+For example, temporary ANDs can be useful for applying phase rotations to many qubits simultaneously.
+Given a $b$-bit ancilla register $G$ prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle$ (a `phase gradient state'), using our adder construction to add a register $Q$ into $G$ will cause phase kickback that applies the operation $\text{Grad} = \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle \langle k|$ to $V$.
+The $\text{Grad}$ operation is equivalent to applying the phase gate $Z^{2^-k}$ to the qubit position at $k$ for each qubit within $Q$.
+Some quantum Fourier transform circuits involve conditional uses of $\text{Grad}$, so it is possible that temporary-ANDs improve the T-cost of those circuits.
 
-Other examples of operations which benefit from this optimization include:
+Other examples of operations which benefit from cheaper temporary AND gates include:
 
@@
 
+- Integer arithmetic in general.
+
 - Modular arithmetic.
+
+- Expanding a binary register into a unary register.
 
@@
 
-- Expanding a binary register into a unary register.
-
-- Phasing a register by a computable function $f$ of its value (i.e. applying the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$).
+- Phasing a register by a computable function $f$ (i.e. applying the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$).
 
@@
 
-- etc
+We leave further low hanging fruit as exercises for the reader.
 
@@
 
-It was commonly believed that the optimal number of T gates to perform addition was 8, or 7 \cite{AustinDiscussionsAndEmails2017}.
-We showed that the number of T gates is actually 4.
-We are not aware of any method for improving the T cost further, but we do prove the temporary AND gatea lower bound of 2 In \autoref{fig:lower-bound} we show a lower bound of 2
+For over a decade, the T-cost of addition has been $8n + O(1)$ \cite{Barenco1995, Cuccaro2004, Amy2013}).
+This was believed to be optimal \cite{AustinDiscussionsAndEmails2017}.
+In this paper, we halved the T-cost of addition to $4n + O(1)$.
+We did so with a construction that applies to many other circuits: the temporary AND gate.
+
+Our result raises the question of whether further T-cost improvements to addition or to temporary AND gates are possible.
+A simple method for proving lower bounds on the T-cost of an operation is to synthesize T gates by applying that operation.
+In \autoref{fig:lower-bound-temporary-and} we use a T gate and a single temporary-AND to synthesize three T gates.
+This proves that the T-cost of a temporary-AND is at least two.
+Two is less than four, but we are not aware of any method for further reducing the T-cost of temporary-ANDs or for synthesizing more T gates out of a temporary-AND.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &\qT &\qw && &&&\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw &\qw     &\ctrl{3}&\qw     &\ctrl{2}&\qw &\\
+      &\qT &\qw &&=&&&\targ   &\ctrl{1}&\qw     &\ctrl{1}&\qw &\ctrl{1}&\qw     &\ctrl{1}&\targ   &\qw &\\
+      &\qT &\qw && &&&\targ   &\ctrl{1}&\targ   &\targ   &\qT &\targ   &\targ   &\ctrl{1}&\targ   &\qw &\\
+      &    &    && &&&        &        &\targ   &\qw     &\qS &\qw     &\targ   &\qw     &        &    &\\
+    }
+  }
+  \caption{
+	Three T gates can be synthesized by performing a temporary-AND, one T gate, and various Clifford operations.
+	Clifford operations can't produce T gates, so the extra two gates must be coming from the temporary-AND.
+	This proves that at least two T gates went into performing the temporary-AND.
+  }
+  \label{fig:lower-bound-temporary-and}
+\end{figure}
+
+Our work represents a significant and widely applicable improvement to the state of the art in synthesizing quantum circuits with low T gate counts.
 
@@
 
-Austin [reference finding, idea bouncing, review].
+We thank Austin Fowler for assistance in locating relevant references, sampling the opinions of researchers, and for comments that greatly improved the paper.
 
@@
 
-https://arxiv.org/pdf/quant-ph/9503016.pdf   ->   the paired Toffoli construction that uses 4T
-
-https://arxiv.org/pdf/cond-mat/9409111.pdf   -> shows CCNOT can be done with three 2-qubit interactions, ignores single-qubit???
-
-https://arxiv.org/pdf/1308.4134.pdf -> proof of 7T gate minimum with no ancilla
-
-https://arxiv.org/pdf/1212.5069.pdf  ->   does a single Toffoli with four T gates by turning one of the controls classical
-
-https://arxiv.org/pdf/1210.0974.pdf  ->   a phase-neglected 4T gate construction that only uses a single layer of T gates
-
-Quantum Computation and Quantum Information -> shows standard decomposition into 7T gates. But do they say it's optimal?
-
-https://arxiv.org/pdf/1206.0758.pdf -> just an example of the 7T construction;;; probablynot relevant
-
->>>>>>>>>>>>>>>>>>>>>>> https://arxiv.org/pdf/1206.0758.pdf          paper that says addition costs 8N T
-
-https://arxiv.org/pdf/1706.05113.pdf   has a ctrl-add with 21N + O(1). This technique+commutators achieves 8N.
-
-https://pdfs.semanticscholar.org/8b84/d1bc2928922937a16205f9d8c925ff76689b.pdf     has a pretty good ripple carry
-
 \end{document}

### 2017-08-26 paper.tex
Diagram tweaks, consistency

@@
 
-In addition to our $n$-bit adder circuit with T-cost of $4n-4$, we construct an $n$-bit controlled adder circuit with T-cost of $8n-4$, a temporary adder circuit that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed at no T-cost, a phase gradient circuit with T-cost of $4n + O(b \lg n)$ where $b = \lg \frac{1}{\epsilon}$ is the desired bit-precision of the operation, a quantum Fourier transform circuit with T-cost of $8 n \lg b + O(b (\lg n) (\lg b))$, and a circuit to rotate $n$ qubits by any angle $\theta$ with T-cost of $4n + O(b \lg n)$.
+In addition to our $n$-bit adder circuit with T-cost of $4n + O(1)$, we show how to construct an $n$-bit controlled adder circuit with T-cost of $8n + O(1)$, and a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed at no T-cost.
 \end{abstract}
@@
   \caption{
-	Adder building block with a T-cost of 4.
+	Building block, for an adder, with a T-cost of 4.
 	Equivalent to the building block of the Cuccaro adder \cite{Cuccaro2004}, but using our temporary AND gate construction.
@@
     \Qcircuit @R=0.7em @C=0.7em {
-      &\text{control} &&&&\qw     &\qw     &\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\qw     &\ctrl{2}&\qw     &\qw &\\
-      &&c_k            &&&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\qw     &\ctrl{2}&\qw &\\
-      &&i_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\ctrl{1}&\targ   &\qw &\\
-      &&t_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\targ   &\targ   &\qw &\\
-      &               &&&&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &    &\\
+      &\text{control} &&&&\qw     &\qw     &\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\qw     &\ctrl{2}&\qw     &\qw &&&\text{control}               &&&&&&\\
+      &&c_k            &&&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\qw     &\ctrl{2}&\qw &c_k                          &&&&&&&&\\
+      &&i_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\ctrl{1}&\targ   &\qw &i_k                          &&&&&&&&\\
+      &&t_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\targ   &\targ   &\qw &&&&&(t+i \cdot \text{control})_k &&&&&\\
+      &               &&&&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &    &    &&&&&&&&\\
     }
@@
   \caption{
-	Controlled full adder block with a T-cost of 8.
+	Building block, for a controlled adder, with a T-cost of 8.
   }

### 2017-08-28 paper.tex
Applying Austin's comments: - Shortened title to be punchy - Fixed logical-and -> logical-AND - T-cost -> T-count

@@
 
-\title{T gates, temporary ANDs, and halving the cost of quantum addition}
+\title{Halving the cost of quantum addition}
 \author{Craig Gidney}
@@
 We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Cuccaro2004, Amy2013, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
-We do so via a ``temporary AND gate" construction, which uses four T gates to store the logical-and of two qubits into an ancilla and zero T gates to later erase the ancilla.
+We do so via a ``temporary AND gate" construction, which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
 
-Temporary AND gates are a generally useful tool when optimizing T-cost.
+Temporary AND gates are a generally useful tool when optimizing T-count.
 They can be applied to integer arithmetic, modular arithmetic, rotation synthesis, the quantum Fourier transform, Shor's algorithm, Grover oracles, and many other circuits.
@@
 
-In addition to our $n$-bit adder circuit with T-cost of $4n + O(1)$, we show how to construct an $n$-bit controlled adder circuit with T-cost of $8n + O(1)$, and a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed at no T-cost.
+In addition to our $n$-bit adder circuit with T-count of $4n + O(1)$, we show how to construct an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, and a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed at no T-count.
 \end{abstract}
@@
 \autoref{sec:review} discusses how previous work on optimizing T gate counts improved the cost of matched pairs of Toffolis from 14 to 8.
-In \autoref{sec:invest}, we show how to improve the T-cost from 8 to 6 by investing and later recovering an $|A\rangle$ state instead of consuming it to perform T gates.
-\autoref{sec:temporary-and} expands on those ideas with an ancilla erasing trick that improves the T-cost from 6 to 4.
-This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to create low T-cost circuits for several addition tasks and discuss other possible applications of our constructions.
+In \autoref{sec:invest}, we show how to improve the T-count from 8 to 6 by investing and later recovering an $|A\rangle$ state instead of consuming it to perform T gates.
+\autoref{sec:temporary-and} expands on those ideas with an ancilla erasing trick that improves the T-count from 6 to 4.
+This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to create low T-count circuits for several addition tasks and discuss other possible applications of our constructions.
 Finally, \autoref{sec:conclusion} summarizes our contributions, and proves that at least two T gates are needed to perform a temporary AND.
@@
   \caption{
-	The T-cost of $N$ adjacent Toffolis sharing the same controls is $0 \cdot N + O(1)$.
-	The marginal T-cost is 0 because each additional Toffoli can be replaced with CNOTs framing a root Toffoli.
+	The T-count of $N$ adjacent Toffolis sharing the same controls is $0 \cdot N + O(1)$.
+	The marginal T-count is 0 because each additional Toffoli can be replaced with CNOTs framing a root Toffoli.
   }
@@
   \caption{
-	Starting with \autoref{fig:textbook-toffoli} then dropping T gates on the controls produces an operation with a T-cost of 4 that performs the correct permutation.
+	Starting with \autoref{fig:textbook-toffoli} then dropping T gates on the controls produces an operation with a T-count of 4 that performs the correct permutation.
 	However, the operation introduces phase errors.
@@
 	When two Toffolis form a compute/uncompute pair, they can cancel each others' phase errors.
-	Fixes the problem with the construction in \autoref{fig:bad-phase-toffoli}, and achieves a per-Toffoli T-cost of 4 for paired Toffolis \cite{Barenco1995}.
+	Fixes the problem with the construction in \autoref{fig:bad-phase-toffoli}, and achieves a per-Toffoli T-count of 4 for paired Toffolis \cite{Barenco1995}.
   }
@@
 
-A typical $n$-bit quantum adder will contain $2n + O(1)$ Toffoli gates, implying a naive T-cost of $14n + O(1)$.
+A typical $n$-bit quantum adder will contain $2n + O(1)$ Toffoli gates, implying a naive T-count of $14n + O(1)$.
 However, almost all of the Toffoli gates in the first half of an adder are uncomputed by Toffoli gates in the second half.
-This allows the T gates on the controls of the Toffoli gates to be omitted, reducing their T-cost from 7 to 4 and the T-cost of addition to $8n + O(1)$.
+This allows the T gates on the controls of the Toffoli gates to be omitted, reducing their T-count from 7 to 4 and the T-count of addition to $8n + O(1)$.
 
-Even if a Toffoli isn't paired with a second Toffoli that uncomputes its effects, it's still possible to perform the Toffoli with T-cost of by using an ancilla, a measurement, and a conditional fixup operation \cite{Jones2013} (see \autoref{fig:jones-toffoli}).
-Our key improvement over \cite{Jones2013} is to notice that the ancilla's lifetime can be extended, so it can be used both when computing and uncomputing a Toffoli.
+Even if a Toffoli isn't paired with a second Toffoli that uncomputes its effects, it is still possible to perform the Toffoli with T-count of 4 by using an ancilla, a measurement, and a conditional fixup operation \cite{Jones2013} (see \autoref{fig:jones-toffoli}).
 
@@
   \caption{
-	Toffoli construction with T-cost of 4 from \cite{Jones2013}.
+	Toffoli construction with T-count of 4 from \cite{Jones2013}.
   }
@@
 
-\section{Warm up: Investing $|A\rangle$ states}
+\section{Warm up: Reusing $|A\rangle$ states}
 \label{sec:invest}
 
-This paper started with an idea that improved the T-cost of a single Toffoli gate from 7 to 6 in a surprising way, by performing the Toffoli indirectly instead of directly.
-We started with a clean ancilla in the $|0\rangle$ state, applied a Toffoli to store the logical-and of the two controls in the ancilla, used the ancilla to control a CNOT onto the intended target, then uncomputed the ancilla.
+This paper started with an idea that improved the T-count of a single Toffoli gate from 7 to 6 in a surprising way, by performing the Toffoli indirectly instead of directly.
+We started with a clean ancilla in the $|0\rangle$ state, applied a Toffoli to store the logical-AND of the two controls in the ancilla, used the ancilla to control a CNOT onto the intended target, then uncomputed the ancilla.
 
-Initially, this indirect-Toffoli construction appears to have a T-cost of 8.
+Initially, this indirect-Toffoli construction appears to have a T-count of 8.
 However, as shown in \autoref{fig:indirect-toffoli}, the last T gate in the circuit is unnecessary.
 In fact, it is actively harmful.
-Dropping this T gate and the following Hadamard not only reduces the T-cost from 8 to 7, it leaves the ancilla in an $|A\rangle$ state that can be consumed to perform a T gate elsewhere.
-This improves the net T-cost to 6.
+Dropping this T gate and the following Hadamard not only reduces the T-count from 8 to 7, it leaves the ancilla in an $|A\rangle$ state that can be consumed to perform a T gate elsewhere.
+This improves the net T-count to 6.
 
@@
   \caption{
-	Performing a Toffoli indirectly, by applying a pair of Toffolis to an ancilla and using the intermediate value to control a CNOT onto the actual target, appears to have a T-cost of 8.
+	Performing a Toffoli indirectly, by applying a pair of Toffolis to an ancilla and using the intermediate value to control a CNOT onto the actual target, appears to have a T-count of 8.
 	However, the initial T gate is being used to compute an $|A\rangle$ state and the final T gate is being used to uncompute the $|A\rangle$ state.
     Instead of spending $|A\rangle$ states to perform T gates to make and unmake $|A\rangle$ states, we pass in and later recover a single $|A\rangle$ state.
-	This ``investment" reduces the net T-cost to 6.
+	This ``investment" reduces the net T-count to 6.
   }
@@
 
-A Toffoli with a T-cost of 6 isn't optimal, but the reduction from 7 was done in an unexpected way.
+A Toffoli with a T-count of 6 isn't optimal, but the reduction from 7 was done in an unexpected way.
 While searching for other circuits where this optimization might apply, we realized it would be useful when an initial Toffoli was later uncomputed by a second Toffoli.
 Instead of computing and uncomputing the ancilla for the first Toffoli, then recomputing and re-uncomputing the ancilla for the second Toffoli, we simply kept the ancilla around until it was time to uncompute the Toffoli.
-This halved the T-cost of the pair from 12 to 6.
+This halved the T-count of the pair from 12 to 6.
 See \autoref{fig:ancilla-temporary-toffoli}.
@@
   \caption{
-	Computing and uncomputing a Toffoli gate with a net T-cost of 6.
-	Invests an $|A\rangle$ state when computing the Toffoli, holds an ancilla qubit storing the logical-and of the controls until it's time to uncompute the Toffoli, then recovers the $|A\rangle$ state.
+	Computing and uncomputing a Toffoli gate with a net T-count of 6.
+	Invests an $|A\rangle$ state when computing the Toffoli, holds an ancilla qubit storing the logical-AND of the controls until it is time to uncompute the Toffoli, then recovers the $|A\rangle$ state.
   }
@@
 Many circuits involve computing and later uncomputing a Toffoli (e.g. addition).
-Previously, it was believed that each Toffoli had a T-cost of 4 so the pair would have a T-cost of 8.
-Our $|A\rangle$-investing construction state reduces the T-cost of the pair from 8 to 6.
+Previously, it was believed that each Toffoli had a T-count of 4 so the pair would have a T-count of 8.
+Our $|A\rangle$-investing construction state reduces the T-count of the pair from 8 to 6.
 
@@
 Ultimately, the Toffoli gate we started with has been turned into a CZ gate that we apply conditional on the result of measuring the ancilla.
-The final circuit uses only Clifford operations, which have no T-cost.
+The final circuit uses only Clifford operations, which have no T-count.
 
@@
 We can initialize an ancilla storing the logical-AND of two qubits.
-The creation process has a T-cost of 4.
+The creation process has a T-count of 4.
 Then we can keep the ancilla around, using it as a control for operations that would otherwise have been conditioned on both input qubits.
-Finally, once the ancilla is no longer needed, we can erase it at no T-cost.
+Finally, once the ancilla is no longer needed, we can erase it at no T-count.
 
 In diagrams, we draw the computation of the AND gate (i.e.
-the ancilla being initialized) as an ancilla wire emerging vertically from two controls then heading rightward (see \autoref{fig:compute-logical-and}).
-We draw the uncomputation analogously, with the ancilla wire coming in from the left then merging vertically into the two controls that created it (see \autoref{fig:erase-logical-and}).
+the ancilla being initialized) as an ancilla wire emerging vertically from two controls then heading rightward (see \autoref{fig:compute-logical-AND}).
+We draw the uncomputation analogously, with the ancilla wire coming in from the left then merging vertically into the two controls that created it (see \autoref{fig:erase-logical-AND}).
 The initialized lifetime of the ancilla qubit corresponds to the wire coming out of and later merging into the control qubits.
@@
   \caption{
-	Computing the logical-and of two qubits, with a T-cost of 4 and a T-depth of 1.
+	Computing the logical-AND of two qubits, with a T-count of 4 and a T-depth of 1.
   }
-  \label{fig:compute-logical-and}
+  \label{fig:compute-logical-AND}
 \end{figure}
@@
   \caption{
-	Uncomputing the logical-and of two qubits, with a T-cost of 0.
+	Uncomputing the logical-AND of two qubits, with a T-count of 0.
   }
-  \label{fig:erase-logical-and}
+  \label{fig:erase-logical-AND}
 \end{figure}
 
-We can perform a single Toffoli by computing the logical-and of two qubits, applying a CNOT from the logical-and onto the actual target, then uncomputing the logical-and.
+We can perform a single Toffoli by computing the logical-AND of two qubits, applying a CNOT from the logical-AND onto the actual target, then uncomputing the logical-AND.
 This combined construction, shown in \autoref{fig:merge-use-erase-toffoli}, is equivalent to the one from \cite{Jones2013}.
@@
   \caption{
-	Performing a Toffoli gate by computing the logical-and of its controls, using the result, then erasing it.
-	Has a T-cost of 4 and a measure-react depth of 2.
+	Performing a Toffoli gate by computing the logical-AND of its controls, using the result, then erasing it.
+	Has a T-count of 4 and a measure-react depth of 2.
 	Equivalent to the construction from \cite{Jones2013} shown in \autoref{fig:jones-toffoli}.
@@
 In hindsight, our improvement on the construction in \cite{Jones2013} is merely noticing that it can be split it useful pieces that don't all need to be applied at the same time.
-In particular, when a Toffoli is later uncomputed by a second Toffoli, simply delaying erasing the ancilla wire as shown in \autoref{fig:paired-toffoli-to-logical-and} reduces the T-cost of such paired Toffolis from 8 to 4.
+In particular, when a Toffoli is later uncomputed by a second Toffoli, simply delaying erasing the ancilla wire as shown in \autoref{fig:paired-toffoli-to-logical-AND} reduces the T-count of such paired Toffolis from 8 to 4.
 
@@
 	Replacing a pair of compute/uncompute Toffolis with a temporary AND gate and Clifford operations.
-	Improves the T-cost from 8 to 4.
+	Improves the T-count from 8 to 4.
   }
-  \label{fig:paired-toffoli-to-logical-and}
+  \label{fig:paired-toffoli-to-logical-AND}
 \end{figure}
@@
   \caption{
-	Building block, for an adder, with a T-cost of 4.
+	Building block, for an adder, with a T-count of 4.
 	Equivalent to the building block of the Cuccaro adder \cite{Cuccaro2004}, but using our temporary AND gate construction.
@@
   \caption{
-	A 5-bit adder with T-cost of 16.
-	The circuit contains Clifford operations, four logical-AND computations each with a T-cost of 4, and four logical-AND erasures with no T-cost.
-	Generalizes to an $n$-bit adder with a T-cost of $4n - 4$.
+	A 5-bit adder with T-count of 16.
+	The circuit contains Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures with no T-count.
+	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
   }
@@
 With a slight tweak to the adder building block, we can make the sum available for use as soon as the carry signal hits the high bit (instead of needing to wait for the uncomputation sweep to finish).
-This can halve the T-cost of the addition when it is going to be uncomputed.
+This can halve the T-count of the addition when it is going to be uncomputed.
 Instead of paying $4n + O(1)$ T gates to compute the addition, and then $4n + O(1)$ more T gates to uncompute the addition, we simply work with the intermediate state of a single addition computation.
-\autoref{fig:temporary-full-adder-block} shows the tweaked building block, which is also a temporary three-bit adder with T-cost 4.
+\autoref{fig:temporary-full-adder-block} shows the tweaked building block, which is also a temporary three-bit adder with T-count 4.
 
@@
   \caption{
-	  A three-bit adder with T-cost 4 that can later be uncomputed for no T-cost.
+	  A three-bit adder with T-count 4 that can later be uncomputed for no T-count.
 	  A building-block for temporary addition circuits whose result will be used but then uncomputed.
@@
 Some additions, such as the ones performed in Shor's algorithm, are conditioned on a control qubit.
-\cite{Coreas2017} derives a controlled-addition construction with T-cost of $21n + O(1)$.
+\cite{Coreas2017} derives a controlled-addition construction with T-count of $21n + O(1)$.
 We use our temporary AND gate construction to improve this to $8n + O(1)$, and show the building block in \autoref{fig:controlled-full-adder-block}.
@@
   \caption{
-	Building block, for a controlled adder, with a T-cost of 8.
+	Building block, for a controlled adder, with a T-count of 8.
   }
@@
 The $\text{Grad}$ operation is equivalent to applying the phase gate $Z^{2^-k}$ to the qubit position at $k$ for each qubit within $Q$.
-Some quantum Fourier transform circuits involve conditional uses of $\text{Grad}$, so it is possible that temporary-ANDs improve the T-cost of those circuits.
+Some quantum Fourier transform circuits involve conditional uses of $\text{Grad}$, so it is possible that temporary-ANDs improve the T-count of those circuits.
+
+Our constructions reduce many existing estimates of the cost of quantum computation.
+For example, \cite{Fowler2012} estimates that factoring a 2000-bit number would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$.
+Their time estimate is based on each Toffoli having a T-depth of 1, and their $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
+Because Shor's algorithm is dominated by the cost of additions, our techniques (and previous work) multiply the T-count by $\frac{4}{14}$ and the T-depth by $\frac{1}{3}$.
+This reduces the estimate to 9 hours and $6 \cdot 10^{11}$ distilled $|A\rangle$ states.
 
@@
 
-For over a decade, the T-cost of addition has been $8n + O(1)$ \cite{Barenco1995, Cuccaro2004, Amy2013}).
+For over a decade, the T-count of addition has been $8n + O(1)$ \cite{Barenco1995, Cuccaro2004, Amy2013}).
 This was believed to be optimal \cite{AustinDiscussionsAndEmails2017}.
-In this paper, we halved the T-cost of addition to $4n + O(1)$.
+In this paper, we halved the T-count of addition to $4n + O(1)$.
 We did so with a construction that applies to many other circuits: the temporary AND gate.
 
-Our result raises the question of whether further T-cost improvements to addition or to temporary AND gates are possible.
-A simple method for proving lower bounds on the T-cost of an operation is to synthesize T gates by applying that operation.
+Our result raises the question of whether further T-count improvements to addition or to temporary AND gates are possible.
+A simple method for proving lower bounds on the T-count of an operation is to synthesize T gates by applying that operation.
 In \autoref{fig:lower-bound-temporary-and} we use a T gate and a single temporary-AND to synthesize three T gates.
-This proves that the T-cost of a temporary-AND is at least two.
-Two is less than four, but we are not aware of any method for further reducing the T-cost of temporary-ANDs or for synthesizing more T gates out of a temporary-AND.
+This proves that the T-count of a temporary-AND is at least two.
+Two is less than four, but we are not aware of any method for further reducing the T-count of temporary-ANDs or for synthesizing more T gates out of a temporary-AND.
 

### 2017-08-29 paper.tex
Fix figure 5 (fig:jones-toffoli) having the wrong circuit
Second draft almost done

@@
 \documentclass[twocolumn,longbibliography]{quantumarticle-customized}
+\usepackage{amssymb}
 \usepackage{amsmath}
@@
 \begin{abstract}
-We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Cuccaro2004, Amy2013, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
-We do so via a ``temporary AND gate" construction, which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
+We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Amy2013, Cuccaro2004, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
+We do so via a ``temporary logical-AND" construction, which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
 
-Temporary AND gates are a generally useful tool when optimizing T-count.
+Temporary logical-ANDs are a generally useful tool when optimizing T-count.
 They can be applied to integer arithmetic, modular arithmetic, rotation synthesis, the quantum Fourier transform, Shor's algorithm, Grover oracles, and many other circuits.
-Because T gates dominate the cost of quantum computation based on the surface code, and the temporary AND gate is widely applicable, our constructions represent a significant reduction in projected costs of quantum computation.
+Because T gates dominate the cost of quantum computation based on the surface code, and temporary logical-ANDs are widely applicable, our constructions represent a significant reduction in projected costs of quantum computation.
 
-In addition to our $n$-bit adder circuit with T-count of $4n + O(1)$, we show how to construct an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, and a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed at no T-count.
+In addition to our $n$-bit adder circuit with T-count of $4n + O(1)$, we show how to construct an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, and a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed without T gates.
 \end{abstract}
@@
 
-In this paper, we focus on improving the number of T gates needed to perform Toffoli gates that will later be uncomputed by a second Toffoli gate.
-We then apply these improvements.
-The key ideas behind our construction are 1) perform the Toffoli indirectly by targeting a clean ancilla qubit and then using the ancilla to toggle the intended target, 2) don't uncompute and recompute the ancilla if it will be needed again, 3) if a T gate is being used to compute and/or uncompute an $|A\rangle$ state then just pass in and/or recover an $|A\rangle$ state instead, and 4) uncompute the ancilla by measuring it then applying a classically controlled fixup operation.
-We refer to initializing the ancilla as ``computing the logical-AND of the controls", to uncomputing the ancilla as ``erasing the logical-AND", and to the combination of both pieces as a ``temporary AND gate".
+The textbook construction of a Toffoli gate uses seven T gates \cite{Nielsen2009}.
+When Toffoli operations are paired, i.e.
+when an initial Toffoli operation is later uncomputed by a second Toffoli operation, each Toffoli in the pair can omit three of the T gates from the textbook construction.
+This introduces phase errors, assuming this doesn't affect intermediate operatoins, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \cite{Barenco1995}.
+It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and classically conditioned fixup operations \cite{Jones2013}.
 
-Our paper is divided into six sections.
-In \autoref{sec:introduction}, we motivate the problem and introduce terminology.
-\autoref{sec:review} discusses how previous work on optimizing T gate counts improved the cost of matched pairs of Toffolis from 14 to 8.
-In \autoref{sec:invest}, we show how to improve the T-count from 8 to 6 by investing and later recovering an $|A\rangle$ state instead of consuming it to perform T gates.
-\autoref{sec:temporary-and} expands on those ideas with an ancilla erasing trick that improves the T-count from 6 to 4.
-This leads to \autoref{sec:circuit-constructions}, where we demonstrate how to use our constructions to create low T-count circuits for several addition tasks and discuss other possible applications of our constructions.
-Finally, \autoref{sec:conclusion} summarizes our contributions, and proves that at least two T gates are needed to perform a temporary AND.
+Existing Toffoli constructions used at least 4 T gates per Toffoli gate.
+The Cuccaro adder uses $2n-1$ Toffoli gates \cite{Cuccaro2004}.
+Correspondingly, existing adder constructions have T-counts of $8n-4$ \cite{Amy2013}.
+Adders with a T-count of $8n-4$ have been known for over a decade \cite{Barenco1995, Cuccaro2004}, and were believed to be optimal \cite{AustinDiscussionsAndEmails2017}.
 
-\section{Previous Work}
-\label{sec:review}
+Our paper is divided into four sections.
+In \autoref{sec:introduction}, we motivate the problem, discuss existing work, and note the many applications of our construction.
+\autoref{sec:construction} defines our $4n + O(1)$ adder and explains its pieces.
+\autoref{sec:other-applications} covers circuits for other adding tasks and discusses other kinds of circuits where the temporary-AND construction can improve T-counts.
+Finally, \autoref{sec:conclusion} summarizes our contributions and proves that at least two T gates are needed to perform a temporary AND while discussing future work.
 
-The textbook construction of a Toffoli gate uses seven T gates \cite{Nielsen2009} (see \autoref{fig:textbook-toffoli}).
-Assuming we aren't permitted to involve other qubits or to share work with other operations, this construction is optimal \cite{Gosset2014}.
+We improve the number of T gates needed to perform an adder by halving the number of T gates needed to perform paired Toffoli gates.
+In this paper we focus on how this halves the T-count of addition, but improving the T-count of paired Toffolis is a widely applicable optimization.
+Other examples of operations whose T-count can be improved by our constructions include:
 
-When optimizations can cross qubit and gate boundaries, the textbook construction isn't optimal.
-For example, when several adjacent Toffoli gates share the same controls, all but one can be replaced by CNOT operations (see \autoref{fig:shared-controls}).
+- Integer arithmetic in general.
 
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1}&\qw & &   & & &\ctrl{1}&\qw  &\ctrl{1}&\qT &\qw     &\qw  &\ctrl{2}&\qw &\qw     &\qw  &\ctrl{2}&\qw &\qw \\
-      &\ctrl{1}&\qw & & = & & &\targ   &\qTi &\targ   &\qT &\ctrl{1}&\qw  &\qw     &\qw &\ctrl{1}&\qw  &\qw     &\qw &\qw \\
-      &\targ   &\qw & &   & & &\qw     &\qw  &\qH     &\qT &\targ   &\qTi &\targ   &\qT &\targ   &\qTi &\targ   &\qH &\qw \\
-    }
-  }
-  \caption{
-	Textbook Toffoli construction from \cite{Nielsen2009}.
-	Uses eight Clifford gates and seven T gates.
-  }
-  \label{fig:textbook-toffoli}
-\end{figure}
+- Modular arithmetic.
 
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1}&\ctrl{1}&\ctrl{1}&\ctrl{1}&\qw && &&&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw &\\
-      &\ctrl{1}&\ctrl{2}&\ctrl{3}&\ctrl{4}&\qw &&=&&&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw &\\
-      &\targ   &\qw     &\qw     &\qw     &\qw && &&&\ctrl{3}&\ctrl{2}&\ctrl{1}&\targ   &\ctrl{1}&\ctrl{2}&\ctrl{3}&\qw &\\
-      &\qw     &\targ   &\qw     &\qw     &\qw && &&&\qw     &\qw     &\targ   &\qw     &\targ   &\qw     &\qw     &\qw &\\
-      &\qw     &\qw     &\targ   &\qw     &\qw && &&&\qw     &\targ   &\qw     &\qw     &\qw     &\targ   &\qw     &\qw &\\
-      &\qw     &\qw     &\qw     &\targ   &\qw && &&&\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &\\
-    }
-  }
-  \caption{
-	The T-count of $N$ adjacent Toffolis sharing the same controls is $0 \cdot N + O(1)$.
-	The marginal T-count is 0 because each additional Toffoli can be replaced with CNOTs framing a root Toffoli.
-  }
-  \label{fig:shared-controls}
-\end{figure}
+- Applying the same rotation to many qubits.
 
-It isn't common for adjacent Toffolis to have the same controls, but it is common for a Toffoli to later be uncomputed by a second matching Toffoli (i.e. for the Toffoli's effect to be temporary).
-When this occurs, the three $T$ gates on the control qubits of the textbook construction can be omitted.
-This introduces phase errors (see \autoref{fig:bad-phase-toffoli}), but the second Toffoli gate can uncompute those errors while uncomputing the state permutation \cite{Barenco1995} (see \autoref{fig:cancelled-bad-phase-toffoli}).
+- The quantum Fourier transform.
 
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1} &\qw & &       & & &\ctrl{1}     &\ctrl{1} &\qw & &   & & &\qw &\qw &\qw      &\qw  &\ctrl{2} &\qw &\qw      &\qw  &\qw &\qw \\
-      &\ctrl{1} &\qw & & \cong & & &\qS          &\ctrl{1} &\qw & & = & & &\qw &\qw &\ctrl{1} &\qw  &\qw      &\qw &\ctrl{1} &\qw  &\qw &\qw \\
-      &\targ    &\qw & &       & & &\gate{Z}\qwx &\targ    &\qw & &   & & &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qw \\
-    }
-  }
-  \caption{
-	Starting with \autoref{fig:textbook-toffoli} then dropping T gates on the controls produces an operation with a T-count of 4 that performs the correct permutation.
-	However, the operation introduces phase errors.
-  }
-  \label{fig:bad-phase-toffoli}
-\end{figure}
+- Operations whose target is indexed by a quantum register.
 
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw && &&&\ctrl{1}     &\ctrl{1}  &\qw &\qw &\qw    &\qw &\qw     &\ctrl{1}&\ctrl{1}     &\qw & \\
-      &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw &&=&&&\qS          &\ctrl{1}  &\qw &\qw &\qw    &\qw &\qw     &\ctrl{1}&\qSi         &\qw & \\
-      &\targ   &\qw &    &\ldots &    &    &\targ   &\qw && &&&\gate{Z}\qwx &\targ     &\qw &    &\ldots &    &        &\targ   &\gate{Z}\qwx &\qw & \\
-    }
-  }
-  \caption{
-	When two Toffolis form a compute/uncompute pair, they can cancel each others' phase errors.
-	Fixes the problem with the construction in \autoref{fig:bad-phase-toffoli}, and achieves a per-Toffoli T-count of 4 for paired Toffolis \cite{Barenco1995}.
-  }
-  \label{fig:cancelled-bad-phase-toffoli}
-\end{figure}
+- Phasing a register by a computable function $f$ (i.e. applying the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$).
 
-A typical $n$-bit quantum adder will contain $2n + O(1)$ Toffoli gates, implying a naive T-count of $14n + O(1)$.
-However, almost all of the Toffoli gates in the first half of an adder are uncomputed by Toffoli gates in the second half.
-This allows the T gates on the controls of the Toffoli gates to be omitted, reducing their T-count from 7 to 4 and the T-count of addition to $8n + O(1)$.
+- Expanding a binary register into a unary register.
 
-Even if a Toffoli isn't paired with a second Toffoli that uncomputes its effects, it is still possible to perform the Toffoli with T-count of 4 by using an ancilla, a measurement, and a conditional fixup operation \cite{Jones2013} (see \autoref{fig:jones-toffoli}).
+The above list is not exhaustive.
+We leave further low hanging fruit as an exercise for the reader.
 
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1} &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw  &\qw       &\qw &\qw    &\ctrl{1}       &\qw \\
-      &\ctrl{1} &\qw & & = & & &     &\qw &\qw &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw  &\qw       &\qw &\qw    &\gate{Z}       &\qw \\
-      &\targ    &\qw & &   & & &     &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\qw  &\targ     &\qw &\qw    &\qw\cwx        &\qw \\
-      &         &    & &   & & & \qO &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qSi &\ctrl{-1} &\qH &\meter &\cw\cwx\bullet & \\
-    }
-  }
-  \caption{
-	Toffoli construction with T-count of 4 from \cite{Jones2013}.
-  }
-  \label{fig:jones-toffoli}
-\end{figure}
+\section{Adder Construction}
+\label{sec:construction}
 
-\section{Warm up: Reusing $|A\rangle$ states}
-\label{sec:invest}
-
-This paper started with an idea that improved the T-count of a single Toffoli gate from 7 to 6 in a surprising way, by performing the Toffoli indirectly instead of directly.
-We started with a clean ancilla in the $|0\rangle$ state, applied a Toffoli to store the logical-AND of the two controls in the ancilla, used the ancilla to control a CNOT onto the intended target, then uncomputed the ancilla.
-
-Initially, this indirect-Toffoli construction appears to have a T-count of 8.
-However, as shown in \autoref{fig:indirect-toffoli}, the last T gate in the circuit is unnecessary.
-In fact, it is actively harmful.
-Dropping this T gate and the following Hadamard not only reduces the T-count from 8 to 7, it leaves the ancilla in an $|A\rangle$ state that can be consumed to perform a T gate elsewhere.
-This improves the net T-count to 6.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1} &\qw && &&&    &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw       &\qw &\qw &\qw      &\qw  &\ctrl{3} &\qw &\qw      &\qw  &\qw &\qw &&\\
-      &\ctrl{1} &\qw &&=&&&    &\qw &\qw &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw       &\qw &\qw &\ctrl{2} &\qw  &\qw      &\qw &\ctrl{2} &\qw  &\qw &\qw &&\\
-      &\targ    &\qw && &&&    &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\targ     &\qw &\qw &\qw      &\qw  &\qw      &\qw &\qw      &\qw  &\qw &\qw &&\\
-      &         &    && &&&\qO &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\ctrl{-1} &\qH &\qT &\targ    &\qTi &\targ    &\qT &\targ    &\qTi &\qH &\qw 
-          \gategroup{4}{6}{4}{10}{.7em}{--} \gategroup{4}{26}{4}{30}{.7em}{--} &|0\rangle & \\
-      &         &    && &&&    &    |A\rangle &  & &     &         &    &         &     &    &          &    &    &         &    &         &     &         &     & |A\rangle
-    }
-  }
-  \caption{
-	Performing a Toffoli indirectly, by applying a pair of Toffolis to an ancilla and using the intermediate value to control a CNOT onto the actual target, appears to have a T-count of 8.
-	However, the initial T gate is being used to compute an $|A\rangle$ state and the final T gate is being used to uncompute the $|A\rangle$ state.
-    Instead of spending $|A\rangle$ states to perform T gates to make and unmake $|A\rangle$ states, we pass in and later recover a single $|A\rangle$ state.
-	This ``investment" reduces the net T-count to 6.
-  }
-  \label{fig:indirect-toffoli}
-\end{figure}
-
-A Toffoli with a T-count of 6 isn't optimal, but the reduction from 7 was done in an unexpected way.
-While searching for other circuits where this optimization might apply, we realized it would be useful when an initial Toffoli was later uncomputed by a second Toffoli.
-Instead of computing and uncomputing the ancilla for the first Toffoli, then recomputing and re-uncomputing the ancilla for the second Toffoli, we simply kept the ancilla around until it was time to uncompute the Toffoli.
-This halved the T-count of the pair from 12 to 6.
-See \autoref{fig:ancilla-temporary-toffoli}.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw && &&&    &\qw     &\qw  &\ctrl{3}&\qw &\qw     &\qw  &\qw &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\qw &\qw &\qw     &\qw  &\ctrl{3}&\qw &\qw     &\qw &&\\
-      &\ctrl{1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{1}&\qw &&=&&&    &\ctrl{2}&\qw  &\qw     &\qw &\ctrl{2}&\qw  &\qw &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\qw &\qw &\ctrl{2}&\qw  &\qw     &\qw &\ctrl{2}&\qw &&\\
-      &\targ   &\qw &    &\ldots &    &    &\targ   &\qw && &&&    &\qw     &\qw  &\qw     &\qw &\qw     &\qw  &\qw &\targ    &\qw &    &\ldots &    &    &\targ    &\qw &\qw &\qw     &\qw  &\qw     &\qw &\qw     &\qw &&\\
-      &        &    &    &       &    &    &        &    && &&&\qA &\targ   &\qTi &\targ   &\qT &\targ   &\qTi &\qH &\ctrl{-1}&\qw &\qw &\qw    &\qw &\qw &\ctrl{-1}&\qH &\qT &\targ   &\qTi &\targ   &\qT &\targ   &\qw &|A\rangle &\\
-    }
-  }
-  \caption{
-	Computing and uncomputing a Toffoli gate with a net T-count of 6.
-	Invests an $|A\rangle$ state when computing the Toffoli, holds an ancilla qubit storing the logical-AND of the controls until it is time to uncompute the Toffoli, then recovers the $|A\rangle$ state.
-  }
-  \label{fig:ancilla-temporary-toffoli}
-\end{figure}
-
-Many circuits involve computing and later uncomputing a Toffoli (e.g. addition).
-Previously, it was believed that each Toffoli had a T-count of 4 so the pair would have a T-count of 8.
-Our $|A\rangle$-investing construction state reduces the T-count of the pair from 8 to 6.
-
-We further improve the cost of paired Toffoli gates in the next section, obsoleting the $|A\rangle$-investing construction we just explained.
-We covered it regardless, because we believe that investing and recovering $|A\rangle$ states is a useful optimization likely to apply more generally to other kinds of circuits.
-
-\section{Constructing a temporary AND}
-\label{sec:temporary-and}
-
-In the previous section, we showed how to compute and uncompute a Toffoli gate by investing an $|A\rangle$ state and performing six $T$ gates.
-This construction be improved in several ways, ultimately producing our temporary AND gate construction.
-
-First, the computation step is introducing phase errors (the same ones as in \autoref{fig:bad-phase-toffoli}).
-In general, correcting these phase errors requires applying a controlled-S gate to the two controls, and pre-applying a controlled-Z gate involving the target.
-However, we know our target is $|0\rangle$ beforehand, so the pre-applied controlled-Z has no effect and can be ignored.
-Furthermore, the effect of the controlled-S gate is to apply a phase factor of $i$ to the amplitudes of computational basis states where both controls on.
-Our output happens to be a qubit storing whether or not both controls are on, so we can replace the controlled S gate on the controls with an uncontrolled S gate on the ancilla.
-This corrects all of the phase errors that were being introduced.
-
-Second, the construction from the previous section is unnecessarily serial.
-In the surface code, every T gate hides a measurement and an S gate performed conditional on the outcome of that measurement.
-It is beneficial to be waiting for several measurement results in parallel, instead of one after another serially.
-As shown in \cite{Jones2013}, the T gates of a Toffoli can all be performed at the same time, and we follow their lead.
-Note that we still pass an $|A\rangle$ state into the circuit instead of performing a fourth T gate, because passing in an $|A\rangle$ state already doesn't involve measuring and reacting.
-
-Third, taking another hint from \cite{Jones2013}, we can uncompute the ancilla by using a measure-and-correct process instead of with a mirror of the circuit that computed the ancilla.
-To explain the uncomputatoin process, we start with a circuit that obviously performs the uncomputation and then perform simple circuit moves until we arrive at the intended destination.
-Our starting point for the uncomputation circuit is simply a Toffoli gate, which obviously clears the ancilla since it was computed with a Toffoli and Toffolis are their own inverse.
-Then, since the cleared ancilla is going to be discarded anyways, we are free to apply a Hadamard gate and a measurement to it after the Toffoli but before discarding it.
-We then hop the Hadamard over the Toffoli, transforming it into a CCZ operation.
-Next we re-arrange the CCZ so that the ancilla is one of the controls, which is possible because the controls and targets of a CCZ are interchangeable.
-Finally, we invoke the deferred measurement principle \cite{Nielsen2009} to hop the Measurement over the CCZ, turning the quantum control into a classical control.
-Ultimately, the Toffoli gate we started with has been turned into a CZ gate that we apply conditional on the result of measuring the ancilla.
-The final circuit uses only Clifford operations, which have no T-count.
-
-With these three improvements applied, we are left with the key ingredients of our temporary AND gate.
-We can initialize an ancilla storing the logical-AND of two qubits.
-The creation process has a T-count of 4.
-Then we can keep the ancilla around, using it as a control for operations that would otherwise have been conditioned on both input qubits.
-Finally, once the ancilla is no longer needed, we can erase it at no T-count.
-
-In diagrams, we draw the computation of the AND gate (i.e.
-the ancilla being initialized) as an ancilla wire emerging vertically from two controls then heading rightward (see \autoref{fig:compute-logical-AND}).
-We draw the uncomputation analogously, with the ancilla wire coming in from the left then merging vertically into the two controls that created it (see \autoref{fig:erase-logical-AND}).
-The initialized lifetime of the ancilla qubit corresponds to the wire coming out of and later merging into the control qubits.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &x &&\ctrl{1} &\qw & x  && &&          &&\targ    &\ctrl{2}&\qw     &\qTi &\qw     &\ctrl{2}&\targ     &\qw &\qw &\qw &\\
-      &y &&\ctrl{1} &\qw & y  &&=&&          &&\targ    &\qw     &\ctrl{1}&\qTi &\ctrl{1}&\qw     &\targ     &\qw &\qw &\qw &\\
-      &  &&         &\qw & xy && &&|A\rangle &&\ctrl{-2}&\targ   &\targ   &\qT  &\targ   &\targ   &\ctrl{-2} &\qH &\qS &\qw &\\
-    }
-  }
-  \caption{
-	Computing the logical-AND of two qubits, with a T-count of 4 and a T-depth of 1.
-  }
-  \label{fig:compute-logical-AND}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &x  &&\ctrl{1} &\qw &x && &&\qw &\qw    &\ctrl{1} &\qw \\
-      &y  &&\ctrl{1} &\qw &y &&=&&\qw &\qw    &\gate{Z} &\qw \\
-      &xy &&\qw      &    &  && &&\qH &\meter &\cw \cwx \bullet &    \\
-    }
-  }
-  \caption{
-	Uncomputing the logical-AND of two qubits, with a T-count of 0.
-  }
-  \label{fig:erase-logical-AND}
-\end{figure}
-
-We can perform a single Toffoli by computing the logical-AND of two qubits, applying a CNOT from the logical-AND onto the actual target, then uncomputing the logical-AND.
-This combined construction, shown in \autoref{fig:merge-use-erase-toffoli}, is equivalent to the one from \cite{Jones2013}.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\targ    &\ctrl{2}&\qw     &\qTi &\qw     &\ctrl{2}&\targ     &\qw &\qw &\qw      &\qw &\qw    &\ctrl{1}         &\qw \\
-      &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\ctrl{1} &\qw && &&          &&\targ    &\qw     &\ctrl{1}&\qTi &\ctrl{1}&\qw     &\targ     &\qw &\qw &\qw      &\qw &\qw    &\gate{Z}         &\qw \\
-      &         &    &&=&&         &\ctrl{1} &\qw      &    &&=&&|A\rangle &&\ctrl{-2}&\targ   &\targ   &\qT  &\targ   &\targ   &\ctrl{-2} &\qH &\qS &\ctrl{1} &\qH &\meter &\cw \cwx \bullet &    \\
-      &\targ    &\qw && &&\qw      &\targ    &\qw      &\qw && &&          &&\qw      &\qw     &\qw     &\qw  &\qw     &\qw     &\qw       &\qw &\qw &\targ    &\qw &\qw    &\qw              &\qw \\
-    }
-  }
-  \caption{
-	Performing a Toffoli gate by computing the logical-AND of its controls, using the result, then erasing it.
-	Has a T-count of 4 and a measure-react depth of 2.
-	Equivalent to the construction from \cite{Jones2013} shown in \autoref{fig:jones-toffoli}.
-  }
-  \label{fig:merge-use-erase-toffoli}
-\end{figure}
-
-In hindsight, our improvement on the construction in \cite{Jones2013} is merely noticing that it can be split it useful pieces that don't all need to be applied at the same time.
-In particular, when a Toffoli is later uncomputed by a second Toffoli, simply delaying erasing the ancilla wire as shown in \autoref{fig:paired-toffoli-to-logical-AND} reduces the T-count of such paired Toffolis from 8 to 4.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
-      &\ctrl{2} &\qw &\qw &\qw    &\qw &\qw &\ctrl{2} &\qw && &&\ctrl{1} &\qw      &\qw &\qw &\qw    &\qw &\qw &\qw      &\ctrl{1} &\qw & \\
-      &         &    &    &       &    &    &         &    &&=&&         &\ctrl{1} &\qw &\qw &\qw    &\qw &\qw &\ctrl{1} &\qw      &    & \\
-      &\targ    &\qw &    &\ldots &    &    &\targ    &\qw && &&\qw      &\targ    &\qw &    &\ldots &    &    &\targ    &\qw      &\qw & \\
-    }
-  }
-  \caption{
-	Replacing a pair of compute/uncompute Toffolis with a temporary AND gate and Clifford operations.
-	Improves the T-count from 8 to 4.
-  }
-  \label{fig:paired-toffoli-to-logical-AND}
-\end{figure}
-
-\section{Applications}
-\label{sec:circuit-constructions}
-
-In this section we will show how to apply temporary Toffolis to improve the complexity of several circuits.
-
-Existing adder constructions such as the Cuccaro adder \cite{Cuccaro2004} contain many Toffolis that are later uncomputed by another Toffoli.
-Our improved construction does not fundamentally change the structure of the adder.
-We simply perform the temporary Toffoli gates with our temporary AND gate construction, halving the cost.
-\autoref{fig:full-adder-block} shows the per-bit building block of our adder construction.
-\autoref{fig:multi-bit-adder-example} shows how to tile the building block into a five bit adder.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &c_k &&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\ctrl{1}&\qw     &\qw &c_k     &&&&& \\
-      &i_k &&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\targ   &\ctrl{1}&\qw &i_k     &&&&& \\
-      &t_k &&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\qw     &\targ   &\qw &&&(t+i)_k &&& \\
-      &    &&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &    &&&        &&& \\
-    }
-  }
-  \caption{
-	Building block, for an adder, with a T-count of 4.
-	Equivalent to the building block of the Cuccaro adder \cite{Cuccaro2004}, but using our temporary AND gate construction.
-  }
-  \label{fig:full-adder-block}
-\end{figure}
+In \autoref{fig:multi-bit-adder-example}, we construct a 5-bit adder with a T-count of 16.
+It performs 4 temporary logical-ANDs, each with a T-count of 4.
 
@@
 	A 5-bit adder with T-count of 16.
-	The circuit contains Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures with no T-count.
+	Uses Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures requiring no T gates.
 	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
@@
 
-With a slight tweak to the adder building block, we can make the sum available for use as soon as the carry signal hits the high bit (instead of needing to wait for the uncomputation sweep to finish).
-This can halve the T-count of the addition when it is going to be uncomputed.
-Instead of paying $4n + O(1)$ T gates to compute the addition, and then $4n + O(1)$ more T gates to uncompute the addition, we simply work with the intermediate state of a single addition computation.
-\autoref{fig:temporary-full-adder-block} shows the tweaked building block, which is also a temporary three-bit adder with T-count 4.
+We construct $n$-bit adders by nesting $n$ copies of the building block shown in \autoref{fig:full-adder-block} inside of each other.
+We then specialize the outer-most and inner-most blocks (which act on the low bit and high bit respectively) based on the fact that they either have no carry input or no carry output.
 
@@
     \Qcircuit @R=0.7em @C=0.7em {
-      &a &&\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw &a          &&&&&&& \\
-      &b &&\targ   &\ctrl{1}&\qw     &\ctrl{1}&\qw &&b \oplus a &&&&&& \\
-      &c &&\targ   &\ctrl{1}&\qw     &\targ   &\qw &&&&(a+b+c)_0  &&&& \\
-      &  &&        &        &\targ   &\qw     &\qw &&&&(a+b+c)_1  &&&& \\
+      &c_k &&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\ctrl{1}&\qw     &\qw &c_k     &&&&& \\
+      &i_k &&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\targ   &\ctrl{1}&\qw &i_k     &&&&& \\
+      &t_k &&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\qw     &\targ   &\qw &&&(t+i)_k &&& \\
+      &    &&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &    &&&        &&& \\
     }
@@
   \caption{
-	  A three-bit adder with T-count 4 that can later be uncomputed for no T-count.
-	  A building-block for temporary addition circuits whose result will be used but then uncomputed.
+	Adder circuit building-block with a marginal T-count of 4.
+  }
+  \label{fig:full-adder-block}
+\end{figure}
+
+Our adder circuits use temporary logical-AND operations.
+We define the temporary-AND computation (which we draw as a wire emerging from two controls) in \autoref{fig:compute-logical-AND}.
+We define the corresponding uncomputation (which we draw as a wire merging into two controls) in \autoref{fig:erase-logical-AND}.
+Computing the temporary-AND has a T-count of 4, but when uncomputing a temporary-AND we have access to measurement only need Clifford operations.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &x &&\ctrl{1} &\qw & x  && &&          &&\targ    &\ctrl{2}&\qw     &\qTi &\qw     &\ctrl{2}&\targ     &\qw &\qw &\qw &\\
+      &y &&\ctrl{1} &\qw & y  &&=&&          &&\targ    &\qw     &\ctrl{1}&\qTi &\ctrl{1}&\qw     &\targ     &\qw &\qw &\qw &\\
+      &  &&         &\qw & xy && &&|A\rangle &&\ctrl{-2}&\targ   &\targ   &\qT  &\targ   &\targ   &\ctrl{-2} &\qH &\qS &\qw &\\
+    }
+  }
+  \caption{
+	Computing the logical-AND of two qubits, with a T-count of 4 and a T-depth of 1.
+	The $|A\rangle$ state input on the bottom wire contributes to the T-count (because $|A\rangle$ states are the resource used to perform T gates).
+  }
+  \label{fig:compute-logical-AND}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &x  &&\ctrl{1} &\qw &x && &&\qw &\qw    &\ctrl{1} &\qw \\
+      &y  &&\ctrl{1} &\qw &y &&=&&\qw &\qw    &\gate{Z} &\qw \\
+      &xy &&\qw      &    &  && &&\qH &\meter &\cw \cwx \bullet &    \\
+    }
+  }
+  \caption{
+	Uncomputing the logical-AND of two qubits using only Clifford operations.
+    Has a T-count of 0.
+  }
+  \label{fig:erase-logical-AND}
+\end{figure}
+
+\section{Other Applications}
+\label{sec:other-applications}
+
+With a slight tweak to the adder building block, we can make the sum available half-way through the adder circuit, before any uncomputation of temporary-ANDs occurs.
+When a circuit is going to compute an addition, then later uncompute it, using the sum at the half-way point can halve the T-count.
+Instead of paying $4n + O(1)$ T gates to compute the addition, and then $4n + O(1)$ more T gates to uncompute the addition, we simply work with the intermediate state of a single addition computation then backtrack instead of finishing.
+\autoref{fig:temporary-full-adder-block} shows the temporary adder building block, which is also a temporary three-bit adder with T-count 4.
+
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &a &&\ctrl{2}&\qw     &\qw     &\ctrl{3}&\qw &a          &&&&&&& \\
+      &b &&\targ   &\ctrl{1}&\ctrl{1}&\targ   &\qw &b          &&&&&&& \\
+      &c &&\targ   &\ctrl{1}&\targ   &\qw     &\qw &&&&(a+b+c)_0  &&&& \\
+      &  &&        &        &\qw     &\targ   &\qw &&&&(a+b+c)_1  &&&& \\
+    }
+  }
+  \caption{
+	Temporary-adder circuit building-block with a marginal T-count of 4.
+	Uncomputing the result requires no T gates.
+	Can be used to temporarily add two $n$-bit registers (T-count $4n + O(1)$).
+	Can be used to temporarily compute the pop count of an $n$-bit register (T-count $4n + O(\lg n)$).
   }
@@
 
-Some additions, such as the ones performed in Shor's algorithm, are conditioned on a control qubit.
+Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
 \cite{Coreas2017} derives a controlled-addition construction with T-count of $21n + O(1)$.
-We use our temporary AND gate construction to improve this to $8n + O(1)$, and show the building block in \autoref{fig:controlled-full-adder-block}.
+We use temporary logical-ANDs to construct an improved circuit with a T-count of $8n + O(1)$ (see \autoref{fig:controlled-full-adder-block} for the building-block).
 
@@
   \caption{
-	Building block, for a controlled adder, with a T-count of 8.
+	Controlled-adder circuit building-block with a marginal T-count of 8.
   }
@@
 
-The temporary AND gate is useful for more than just arithmetic.
-For example, temporary ANDs can be useful for applying phase rotations to many qubits simultaneously.
-Given a $b$-bit ancilla register $G$ prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle$ (a `phase gradient state'), using our adder construction to add a register $Q$ into $G$ will cause phase kickback that applies the operation $\text{Grad} = \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle \langle k|$ to $V$.
-The $\text{Grad}$ operation is equivalent to applying the phase gate $Z^{2^-k}$ to the qubit position at $k$ for each qubit within $Q$.
-Some quantum Fourier transform circuits involve conditional uses of $\text{Grad}$, so it is possible that temporary-ANDs improve the T-count of those circuits.
-
-Our constructions reduce many existing estimates of the cost of quantum computation.
-For example, \cite{Fowler2012} estimates that factoring a 2000-bit number would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$.
-Their time estimate is based on each Toffoli having a T-depth of 1, and their $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
-Because Shor's algorithm is dominated by the cost of additions, our techniques (and previous work) multiply the T-count by $\frac{4}{14}$ and the T-depth by $\frac{1}{3}$.
+Decreasing the T-count of addition reduces the T-count of any construction based on addition.
+For example, \cite{Fowler2012} estimate that factoring a 2048-bit number would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$.
+Their time estimate is based on Toffolis having a T-depth of 3, and their $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
+Because Shor's algorithm is dominated by the cost of additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $2$ and $1$ respectively.
 This reduces the estimate to 9 hours and $6 \cdot 10^{11}$ distilled $|A\rangle$ states.
 
-Other examples of operations which benefit from cheaper temporary AND gates include:
+On top of reducing the T-count of obviously related operations like multiplication, reducing the T-count of addition also reduces the T-count quantum-specific operations such as rotating qubits.
+For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a T-cost $4n + O(\text{poly}(\frac{1}{\epsilon}) \lg n)$.
+First, use our temporary adder iteratively to compute the pop count of a register containing the target qubits.
+This takes $4n + O(\lg n)$ T gates to do.
+Then, for each position $k$ in the pop count output register, synthesize and apply the operation $R_Z(\theta \cdot 2^k)$ to the qubit at that position.
+This uses a number of T gates proportional to some polynomial of the desired precision times the number of bits in the output register.
+But the output register is exponentially smaller than $n$, so this cost is asymptotically negligible.
+Finally, uncompute the temporary pop count (using no T gates).
+This applies the desired operation.
 
-- Integer comparisons.
+Another rotation-based operation that can be implemented via an adder is the ``$n$-bit phase gradient operation" $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
+Normally this operation would be implemented by separately applying the operation $Rz(2 \pi 2^{-p})$ to each qubit, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depended on the maximum per-gate error $\epsilon$.
+However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition using only $4n + O(1)$ T gates.
+Just add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
+Ignoring the one-time cost of initializing the reusable phase gradient register, this uses only $4n + O(1)$ T gates.
 
-- Integer multiplication.
+Some quantum Fourier transform circuits use controlled phase gradient operations.
+Controlled phase gradients can be applied by adding into the reusable phase gradient register with controlled adders instead of uncontrolled adders.
+Assume a maximum per-gate error of $\epsilon$, start from the textbook QFT circuit based on the Cooley-Tukey algorithm from \cite{Nielsen2009}, and drop negligible rotations, merge individual controlled rotations into controlled phase gradients, and apply those controlled phase gradients using controlled adders.
+The result is a simple approximate $n$-qubit QFT circuit that uses fewer than $8 n \lg \lg \frac{1}{\epsilon}$ T gates.
 
-- Incrementing and counting.
+The temporary logical-AND is useful for optimizing an even wider variety of circuits than addition is.
+Whenever Toffoli gates appear in compute/uncompute pairs, and intermediate operations are not sensitive to phase errors on the target or controls of the Toffoli gate (see \autoref{fig:pair-definition}), it is possible to save 4 T gates by replacing the pair of Toffoli gates with a temporary logical-AND.
 
-- Integer arithmetic in general.
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw && &&&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &            &\\
+      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw && &&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &            &\\
+      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&=&&&          &\ctrl{2}   &\ghost{U}       &\ctrl{2}&\qw &            &\\
+      &          &\targ      &\ghost{U}       &\targ   &\qw && &&&          &\targ      &\ghost{U}       &\targ   &\qw &            &\\
+      &          &           &                &        &    && &&&\qO       &\targ      &\qw             &\targ   &\qw &| 0 \rangle &\\
+    }
+  }
+  \caption{
+	A sufficient condition for replacing a pair of Toffoli gates with a temporary logical-AND, saving 4 T gates.\\
+	1) The later Toffoli gate must be uncomputing the earlier Toffoli gate (indicated by the ancilla on the RHS starting and ending up in the $|0\rangle$ state).\\
+	2) Intermediate operations must not be sensitive to the presence of entangled copies of the target.
+  }
+  \label{fig:pair-definition}
+\end{figure}
 
-- Modular arithmetic.
+For example, the Grover diffusion operator's cost is dominated by the cost of merging $n$ controls.
+Merging the controls can be computed, and then uncomputed, with Toffoli gates.
+By instead iteratively merging the controls with nested temporary-ANDs, the Grover diffusion operation can be applied using $4n + O(1)$ T gates.
 
-- Expanding a binary register into a unary register.
-
-- Operations with a target qubit indexed by a binary qubit register.
-
-- Phasing a register by a computable function $f$ (i.e. applying the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$).
-
-- Temporary permutations.
-
-- Oracles in Grover's algorithm.
-
-We leave further low hanging fruit as exercises for the reader.
+For other cases where temporary-ANDs can be used to optimize circuits, we refer back to the introduction.
+Even circuits based on simulating physical systems, not just abstract mathematical problems, can benefit \cite{RyanEmails2017}.
 
@@
 
-For over a decade, the T-count of addition has been $8n + O(1)$ \cite{Barenco1995, Cuccaro2004, Amy2013}).
+For over a decade, the T-count of addition has been known to be at most $8n + O(1)$ \cite{Amy2013, Barenco1995, Cuccaro2004}).
 This was believed to be optimal \cite{AustinDiscussionsAndEmails2017}.
 In this paper, we halved the T-count of addition to $4n + O(1)$.
-We did so with a construction that applies to many other circuits: the temporary AND gate.
+We did so with a construction that applies to many other circuits: the temporary logical-AND gate.
 
-Our result raises the question of whether further T-count improvements to addition or to temporary AND gates are possible.
-A simple method for proving lower bounds on the T-count of an operation is to synthesize T gates by applying that operation.
-In \autoref{fig:lower-bound-temporary-and} we use a T gate and a single temporary-AND to synthesize three T gates.
-This proves that the T-count of a temporary-AND is at least two.
+We note that the T-count of the temporary logical-AND is at least 2 in \autoref{fig:lower-bound-temporary-and}.
 Two is less than four, but we are not aware of any method for further reducing the T-count of temporary-ANDs or for synthesizing more T gates out of a temporary-AND.
+We are unsure if addition or temporary logical-AND gates can be done with fewer, but certain that T-count improvements as significant as the ones we have made in this paper are possible for other basic operations used by many circuits.
 
@@
 
-We thank Austin Fowler for assistance in locating relevant references, sampling the opinions of researchers, and for comments that greatly improved the paper.
+We thank Austin Fowler for assistance in locating relevant references, sampling the opinions of other researchers, and for comments that greatly improved our paper.
 

### 2017-08-29 paper.tex
Polish - Moving adder figure to first page - Merged AND compute/uncompute figures - Annealing the text

@@
 The surface code is a quantum error correcting code that works on a 2D nearest-neighbour array of qubits and achieves a threshold error rate of approximately 1\% [[[[cite so much]]]].
-This makes the surface code a likely component in the architecture of future error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies [[[[cite so much]]]] and other known error correcting codes have lower thresholds or require stronger connectivity [[[[cite so much]]]].
+This makes the surface code a likely component in the architecture of future error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies [[[[cite so much]]]] and other well understood error correcting codes either have lower thresholds or require stronger connectivity [[[[cite so much]]]].
 
@@
 Because T gates are so expensive for the surface code, and the surface code is a likely component of future quantum computers, it is important to consider and optimize the number of T gates used by quantum circuits.
+Optimizing the T-count of basic elements of quantum circuits, such as the construction of adders and Toffoli gates, is even more important.
 
@@
 when an initial Toffoli operation is later uncomputed by a second Toffoli operation, each Toffoli in the pair can omit three of the T gates from the textbook construction.
-This introduces phase errors, assuming this doesn't affect intermediate operatoins, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \cite{Barenco1995}.
-It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and classically conditioned fixup operations \cite{Jones2013}.
+This introduces phase errors but, assuming intermediate operations aren't poisoned by the phase errors, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \cite{Barenco1995, Nielsen2009}.
+It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and a classically conditioned fixup operations \cite{Jones2013}.
 
@@
 Correspondingly, existing adder constructions have T-counts of $8n-4$ \cite{Amy2013}.
-Adders with a T-count of $8n-4$ have been known for over a decade \cite{Barenco1995, Cuccaro2004}, and were believed to be optimal \cite{AustinDiscussionsAndEmails2017}.
+Adders with a T-count of $8n + O(1)$ have been known for over a decade \cite{Barenco1995, Cuccaro2004}.
+The leading factor of 8 has been conjectured to be optimal \cite{AustinDiscussionsAndEmails2017}.
+We show that this is not true by reducing the T-count of addition to $4n + O(1)$.
 
-Our paper is divided into four sections.
-In \autoref{sec:introduction}, we motivate the problem, discuss existing work, and note the many applications of our construction.
-\autoref{sec:construction} defines our $4n + O(1)$ adder and explains its pieces.
-\autoref{sec:other-applications} covers circuits for other adding tasks and discusses other kinds of circuits where the temporary-AND construction can improve T-counts.
-Finally, \autoref{sec:conclusion} summarizes our contributions and proves that at least two T gates are needed to perform a temporary AND while discussing future work.
+\begin{figure}
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &i_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\qw &i_0     &&&&&\\
+      &t_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw &&&(t+i)_0 &&&\\
+      &    &&         &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\ctrl{1}&\qw     &        &    &&&        &&&\\
+      &i_1 &&\qw      &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\ctrl{1}&\qw &i_1     &&&&&\\
+      &t_1 &&\qw      &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\targ   &\qw &&&(t+i)_1 &&&\\
+      &    &&         &        &        &\targ   &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\ctrl{1}&\targ   &\qw     &        &        &        &    &&&        &&&\\
+      &i_2 &&\qw      &\qw     &\qw     &\qw     &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_2     &&&&&\\
+      &t_2 &&\qw      &\qw     &\qw     &\qw     &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_2 &&&\\
+      &    &&         &        &        &        &        &        &\targ   &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\ctrl{3}&\qw     &\ctrl{1}&\targ   &\qw     &        &        &        &        &        &        &    &&&        &&&\\
+      &i_3 &&\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_3     &&&&&\\
+      &t_3 &&\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_3 &&&\\
+      &    &&         &        &        &        &        &        &        &        &        &\targ   &\ctrl{2}&\targ   &\qw     &        &        &        &        &        &        &        &        &        &    &&&        &&&\\
+      &i_4 &&\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_4     &&&&&\\
+      &t_4 &&\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_4 &&&\\
+    }
+  }
+  \caption{
+	A 5-bit adder with T-count of 16.
+	Uses Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures requiring no T gates.
+	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
+	See \autoref{fig:full-adder-block} for the adder building-block and \autoref{fig:temporary-logical-AND} for the logical-AND computation and uncomputation circuits.
+  }
+  \label{fig:multi-bit-adder-example}
+\end{figure}
 
-We improve the number of T gates needed to perform an adder by halving the number of T gates needed to perform paired Toffoli gates.
-In this paper we focus on how this halves the T-count of addition, but improving the T-count of paired Toffolis is a widely applicable optimization.
+We halve the number of T gates needed to perform an adder by halving the number of T gates needed to perform Toffoli gates that appear in compute/uncompute pairs.
+In this paper we focus on how this optimization halves the T-count of addition, but improving the T-count of paired Toffolis is a widely applicable optimization.
 Other examples of operations whose T-count can be improved by our constructions include:
@@
 
+Our paper is divided into four sections.
+In \autoref{sec:introduction}, we motivate the need to optimize T-counts, discuss existing work, and note possible applications of our low T-count constructions.
+\autoref{sec:construction} presents our $4n + O(1)$ T-count adder and explains its pieces, including the temporary logical-AND.
+\autoref{sec:other-applications} discusses several circuit constructions and estimates improved by our adder and temporary logical-AND constructions.
+Finally, \autoref{sec:conclusion} summarizes our contributions and discusses future work.
+
 \section{Adder Construction}
@@
 
-In \autoref{fig:multi-bit-adder-example}, we construct a 5-bit adder with a T-count of 16.
+In \autoref{fig:multi-bit-adder-example}, we present a 5-bit adder with a T-count of 16.
 It performs 4 temporary logical-ANDs, each with a T-count of 4.
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &i_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\qw &i_0     &&&&&\\
-      &t_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw &&&(t+i)_0 &&&\\
-      &    &&         &\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\ctrl{1}&\qw     &        &    &&&        &&&\\
-      &i_1 &&\qw      &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\ctrl{1}&\qw &i_1     &&&&&\\
-      &t_1 &&\qw      &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\targ   &\qw &&&(t+i)_1 &&&\\
-      &    &&         &        &        &\targ   &\qw &\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\ctrl{1}&\targ   &\qw     &        &        &        &    &&&        &&&\\
-      &i_2 &&\qw      &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_2     &&&&&\\
-      &t_2 &&\qw      &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_2 &&&\\
-      &    &&         &        &        &        &    &        &        &\targ   &\qw &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\ctrl{3}&\qw     &\ctrl{1}&\targ   &\qw     &        &        &        &        &        &        &    &&&        &&&\\
-      &i_3 &&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_3     &&&&&\\
-      &t_3 &&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_3 &&&\\
-      &    &&         &        &        &        &    &        &        &        &    &        &        &\targ   &\ctrl{2}&\targ   &\qw     &        &        &        &        &        &        &        &        &        &    &&&        &&&\\
-      &i_4 &&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_4     &&&&&\\
-      &t_4 &&\qw      &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\qw &\qw     &\qw     &\qw     &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_4 &&&\\
-    }
-  }
-  \caption{
-	A 5-bit adder with T-count of 16.
-	Uses Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures requiring no T gates.
-	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
-  }
-  \label{fig:multi-bit-adder-example}
-\end{figure}
 
@@
   \caption{
-	Adder circuit building-block with a marginal T-count of 4.
+	Adder circuit building-block with a T-count of 4.
   }
@@
 Our adder circuits use temporary logical-AND operations.
-We define the temporary-AND computation (which we draw as a wire emerging from two controls) in \autoref{fig:compute-logical-AND}.
-We define the corresponding uncomputation (which we draw as a wire merging into two controls) in \autoref{fig:erase-logical-AND}.
-Computing the temporary-AND has a T-count of 4, but when uncomputing a temporary-AND we have access to measurement only need Clifford operations.
+We define the temporary logical-AND computation (which we draw as a wire emerging from two controls), and the corresponding uncomputation (which we draw as a wire merging into two controls), in \autoref{fig:compute-logical-AND}.
+Computing the temporary logical-AND has a T-count of 4, but when uncomputing a temporary logical-AND we have access to measurement and only need Clifford operations.
 
@@
     \Qcircuit @R=0.7em @C=0.7em {
-      &x &&\ctrl{1} &\qw & x  && &&          &&\targ    &\ctrl{2}&\qw     &\qTi &\qw     &\ctrl{2}&\targ     &\qw &\qw &\qw &\\
-      &y &&\ctrl{1} &\qw & y  &&=&&          &&\targ    &\qw     &\ctrl{1}&\qTi &\ctrl{1}&\qw     &\targ     &\qw &\qw &\qw &\\
-      &  &&         &\qw & xy && &&|A\rangle &&\ctrl{-2}&\targ   &\targ   &\qT  &\targ   &\targ   &\ctrl{-2} &\qH &\qS &\qw &\\
+      &x &&\ctrl{1} &\qw &x  && &&          &&\ctrl{2}&\qw     &\targ    &\qTi &\targ     &\qw &\qw &\qw &\\
+      &y &&\ctrl{1} &\qw &y  &&=&&          &&\qw     &\ctrl{1}&\targ    &\qTi &\targ     &\qw &\qw &\qw &\\
+      &  &&         &\qw &xy && &&|A\rangle &&\targ   &\targ   &\ctrl{-2}&\qT  &\ctrl{-2} &\qH &\qS &\qw &\\
+    }
+  }
+  \\ \vspace*{0.25cm} \\
+  \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &x  &&\ctrl{1} &\qw &x && &&&\qw &\qw    &\ctrl{1}         &\qw &&&&&&&&&&\\
+      &y  &&\ctrl{1} &\qw &y &&=&&&\qw &\qw    &\gate{Z}         &\qw &&&&&&&&&&\\
+      &xy &&\qw      &    &  && &&&\qH &\meter &\cw \cwx \bullet &    &&&&&&&&&&\\
     }
@@
   \caption{
-	Computing the logical-AND of two qubits, with a T-count of 4 and a T-depth of 1.
-	The $|A\rangle$ state input on the bottom wire contributes to the T-count (because $|A\rangle$ states are the resource used to perform T gates).
+	How to compute and uncompute the logical-AND of two qubits.
+	The computation circuit has a T-count of 4 and a T-depth of 1.
+	Note that the $|A\rangle$ state input contributes to the T-count, because $|A\rangle$ states are the resource used to perform T gates.
+	The uncomputation circuit only uses Clifford gates, and so has a T-count of zero.
+    \\
+    An alternative uncomputation construction is to simply do the reverse of the computation circuit.
+    This alternative approach has a net T-count of 2 (because an $|A\rangle$ state is recovered).
+    That would still be an improvement on existing work, but inferior to the measure-and-fixup construction shown above.
   }
-  \label{fig:compute-logical-AND}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &x  &&\ctrl{1} &\qw &x && &&\qw &\qw    &\ctrl{1} &\qw \\
-      &y  &&\ctrl{1} &\qw &y &&=&&\qw &\qw    &\gate{Z} &\qw \\
-      &xy &&\qw      &    &  && &&\qH &\meter &\cw \cwx \bullet &    \\
-    }
-  }
-  \caption{
-	Uncomputing the logical-AND of two qubits using only Clifford operations.
-    Has a T-count of 0.
-  }
-  \label{fig:erase-logical-AND}
+  \label{fig:temporary-logical-AND}
 \end{figure}
@@
 
-With a slight tweak to the adder building block, we can make the sum available half-way through the adder circuit, before any uncomputation of temporary-ANDs occurs.
+With a slight tweak to the adder building block, we can make the sum available half-way through the adder circuit, before any uncomputation of temporary logical-ANDs occurs.
 When a circuit is going to compute an addition, then later uncompute it, using the sum at the half-way point can halve the T-count.
@@
   \caption{
-	Temporary-adder circuit building-block with a marginal T-count of 4.
+	Temporary-adder circuit building-block with a T-count of 4.
 	Uncomputing the result requires no T gates.
@@
   \caption{
-	Controlled-adder circuit building-block with a marginal T-count of 8.
+	Controlled-adder circuit building-block with a T-count of 8.
   }
@@
 Decreasing the T-count of addition reduces the T-count of any construction based on addition.
-For example, \cite{Fowler2012} estimate that factoring a 2048-bit number would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$.
-Their time estimate is based on Toffolis having a T-depth of 3, and their $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
+For example, in \cite{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$.
+The time estimate is based on Toffolis having a T-depth of 3, and the $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
 Because Shor's algorithm is dominated by the cost of additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $2$ and $1$ respectively.
-This reduces the estimate to 9 hours and $6 \cdot 10^{11}$ distilled $|A\rangle$ states.
+This reduces the estimates to 9 hours and $6 \cdot 10^{11}$ distilled $|A\rangle$ states respectively.
 
-On top of reducing the T-count of obviously related operations like multiplication, reducing the T-count of addition also reduces the T-count quantum-specific operations such as rotating qubits.
-For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a T-cost $4n + O(\text{poly}(\frac{1}{\epsilon}) \lg n)$.
-First, use our temporary adder iteratively to compute the pop count of a register containing the target qubits.
+On top of reducing the T-count of obviously related operations like multiplication and exponentiation, reducing the T-count of addition also reduces the T-count of quantum-specific operations such as rotating qubits.
+For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a total T-cost $4n + O(\text{poly}(\frac{1}{\epsilon}) \lg n)$.
+First, iteratively apply our temporary adder to compute the pop count of a register containing the target qubits.
 This takes $4n + O(\lg n)$ T gates to do.
@@
 This uses a number of T gates proportional to some polynomial of the desired precision times the number of bits in the output register.
-But the output register is exponentially smaller than $n$, so this cost is asymptotically negligible.
+But the output register is exponentially smaller than $n$, so this cost is negligible for large $n$.
 Finally, uncompute the temporary pop count (using no T gates).
-This applies the desired operation.
+This applies the desired $R_Z(\theta)$ operations.
 
-Another rotation-based operation that can be implemented via an adder is the ``$n$-bit phase gradient operation" $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
-Normally this operation would be implemented by separately applying the operation $Rz(2 \pi 2^{-p})$ to each qubit, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depended on the maximum per-gate error $\epsilon$.
+Another rotation-based operation that can be implemented via an adder is the $n$-bit phase gradient operation $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
+Normally this operation would be implemented by separately applying the operation $R_z(\pi 2^{-p})$ to each qubit, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
 However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition using only $4n + O(1)$ T gates.
@@
 Controlled phase gradients can be applied by adding into the reusable phase gradient register with controlled adders instead of uncontrolled adders.
-Assume a maximum per-gate error of $\epsilon$, start from the textbook QFT circuit based on the Cooley-Tukey algorithm from \cite{Nielsen2009}, and drop negligible rotations, merge individual controlled rotations into controlled phase gradients, and apply those controlled phase gradients using controlled adders.
+Assume a maximum per-gate error of $\epsilon$, start from the textbook QFT circuit based on the Cooley-Tukey algorithm from \cite{Nielsen2009}, drop negligible rotations, merge individual controlled rotations into controlled phase gradients, and apply those controlled phase gradients using controlled adders.
 The result is a simple approximate $n$-qubit QFT circuit that uses fewer than $8 n \lg \lg \frac{1}{\epsilon}$ T gates.
@@
 The temporary logical-AND is useful for optimizing an even wider variety of circuits than addition is.
-Whenever Toffoli gates appear in compute/uncompute pairs, and intermediate operations are not sensitive to phase errors on the target or controls of the Toffoli gate (see \autoref{fig:pair-definition}), it is possible to save 4 T gates by replacing the pair of Toffoli gates with a temporary logical-AND.
+Whenever Toffoli gates appear in compute/uncompute pairs, and intermediate operations are not sensitive to phase errors on the controls of the Toffoli gate (i.e.
+the condition in \autoref{fig:pair-definition} is satisfied), it is possible to save 4 T gates by replacing the pair of Toffoli gates with a temporary logical-AND.
 
@@
     \Qcircuit @R=0.7em @C=0.7em {
-      &\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw && &&&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &            &\\
-      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw && &&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &            &\\
-      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&=&&&          &\ctrl{2}   &\ghost{U}       &\ctrl{2}&\qw &            &\\
-      &          &\targ      &\ghost{U}       &\targ   &\qw && &&&          &\targ      &\ghost{U}       &\targ   &\qw &            &\\
-      &          &           &                &        &    && &&&\qO       &\targ      &\qw             &\targ   &\qw &| 0 \rangle &\\
+      &\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &&               &&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &\\
+      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&               &&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &\\
+      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&\stackrel{?}{=}&&          &\ctrl{2}   &\ghost{U}       &\ctrl{2}&\qw &\\
+      &          &\targ      &\ghost{U}       &\targ   &\qw &&               &&          &\targ      &\ghost{U}       &\targ   &\qw &\\
+      &          &\qw        &\qw             &\qw     &\qw &&               &&          &\targ      &\qw             &\targ   &\qw &\\
     }
@@
 	A sufficient condition for replacing a pair of Toffoli gates with a temporary logical-AND, saving 4 T gates.\\
-	1) The later Toffoli gate must be uncomputing the earlier Toffoli gate (indicated by the ancilla on the RHS starting and ending up in the $|0\rangle$ state).\\
-	2) Intermediate operations must not be sensitive to the presence of entangled copies of the target.
+	1) The later Toffoli gate must be uncomputing the earlier Toffoli gate.\\
+	2) Intermediate operations must not be sensitive to the presence of the entangled ancilla.
   }
@@
 Merging the controls can be computed, and then uncomputed, with Toffoli gates.
-By instead iteratively merging the controls with nested temporary-ANDs, the Grover diffusion operation can be applied using $4n + O(1)$ T gates.
+By instead iteratively merging the controls with nested temporary logical-ANDs, the Grover diffusion operation can be applied using $4n + O(1)$ T gates.
 
-For other cases where temporary-ANDs can be used to optimize circuits, we refer back to the introduction.
+For other cases where temporary logical-ANDs can be used to optimize circuits, we refer back to the introduction.
 Even circuits based on simulating physical systems, not just abstract mathematical problems, can benefit \cite{RyanEmails2017}.
@@
 
-For over a decade, the T-count of addition has been known to be at most $8n + O(1)$ \cite{Amy2013, Barenco1995, Cuccaro2004}).
-This was believed to be optimal \cite{AustinDiscussionsAndEmails2017}.
-In this paper, we halved the T-count of addition to $4n + O(1)$.
-We did so with a construction that applies to many other circuits: the temporary logical-AND gate.
+For over a decade, the T-count of addition has been known to be at most $8n - 4$ \cite{Amy2013, Barenco1995, Cuccaro2004}.
+This was conjectured to be optimal \cite{AustinDiscussionsAndEmails2017}.
+We halved the leading factor of this cost, and we did so via a construction that applies to many other circuits: the temporary logical-AND gate.
+Our work represents a significant and widely applicable improvement to the state of the art in synthesizing quantum circuits with low T gate counts.
 
-We note that the T-count of the temporary logical-AND is at least 2 in \autoref{fig:lower-bound-temporary-and}.
-Two is less than four, but we are not aware of any method for further reducing the T-count of temporary-ANDs or for synthesizing more T gates out of a temporary-AND.
-We are unsure if addition or temporary logical-AND gates can be done with fewer, but certain that T-count improvements as significant as the ones we have made in this paper are possible for other basic operations used by many circuits.
+We are unsure if addition or temporary logical-AND gates can be done with fewer.
+A simple argument based on synthesizing T gates proves that the T-count of the temporary logical-AND is at least 2 in \autoref{fig:lower-bound-logical-AND}.
+Two is less than four, but we are not aware of any method for further reducing the T-count of temporary logical-ANDs or for synthesizing more T gates out of a temporary logical-AND.
 
@@
   \caption{
-	Three T gates can be synthesized by performing a temporary-AND, one T gate, and various Clifford operations.
-	Clifford operations can't produce T gates, so the extra two gates must be coming from the temporary-AND.
-	This proves that at least two T gates went into performing the temporary-AND.
+	Three T gates can be synthesized by performing a temporary logical-AND, one T gate, and various Clifford operations.
+	Clifford operations can't produce T gates, so the extra two gates must be coming from the temporary logical-AND.
+	This proves that at least two T gates went into performing the temporary logical-AND.
   }
-  \label{fig:lower-bound-temporary-and}
+  \label{fig:lower-bound-logical-AND}
 \end{figure}
 
-Our work represents a significant and widely applicable improvement to the state of the art in synthesizing quantum circuits with low T gate counts.
+Regardless, it is clear that other T-count improvements as significant as the ones we have made in this paper are waiting to be found.
+Not just in the construction of basic low-level operations, but in medium-level constructions and high-level constructions and generally across the whole technology stack that will be needed to perform error corrected quantum computation for the first time.
 

### 2017-08-29 paper.tex
Citations and bibliography style

@@
-\documentclass[twocolumn,longbibliography]{quantumarticle-customized}
+\documentclass{quantumarticle-customized}
 \usepackage{amssymb}
@@
 \usepackage{graphicx}
+\usepackage[sort,numbers]{natbib}
 \usepackage[pdfpagelabels,pdftex,bookmarks,breaklinks]{hyperref}
@@
 \begin{abstract}
-We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Amy2013, Cuccaro2004, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
+We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \citep{Amy2013, Cuccaro2004, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
 We do so via a ``temporary logical-AND" construction, which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
@@
 
-The surface code is a quantum error correcting code that works on a 2D nearest-neighbour array of qubits and achieves a threshold error rate of approximately 1\% [[[[cite so much]]]].
-This makes the surface code a likely component in the architecture of future error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies [[[[cite so much]]]] and other well understood error correcting codes either have lower thresholds or require stronger connectivity [[[[cite so much]]]].
+The surface code \citep{Brav98,Denn02,Raus07,Raus07d,Fowler2012} is a quantum error correcting code that works on a 2D nearest-neighbour array of qubits and achieves a threshold error rate of approximately 1\%.
+This makes the surface code a likely component in the architecture of future error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies \citep{Schl11,Bare13,Gamb17,Leik17,Laht17} and other well understood error correcting codes either have lower thresholds or require stronger connectivity.
 
@@
 
-The textbook construction of a Toffoli gate uses seven T gates \cite{Nielsen2009}.
+The textbook construction of a Toffoli gate uses seven T gates \citep{Nielsen2009}.
 When Toffoli operations are paired, i.e.
 when an initial Toffoli operation is later uncomputed by a second Toffoli operation, each Toffoli in the pair can omit three of the T gates from the textbook construction.
-This introduces phase errors but, assuming intermediate operations aren't poisoned by the phase errors, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \cite{Barenco1995, Nielsen2009}.
-It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and a classically conditioned fixup operations \cite{Jones2013}.
+This introduces phase errors but, assuming intermediate operations aren't poisoned by the phase errors, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \citep{Barenco1995, Nielsen2009}.
+It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and a classically conditioned fixup operations \citep{Jones2013}.
 
 Existing Toffoli constructions used at least 4 T gates per Toffoli gate.
-The Cuccaro adder uses $2n-1$ Toffoli gates \cite{Cuccaro2004}.
-Correspondingly, existing adder constructions have T-counts of $8n-4$ \cite{Amy2013}.
-Adders with a T-count of $8n + O(1)$ have been known for over a decade \cite{Barenco1995, Cuccaro2004}.
-The leading factor of 8 has been conjectured to be optimal \cite{AustinDiscussionsAndEmails2017}.
+The Cuccaro adder uses $2n-1$ Toffoli gates \citep{Cuccaro2004}.
+Correspondingly, existing adder constructions have T-counts of $8n-4$ \citep{Amy2013}.
+Adders with a T-count of $8n + O(1)$ have been known for over a decade \citep{Barenco1995, Cuccaro2004}.
+The leading factor of 8 has been conjectured to be optimal \citep{AustinDiscussionsAndEmails2017}.
 We show that this is not true by reducing the T-count of addition to $4n + O(1)$.
@@
 Our adder circuits use temporary logical-AND operations.
-We define the temporary logical-AND computation (which we draw as a wire emerging from two controls), and the corresponding uncomputation (which we draw as a wire merging into two controls), in \autoref{fig:compute-logical-AND}.
+We define the temporary logical-AND computation (which we draw as a wire emerging from two controls), and the corresponding uncomputation (which we draw as a wire merging into two controls), in \autoref{fig:temporary-logical-AND}.
 Computing the temporary logical-AND has a T-count of 4, but when uncomputing a temporary logical-AND we have access to measurement and only need Clifford operations.
@@
 Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
-\cite{Coreas2017} derives a controlled-addition construction with T-count of $21n + O(1)$.
+\citep{Coreas2017} derives a controlled-addition construction with T-count of $21n + O(1)$.
 We use temporary logical-ANDs to construct an improved circuit with a T-count of $8n + O(1)$ (see \autoref{fig:controlled-full-adder-block} for the building-block).
@@
 Decreasing the T-count of addition reduces the T-count of any construction based on addition.
-For example, in \cite{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$.
+For example, in \citep{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$.
 The time estimate is based on Toffolis having a T-depth of 3, and the $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
@@
 Controlled phase gradients can be applied by adding into the reusable phase gradient register with controlled adders instead of uncontrolled adders.
-Assume a maximum per-gate error of $\epsilon$, start from the textbook QFT circuit based on the Cooley-Tukey algorithm from \cite{Nielsen2009}, drop negligible rotations, merge individual controlled rotations into controlled phase gradients, and apply those controlled phase gradients using controlled adders.
+Assume a maximum per-gate error of $\epsilon$, start from the textbook QFT circuit based on the Cooley-Tukey algorithm from \citep{Nielsen2009}, drop negligible rotations, merge individual controlled rotations into controlled phase gradients, and apply those controlled phase gradients using controlled adders.
 The result is a simple approximate $n$-qubit QFT circuit that uses fewer than $8 n \lg \lg \frac{1}{\epsilon}$ T gates.
@@
 For other cases where temporary logical-ANDs can be used to optimize circuits, we refer back to the introduction.
-Even circuits based on simulating physical systems, not just abstract mathematical problems, can benefit \cite{RyanEmails2017}.
+Even circuits based on simulating physical systems, not just abstract mathematical problems, can benefit \citep{RyanEmails2017}.
 
@@
 
-For over a decade, the T-count of addition has been known to be at most $8n - 4$ \cite{Amy2013, Barenco1995, Cuccaro2004}.
-This was conjectured to be optimal \cite{AustinDiscussionsAndEmails2017}.
+For over a decade, the T-count of addition has been known to be at most $8n - 4$ \citep{Amy2013, Barenco1995, Cuccaro2004}.
+This was conjectured to be optimal \citep{AustinDiscussionsAndEmails2017}.
 We halved the leading factor of this cost, and we did so via a construction that applies to many other circuits: the temporary logical-AND gate.
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 \bibliography{citations}

### 2017-08-29 paper.tex
Continuing to polish

@@
 
-Existing Toffoli constructions used at least 4 T gates per Toffoli gate.
+Existing Toffoli constructions use at least 4 T gates per Toffoli gate.
 The Cuccaro adder uses $2n-1$ Toffoli gates \citep{Cuccaro2004}.
@@
 The leading factor of 8 has been conjectured to be optimal \citep{AustinDiscussionsAndEmails2017}.
-We show that this is not true by reducing the T-count of addition to $4n + O(1)$.
 
@@
 We halve the number of T gates needed to perform an adder by halving the number of T gates needed to perform Toffoli gates that appear in compute/uncompute pairs.
-In this paper we focus on how this optimization halves the T-count of addition, but improving the T-count of paired Toffolis is a widely applicable optimization.
-Other examples of operations whose T-count can be improved by our constructions include:
+Although we focus on adders in this paper, paired Toffoli gates appear in a wide variety of circuits.
+Other examples of operations with circuits whose T-count can be improved by our constructions include:
 
-- Integer arithmetic in general.
-
-- Modular arithmetic.
-
-- Applying the same rotation to many qubits.
-
-- The quantum Fourier transform.
-
-- Operations whose target is indexed by a quantum register.
-
-- Phasing a register by a computable function $f$ (i.e. applying the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$).
-
-- Expanding a binary register into a unary register.
+\begin{itemize}
+  \item Integer arithmetic in general.
+  \item Modular arithmetic.
+  \item Applying the same rotation $R_Z(\theta)$ to many qubits.
+  \item The quantum Fourier transform.
+  \item Operations whose target is indexed by a quantum register.
+  \item Phasing a register by a computable function $f$, i.e. applying the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$.
+  \item Expanding a binary register into a unary register.
+\end{itemize}
 
@@
 Our paper is divided into four sections.
-In \autoref{sec:introduction}, we motivate the need to optimize T-counts, discuss existing work, and note possible applications of our low T-count constructions.
+In \autoref{sec:introduction}, we motivate the need to optimize T-counts, discuss existing work, and note the wide applicability of our low T-count constructions.
 \autoref{sec:construction} presents our $4n + O(1)$ T-count adder and explains its pieces, including the temporary logical-AND.
@@
 
-With a slight tweak to the adder building block, we can make the sum available half-way through the adder circuit, before any uncomputation of temporary logical-ANDs occurs.
-When a circuit is going to compute an addition, then later uncompute it, using the sum at the half-way point can halve the T-count.
-Instead of paying $4n + O(1)$ T gates to compute the addition, and then $4n + O(1)$ more T gates to uncompute the addition, we simply work with the intermediate state of a single addition computation then backtrack instead of finishing.
-\autoref{fig:temporary-full-adder-block} shows the temporary adder building block, which is also a temporary three-bit adder with T-count 4.
+\autoref{fig:variations-on-full-adder-block} shows the building-blocks for two variations on our adder: a controlled adder and a temporary adder.
+
+Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
+Our controlled adder reduces the cost of these additions from $21n + O(1)$ \citep{Coreas2017} to $8n + O(1)$.
+
+Temporary adders are useful when a circuit is going to compute an addition, then later uncompute it.
+Instead of paying $4n + O(1)$ T gates to compute an addition, and then $4n + O(1)$ more T gates to uncompute the addition with a subtraction, our temporary adder stops halfway through the normal addition (after all the logical-ANDs have been computed, but before they've been uncomputed) and applies a few cleanup operations so that the sum is available for use.
+When the sum is no longer needed, the partially-performed addition is uncomputed instead of completed.
 
 \begin{figure}
-  \resizebox{\linewidth}{!}{
+  (a) \resizebox{\linewidth}{!}{
     \Qcircuit @R=0.7em @C=0.7em {
-      &a &&\ctrl{2}&\qw     &\qw     &\ctrl{3}&\qw &a          &&&&&&& \\
-      &b &&\targ   &\ctrl{1}&\ctrl{1}&\targ   &\qw &b          &&&&&&& \\
-      &c &&\targ   &\ctrl{1}&\targ   &\qw     &\qw &&&&(a+b+c)_0  &&&& \\
-      &  &&        &        &\qw     &\targ   &\qw &&&&(a+b+c)_1  &&&& \\
+      &&\text{control} &&&&\qw     &\qw     &\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\qw     &\ctrl{2}&\qw     &\ctrl{2}&\qw     &\qw &&&\text{control}               &&&&&&\\
+      &&&c_k            &&&\ctrl{3}&\qw     &\ctrl{4}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{4}&\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw &c_k                          &&&&&&&&\\
+      &&&i_k            &&&\targ   &\ctrl{2}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{2}&\ctrl{1}&\qw     &\ctrl{1}&\targ   &\qw &i_k                          &&&&&&&&\\
+      &&               &&&&        &        &        &    &    &       &    &    &    &       &    &    &    &       &    &    &        &        &        &\ctrl{1}&\qw     &        &    &    &&&&&&&&\\
+      &&&t_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\qw     &\targ   &\qw     &\targ   &\qw &&&&&(t+i \cdot \text{control})_k &&&&&\\
+      &&               &&&&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &        &        &    &    &&&&&&&&\\
+    }
+  }
+  \\ \vspace*{0.25cm} \\
+  (b) \resizebox{\linewidth}{!}{
+    \Qcircuit @R=0.7em @C=0.7em {
+      &a &&\ctrl{2}&\qw     &\qw     &\ctrl{3}&\qw &a          &&&&&&&&&&&&&&&&&&&&&&\\
+      &b &&\targ   &\ctrl{1}&\ctrl{1}&\targ   &\qw &b          &&&&&&&&&&&&&&&&&&&&&&\\
+      &c &&\targ   &\ctrl{1}&\targ   &\qw     &\qw &&&&(a+b+c)_0  &&&&&&&&&&&&&&&&&&&\\
+      &  &&        &        &\qw     &\targ   &\qw &&&&(a+b+c)_1  &&&&&&&&&&&&&&&&&&&\\
     }
@@
   \caption{
-	Temporary-adder circuit building-block with a T-count of 4.
-	Uncomputing the result requires no T gates.
-	Can be used to temporarily add two $n$-bit registers (T-count $4n + O(1)$).
-	Can be used to temporarily compute the pop count of an $n$-bit register (T-count $4n + O(\lg n)$).
+	Variations on our adder construction.
+	(a) is a controlled-adder building-block with a T-count of 8.
+	The addition will only occur when the control is on.
+	(b) is a temporary-adder building-block with a T-count of 4.
+	Has the same T-count as a normal addition, but makes the result available before any uncomputation steps are needed.
   }
-  \label{fig:temporary-full-adder-block}
-\end{figure}
-
-Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
-\citep{Coreas2017} derives a controlled-addition construction with T-count of $21n + O(1)$.
-We use temporary logical-ANDs to construct an improved circuit with a T-count of $8n + O(1)$ (see \autoref{fig:controlled-full-adder-block} for the building-block).
-
-\begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\text{control} &&&&\qw     &\qw     &\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\qw     &\ctrl{2}&\qw     &\qw &&&\text{control}               &&&&&&\\
-      &&c_k            &&&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\qw     &\ctrl{2}&\qw &c_k                          &&&&&&&&\\
-      &&i_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\ctrl{1}&\targ   &\qw &i_k                          &&&&&&&&\\
-      &&t_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\targ   &\targ   &\qw &&&&&(t+i \cdot \text{control})_k &&&&&\\
-      &               &&&&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &    &    &&&&&&&&\\
-    }
-  }
-  \caption{
-	Controlled-adder circuit building-block with a T-count of 8.
-  }
-  \label{fig:controlled-full-adder-block}
+  \label{fig:variations-on-full-adder-block}
 \end{figure}
@@
 Decreasing the T-count of addition reduces the T-count of any construction based on addition.
-For example, in \citep{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$.
+For example, in \citep{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$ states.
 The time estimate is based on Toffolis having a T-depth of 3, and the $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
 Because Shor's algorithm is dominated by the cost of additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $2$ and $1$ respectively.
-This reduces the estimates to 9 hours and $6 \cdot 10^{11}$ distilled $|A\rangle$ states respectively.
+This reduces the estimates to 9 hours and $6 \cdot 10^{11}$ distilled $|A\rangle$ states.
 
 On top of reducing the T-count of obviously related operations like multiplication and exponentiation, reducing the T-count of addition also reduces the T-count of quantum-specific operations such as rotating qubits.
+
 For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a total T-cost $4n + O(\text{poly}(\frac{1}{\epsilon}) \lg n)$.
@@
 Then, for each position $k$ in the pop count output register, synthesize and apply the operation $R_Z(\theta \cdot 2^k)$ to the qubit at that position.
-This uses a number of T gates proportional to some polynomial of the desired precision times the number of bits in the output register.
+This uses a number of T gates proportional to some polynomial of the desired bit precision, times the number of bits in the output register.
 But the output register is exponentially smaller than $n$, so this cost is negligible for large $n$.
@@
 Another rotation-based operation that can be implemented via an adder is the $n$-bit phase gradient operation $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
-Normally this operation would be implemented by separately applying the operation $R_z(\pi 2^{-p})$ to each qubit, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
+Normally this operation would be implemented by separately applying the operation $R_z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
 However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition using only $4n + O(1)$ T gates.
-Just add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
-Ignoring the one-time cost of initializing the reusable phase gradient register, this uses only $4n + O(1)$ T gates.
+Add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
+Ignoring the one-time cost of initializing the reusable phase gradient register, this uses $4n + O(1)$ T gates.
 
 Some quantum Fourier transform circuits use controlled phase gradient operations.
-Controlled phase gradients can be applied by adding into the reusable phase gradient register with controlled adders instead of uncontrolled adders.
-Assume a maximum per-gate error of $\epsilon$, start from the textbook QFT circuit based on the Cooley-Tukey algorithm from \citep{Nielsen2009}, drop negligible rotations, merge individual controlled rotations into controlled phase gradients, and apply those controlled phase gradients using controlled adders.
+Controlled phase gradients can be applied in the same way as normal phase gradients, by adding into a reusable phase gradient register.
+However, controlled adders are used instead of uncontrolled adders.
+
+Assume a maximum per-gate error of $\epsilon$, start from the textbook QFT circuit based on the Cooley-Tukey algorithm \citep{Nielsen2009}, drop negligible rotations, merge individual controlled rotations into controlled phase gradients, and apply those controlled phase gradients using controlled adders.
 The result is a simple approximate $n$-qubit QFT circuit that uses fewer than $8 n \lg \lg \frac{1}{\epsilon}$ T gates.
@@
 The temporary logical-AND is useful for optimizing an even wider variety of circuits than addition is.
-Whenever Toffoli gates appear in compute/uncompute pairs, and intermediate operations are not sensitive to phase errors on the controls of the Toffoli gate (i.e.
+Whenever Toffoli gates appear in compute/uncompute pairs, and intermediate operations are not sensitive to phase errors on the controls of the Toffoli gate (e.g.
 the condition in \autoref{fig:pair-definition} is satisfied), it is possible to save 4 T gates by replacing the pair of Toffoli gates with a temporary logical-AND.
@@
     \Qcircuit @R=0.7em @C=0.7em {
-      &\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &&               &&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &\\
-      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&               &&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &\\
-      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&\stackrel{?}{=}&&          &\ctrl{2}   &\ghost{U}       &\ctrl{2}&\qw &\\
-      &          &\targ      &\ghost{U}       &\targ   &\qw &&               &&          &\targ      &\ghost{U}       &\targ   &\qw &\\
-      &          &\qw        &\qw             &\qw     &\qw &&               &&          &\targ      &\qw             &\targ   &\qw &\\
+      &&&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &&               &&&&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &&            &&\ustick{n}&\diagup\qw &\qw      &\multigate{3}{U}&\qw      &\qw     &\qw &\\
+      &&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&               &&&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&            &&          &\ctrl{1}   &\qw      &\ghost{U}       &\qw      &\ctrl{1}&\qw &\\
+      &&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&\stackrel{?}{=}&&&&          &\ctrl{2}   &\ghost{U}       &\ctrl{2}&\qw &&\Rightarrow &&          &\ctrl{2}   &\qw      &\ghost{U}       &\qw      &\ctrl{2}&\qw &\\
+      &&&          &\targ      &\ghost{U}       &\targ   &\qw &&               &&&&          &\targ      &\ghost{U}       &\targ   &\qw &&            &&          &\qw        &\targ    &\ghost{U}       &\targ    &\qw     &\qw &\\
+      &&&\qO       &\qw        &\qw             &\qw     &\qw &&               &&&&\qO       &\targ      &\qw             &\targ   &\qw &&            &&          &           &\ctrl{-1}&\qw             &\ctrl{-1}&\qw     &    &\\
     }
@@
 
-For over a decade, the T-count of addition has been known to be at most $8n - 4$ \citep{Amy2013, Barenco1995, Cuccaro2004}.
+For over a decade, the T-count of addition has been $8n - 4$ \citep{Amy2013, Barenco1995, Cuccaro2004}.
 This was conjectured to be optimal \citep{AustinDiscussionsAndEmails2017}.
@@
 
-We are unsure if addition or temporary logical-AND gates can be done with fewer.
+We are unsure if addition or temporary logical-AND gates can be done with fewer T gates.
 A simple argument based on synthesizing T gates proves that the T-count of the temporary logical-AND is at least 2 in \autoref{fig:lower-bound-logical-AND}.
@@
 	Three T gates can be synthesized by performing a temporary logical-AND, one T gate, and various Clifford operations.
-	Clifford operations can't produce T gates, so the extra two gates must be coming from the temporary logical-AND.
+	Clifford operations can't produce T gates, so the extra T gates must be coming from the temporary logical-AND.
 	This proves that at least two T gates went into performing the temporary logical-AND.
@@
 Regardless, it is clear that other T-count improvements as significant as the ones we have made in this paper are waiting to be found.
-Not just in the construction of basic low-level operations, but in medium-level constructions and high-level constructions and generally across the whole technology stack that will be needed to perform error corrected quantum computation for the first time.
+Not just in the construction of basic low-level operations, but in medium-level constructions and high-level constructions and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
 

### 2017-09-02 paper.tex
revtex4
Polishing

@@
-\documentclass{quantumarticle-customized}
+\documentclass[twocolumn]{revtex4-1}
 \usepackage{amssymb}
@@
 \usepackage{graphicx}
-\usepackage[sort,numbers]{natbib}
 \usepackage[pdfpagelabels,pdftex,bookmarks,breaklinks]{hyperref}
@@
 
+\def\sectionautorefname{Section}
+
+\begin{document}
 \title{Halving the cost of quantum addition}
@@
 \email{craiggidney@google.com}
-
-\def\sectionautorefname{Section}
-
-\begin{document}
-\maketitle
 
@@
 
-In addition to our $n$-bit adder circuit with T-count of $4n + O(1)$, we show how to construct an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, and a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until later uncomputed without T gates.
+We also show how to construct an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, and a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until it is later uncomputed without using T gates.
 \end{abstract}
+
+\maketitle
 
@@
 Because T gates are so expensive for the surface code, and the surface code is a likely component of future quantum computers, it is important to consider and optimize the number of T gates used by quantum circuits.
-Optimizing the T-count of basic elements of quantum circuits, such as the construction of adders and Toffoli gates, is even more important.
+Optimizing the T-count of basic elements of quantum circuits, such as the construction of adders and Toffoli gates, is particularly important because any improvement is widely applicable.
 
@@
 when an initial Toffoli operation is later uncomputed by a second Toffoli operation, each Toffoli in the pair can omit three of the T gates from the textbook construction.
-This introduces phase errors but, assuming intermediate operations aren't poisoned by the phase errors, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \citep{Barenco1995, Nielsen2009}.
-It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and a classically conditioned fixup operations \citep{Jones2013}.
+This introduces phase errors but, assuming intermediate operations aren't sensitive to the phase errors, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \citep{Barenco1995, Nielsen2009}.
+It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and a classically conditioned fixup operation \citep{Jones2013}.
 
@@
 The Cuccaro adder uses $2n-1$ Toffoli gates \citep{Cuccaro2004}.
-Correspondingly, existing adder constructions have T-counts of $8n-4$ \citep{Amy2013}.
-Adders with a T-count of $8n + O(1)$ have been known for over a decade \citep{Barenco1995, Cuccaro2004}.
+Correspondingly, existing adder constructions have T-counts of $8n + O(1)$ \citep{Amy2013} and this has been the case for over a decade \citep{Barenco1995, Cuccaro2004}.
 The leading factor of 8 has been conjectured to be optimal \citep{AustinDiscussionsAndEmails2017}.
@@
 Although we focus on adders in this paper, paired Toffoli gates appear in a wide variety of circuits.
-Other examples of operations with circuits whose T-count can be improved by our constructions include:
-
-\begin{itemize}
-  \item Integer arithmetic in general.
-  \item Modular arithmetic.
-  \item Applying the same rotation $R_Z(\theta)$ to many qubits.
-  \item The quantum Fourier transform.
-  \item Operations whose target is indexed by a quantum register.
-  \item Phasing a register by a computable function $f$, i.e. applying the operation $U_f = \exp\left( i \sum_k f(k) |k\rangle \langle k| \right)$.
-  \item Expanding a binary register into a unary register.
-\end{itemize}
-
-The above list is not exhaustive.
-We leave further low hanging fruit as an exercise for the reader.
+We discuss some specific examples in \autoref{sec:other-applications} (e.g.
+applying the same rotation $R_Z(\theta)$ to several qubits), but leave further low hanging fruit as an exercise for the reader.
 
@@
 It performs 4 temporary logical-ANDs, each with a T-count of 4.
+All other operations are Clifford operations, with no T-count.
 
-We construct $n$-bit adders by nesting $n$ copies of the building block shown in \autoref{fig:full-adder-block} inside of each other.
-We then specialize the outer-most and inner-most blocks (which act on the low bit and high bit respectively) based on the fact that they either have no carry input or no carry output.
+The building block of our adder is shown in \autoref{fig:full-adder-block}.
+We construct $n$-bit adders by nesting $n$ copies of the building block inside of each other.
+The outer-most and inner-most blocks (which act on the low bit and high bit respectively) are then specialized based on the fact that they either have no carry input or no carry output.
 
@@
 
-Our adder circuits use temporary logical-AND operations.
-We define the temporary logical-AND computation (which we draw as a wire emerging from two controls), and the corresponding uncomputation (which we draw as a wire merging into two controls), in \autoref{fig:temporary-logical-AND}.
-Computing the temporary logical-AND has a T-count of 4, but when uncomputing a temporary logical-AND we have access to measurement and only need Clifford operations.
+Our adder uses temporary logical-AND operations, which we draw as wires emerging out of a pair of controls then later merging into an equivalent pair of controls.
+\autoref{fig:temporary-logical-AND} shows how we compute the logical-AND of the two controls (which we draw as a wire emerging from two controls then heading rightward), and also the corresponding uncomputation (which we draw as a wire turning upward into two controls).
+
+Computing the temporary logical-AND has a T-count of 4, but uncomputing it has a T-count of 0.
+This asymmetry is due to the fact that measurement is not reversible.
+The uncomputation uses measurement in a way that the computation can not.
+
+\section{Other Applications}
+\label{sec:other-applications}
+
+\autoref{fig:variations-on-full-adder-block} shows the building-blocks for two variations on our adder: a controlled adder and a temporary adder.
+
+Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
+Our controlled adder reduces the cost of these additions from $21n + O(1)$ \citep{Coreas2017} to $8n + O(1)$.
+
+Temporary adders are useful when a circuit is going to compute an addition, then later uncompute it.
+Instead of paying $4n + O(1)$ T gates to compute an addition, and then $4n + O(1)$ more T gates to uncompute the addition with a subtraction, our temporary adder stops halfway through the normal addition (after all the logical-ANDs have been computed, but before they've been uncomputed) and applies a few cleanup operations so that the sum is available for use.
+When the sum is no longer needed, the partially-performed addition is uncomputed instead of completed.
 
@@
   }
-  \\ \vspace*{0.25cm} \\
+  \vspace*{0.25cm} \\
   \resizebox{\linewidth}{!}{
@@
     This alternative approach has a net T-count of 2 (because an $|A\rangle$ state is recovered).
-    That would still be an improvement on existing work, but inferior to the measure-and-fixup construction shown above.
+    The resulting total T-count of 6 for the entire temporary logical-AND would still be an improvement on existing work, but is inferior to the measure-and-fixup construction shown above.
   }
@@
 
-\section{Other Applications}
-\label{sec:other-applications}
+Decreasing the T-count of addition reduces the T-count of any construction based on addition.
+For example, in \citep{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$ states.
+The time estimate is based on Toffolis having a T-depth of 3, and the $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
+Because Shor's algorithm is dominated by the cost of controlled additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $\approx 2.7$ and $1$ respectively.
+This reduces the estimates to 9 hours and $8 \cdot 10^{11}$ distilled $|A\rangle$ states.
 
-\autoref{fig:variations-on-full-adder-block} shows the building-blocks for two variations on our adder: a controlled adder and a temporary adder.
+On top of reducing the T-count of obviously-related classical operations like multiplication and exponentiation, reducing the T-count of addition also reduces the T-count of quantum-specific operations such as rotating qubits.
 
-Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
-Our controlled adder reduces the cost of these additions from $21n + O(1)$ \citep{Coreas2017} to $8n + O(1)$.
+For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a total T-cost $4n + O(\text{poly}(\lg \frac{1}{\epsilon}) \lg n)$ as follows.
+First, iteratively apply our temporary adder to combine the target qubits into a single binary register storing their pop count (the number that are in the state $|1\rangle$).
+This takes $4n + O(\lg n)$ T gates to do.
+Then, for each position $k$ in the pop count register, synthesize and apply the operation $R_Z(\theta \cdot 2^k)$ to the register qubit at that position.
+This uses a number of T gates proportional to some polynomial of the desired bit precision, times the number of bits in the output register.
+The output register is exponentially smaller than $n$, so this cost is negligible for large $n$.
+Finally, uncompute the temporary pop count (using no T gates).
+This applies the $n$ desired $R_Z(\theta)$ operations.
 
-Temporary adders are useful when a circuit is going to compute an addition, then later uncompute it.
-Instead of paying $4n + O(1)$ T gates to compute an addition, and then $4n + O(1)$ more T gates to uncompute the addition with a subtraction, our temporary adder stops halfway through the normal addition (after all the logical-ANDs have been computed, but before they've been uncomputed) and applies a few cleanup operations so that the sum is available for use.
-When the sum is no longer needed, the partially-performed addition is uncomputed instead of completed.
+Another rotation-based operation that can be implemented via an adder is the $n$-bit phase gradient operation $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
+Normally this operation would be implemented by separately applying the operation $R_z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
+However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition using only $4n + O(1)$ T gates.
+Add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
+Ignoring the one-time cost of initializing the reusable phase gradient register, this uses $4n + O(1)$ T gates.
 
@@
   }
-  \\ \vspace*{0.25cm} \\
+  \vspace*{0.25cm} \\
   (b) \resizebox{\linewidth}{!}{
@@
 \end{figure}
-
-Decreasing the T-count of addition reduces the T-count of any construction based on addition.
-For example, in \citep{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$ states.
-The time estimate is based on Toffolis having a T-depth of 3, and the $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
-Because Shor's algorithm is dominated by the cost of additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $2$ and $1$ respectively.
-This reduces the estimates to 9 hours and $6 \cdot 10^{11}$ distilled $|A\rangle$ states.
-
-On top of reducing the T-count of obviously related operations like multiplication and exponentiation, reducing the T-count of addition also reduces the T-count of quantum-specific operations such as rotating qubits.
-
-For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a total T-cost $4n + O(\text{poly}(\frac{1}{\epsilon}) \lg n)$.
-First, iteratively apply our temporary adder to compute the pop count of a register containing the target qubits.
-This takes $4n + O(\lg n)$ T gates to do.
-Then, for each position $k$ in the pop count output register, synthesize and apply the operation $R_Z(\theta \cdot 2^k)$ to the qubit at that position.
-This uses a number of T gates proportional to some polynomial of the desired bit precision, times the number of bits in the output register.
-But the output register is exponentially smaller than $n$, so this cost is negligible for large $n$.
-Finally, uncompute the temporary pop count (using no T gates).
-This applies the desired $R_Z(\theta)$ operations.
-
-Another rotation-based operation that can be implemented via an adder is the $n$-bit phase gradient operation $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
-Normally this operation would be implemented by separately applying the operation $R_z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
-However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition using only $4n + O(1)$ T gates.
-Add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
-Ignoring the one-time cost of initializing the reusable phase gradient register, this uses $4n + O(1)$ T gates.
-
-Some quantum Fourier transform circuits use controlled phase gradient operations.
-Controlled phase gradients can be applied in the same way as normal phase gradients, by adding into a reusable phase gradient register.
-However, controlled adders are used instead of uncontrolled adders.
-
-Assume a maximum per-gate error of $\epsilon$, start from the textbook QFT circuit based on the Cooley-Tukey algorithm \citep{Nielsen2009}, drop negligible rotations, merge individual controlled rotations into controlled phase gradients, and apply those controlled phase gradients using controlled adders.
-The result is a simple approximate $n$-qubit QFT circuit that uses fewer than $8 n \lg \lg \frac{1}{\epsilon}$ T gates.
 
@@
 
-For example, the Grover diffusion operator's cost is dominated by the cost of merging $n$ controls.
-Merging the controls can be computed, and then uncomputed, with Toffoli gates.
-By instead iteratively merging the controls with nested temporary logical-ANDs, the Grover diffusion operation can be applied using $4n + O(1)$ T gates.
+For example, a simple way to construct Grover oracles is to translate a classical predicate into a reversible circuit made out of Toffoli gates.
+This classical reversible circuit is then used to compute an output bit, so that a Z gate can be applied to the output bit before the circuit is run in reverse to uncompute the output.
+Every Toffoli gate generated by this construction is paired (each Toffoli in the computation will match a Toffol in the uncomputation).
+Using temporary logical-ANDs halves the T-count of this approach to constructing Grover oracles.
 
-For other cases where temporary logical-ANDs can be used to optimize circuits, we refer back to the introduction.
-Even circuits based on simulating physical systems, not just abstract mathematical problems, can benefit \citep{RyanEmails2017}.
+Matched Toffolis also appear in quantum circuits rooted in physics instead of abstract mathematics.
+For example, we used temporary logical-ANDs to cut the projected T-count of a step in a simulation algorithm under development by nearly 30\% \citep{RyanEmails2017}.
+We re-iterate that this is a very widely applicable optimization.
 
@@
 
-For over a decade, the T-count of addition has been $8n - 4$ \citep{Amy2013, Barenco1995, Cuccaro2004}.
-This was conjectured to be optimal \citep{AustinDiscussionsAndEmails2017}.
-We halved the leading factor of this cost, and we did so via a construction that applies to many other circuits: the temporary logical-AND gate.
-Our work represents a significant and widely applicable improvement to the state of the art in synthesizing quantum circuits with low T gate counts.
+For over a decade, the T-count of addition has been $8n + O(1)$ \citep{Amy2013, Barenco1995, Cuccaro2004}.
+In this paper showed how to halve the leading factor of this cost.
+Furthermore, we did so via a construction that applies to many other circuits: the temporary logical-AND gate.
+Because the temporary logical-AND is so widely applicable, many existing estimates of the T-count of circuits need to be redone.
 
-We are unsure if addition or temporary logical-AND gates can be done with fewer T gates.
-A simple argument based on synthesizing T gates proves that the T-count of the temporary logical-AND is at least 2 in \autoref{fig:lower-bound-logical-AND}.
-Two is less than four, but we are not aware of any method for further reducing the T-count of temporary logical-ANDs or for synthesizing more T gates out of a temporary logical-AND.
+It is interesting to consider if addition or temporary logical-AND gates can be done with still fewer T gates.
+\autoref{fig:lower-bound-logical-AND} presents a simple proof that the T-count of the temporary logical-AND is at least 2.
+We expect that the true lower bound is 4, but that there will be special cases with a lower cost.
+For example, when a register is conditionally incremented several times, it could be possible to cancel some of the T-count-using computation work at the start of an increment against the uncomputation work at the end of the previous increment.
 
@@
 
-Regardless, it is clear that other T-count improvements as significant as the ones we have made in this paper are waiting to be found.
-Not just in the construction of basic low-level operations, but in medium-level constructions and high-level constructions and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
+Regardless of whether there will be further T-count improvements for an operation as basic as the temporary logical-AND, it is clear that many other T-count optimizations are waiting to be found.
+Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
 
@@
 
-We thank Austin Fowler for assistance in locating relevant references, sampling the opinions of other researchers, and for comments that greatly improved our paper.
+We thank Austin Fowler for sampling the opinions of other researchers, assistance locating references, and comments that greatly improved our paper.
 

### 2017-09-02 paper.tex
Tweak polish

@@
 
-We also show how to construct an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, and a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until it is later uncomputed without using T gates.
+We also present an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until it is later uncomputed without using T gates, and discuss some other constructions whose T-count is improved by the temporary logical-AND.
 \end{abstract}
@@
 Existing Toffoli constructions use at least 4 T gates per Toffoli gate.
-The Cuccaro adder uses $2n-1$ Toffoli gates \citep{Cuccaro2004}.
+The Cuccaro adder uses $2n + O(1)$ Toffoli gates \citep{Cuccaro2004}.
 Correspondingly, existing adder constructions have T-counts of $8n + O(1)$ \citep{Amy2013} and this has been the case for over a decade \citep{Barenco1995, Cuccaro2004}.
@@
 Although we focus on adders in this paper, paired Toffoli gates appear in a wide variety of circuits.
-We discuss some specific examples in \autoref{sec:other-applications} (e.g.
-applying the same rotation $R_Z(\theta)$ to several qubits), but leave further low hanging fruit as an exercise for the reader.
+We discuss some specific examples in \autoref{sec:other-applications} (e.g. applying the same rotation $R_Z(\theta)$ to several qubits).
 
 Our paper is divided into four sections.
-In \autoref{sec:introduction}, we motivate the need to optimize T-counts, discuss existing work, and note the wide applicability of our low T-count constructions.
+We motivate the need to optimize T-counts and discuss existing work in \autoref{sec:introduction}.
 \autoref{sec:construction} presents our $4n + O(1)$ T-count adder and explains its pieces, including the temporary logical-AND.
-\autoref{sec:other-applications} discusses several circuit constructions and estimates improved by our adder and temporary logical-AND constructions.
+In \autoref{sec:other-applications}, we discuss several circuit constructions and estimates improved by our adder and temporary logical-AND constructions.
 Finally, \autoref{sec:conclusion} summarizes our contributions and discusses future work.
@@
 For example, we used temporary logical-ANDs to cut the projected T-count of a step in a simulation algorithm under development by nearly 30\% \citep{RyanEmails2017}.
-We re-iterate that this is a very widely applicable optimization.
 
@@
 For over a decade, the T-count of addition has been $8n + O(1)$ \citep{Amy2013, Barenco1995, Cuccaro2004}.
-In this paper showed how to halve the leading factor of this cost.
+In this paper we showed how to halve the leading factor of this cost.
 Furthermore, we did so via a construction that applies to many other circuits: the temporary logical-AND gate.
-Because the temporary logical-AND is so widely applicable, many existing estimates of the T-count of circuits need to be redone.
+
+Because the temporary logical-AND is so widely applicable, the T-count of many circuits is lower than previously thought.
+We covered a few examples in this paper but, for brevity, we leave further low hanging fruit as an exercise for the reader.
 
@@
 \autoref{fig:lower-bound-logical-AND} presents a simple proof that the T-count of the temporary logical-AND is at least 2.
-We expect that the true lower bound is 4, but that there will be special cases with a lower cost.
-For example, when a register is conditionally incremented several times, it could be possible to cancel some of the T-count-using computation work at the start of an increment against the uncomputation work at the end of the previous increment.
+Regardless of whether the true lower bound is 4, we expect that there are special cases with lower average T-counts.
+For example, consider a register being conditionally incremented several times.
+Incrementing is performed with paired Toffoli gates that can be replaced by temporary logical-ANDs.
+An alternative way to reduce the T-count is by cancelling computation work at the start of an increment against uncomputation work from the end of the previous increment.
+The alternative approach happens to have the same T-count in this case, but that may not always be the case.
 
@@
 	Clifford operations can't produce T gates, so the extra T gates must be coming from the temporary logical-AND.
-	This proves that at least two T gates went into performing the temporary logical-AND.
+	This proves that performing the temporary logical-AND requires at least two T gates.
   }
@@
 
-Regardless of whether there will be further T-count improvements for an operation as basic as the temporary logical-AND, it is clear that many other T-count optimizations are waiting to be found.
-Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
+Regardless of whether there will be further T-count improvements for operations as basic as the temporary logical-AND, it is clear that many other T-count optimizations are waiting to be found.
+Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions, and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
 

### 2017-09-12 paper.tex
Polishing based on Dave's initial comments from weekly 1:1

@@
 
-Existing Toffoli constructions use at least 4 T gates per Toffoli gate.
 The Cuccaro adder uses $2n + O(1)$ Toffoli gates \citep{Cuccaro2004}.
-Correspondingly, existing adder constructions have T-counts of $8n + O(1)$ \citep{Amy2013} and this has been the case for over a decade \citep{Barenco1995, Cuccaro2004}.
-The leading factor of 8 has been conjectured to be optimal \citep{AustinDiscussionsAndEmails2017}.
+Existing constructions implement the Cuccaro adder's Toffolis using $8n + O(1)$ T gates \citep{Barenco1995, Cuccaro2004, Amy2013}.
+This T-count can be achieved either by applying the ancilla-and-fixup construction from \citep{Jones2013} to each individual Toffoli gate, or by noting that all but one of the adder's Toffoli gates appear in compute/uncompute pairs and then applying the matched-phase-error construction from \citep{Barenco1995, Nielsen2009} to each pair.
+
+The leading factor of 8 in the T-count of an $n$-bit adder has stood for over a decade \citep{Barenco1995, Cuccaro2004}, and has been conjectured to be optimal \cite{AustinDiscussionsAndEmails2017}.
+We disprove this conjecture with a construction based on the temporary logical-AND, which combines ideas from both the matched-phase-error construction and the ancilla-and-fix-up construction.
+The temporary logical-AND halves the number of T gates needed to perform Toffoli gates that appear in compute/uncompute pairs, and thereby halves the T-count of the Cuccaro adder from $8n + O(1)$ to $4n + O(1)$.
 
@@
 
-We halve the number of T gates needed to perform an adder by halving the number of T gates needed to perform Toffoli gates that appear in compute/uncompute pairs.
-Although we focus on adders in this paper, paired Toffoli gates appear in a wide variety of circuits.
-We discuss some specific examples in \autoref{sec:other-applications} (e.g. applying the same rotation $R_Z(\theta)$ to several qubits).
-
 Our paper is divided into four sections.
@@
 \autoref{sec:construction} presents our $4n + O(1)$ T-count adder and explains its pieces, including the temporary logical-AND.
-In \autoref{sec:other-applications}, we discuss several circuit constructions and estimates improved by our adder and temporary logical-AND constructions.
+In \autoref{sec:other-applications}, we justify the assertion that our constructions represent widely applicable optimizations by giving examples of quantum tasks and cost projections that are improved by efficient adders and temporary logical-ANDs.
 Finally, \autoref{sec:conclusion} summarizes our contributions and discusses future work.
@@
 	How to compute and uncompute the logical-AND of two qubits.
-	The computation circuit has a T-count of 4 and a T-depth of 1.
+	The computation circuit (top) has a T-count of 4 and a T-depth of 1.
 	Note that the $|A\rangle$ state input contributes to the T-count, because $|A\rangle$ states are the resource used to perform T gates.
-	The uncomputation circuit only uses Clifford gates, and so has a T-count of zero.
+	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero.
     \\
@@
     This alternative approach has a net T-count of 2 (because an $|A\rangle$ state is recovered).
-    The resulting total T-count of 6 for the entire temporary logical-AND would still be an improvement on existing work, but is inferior to the measure-and-fixup construction shown above.
+    The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.
   }
@@
 The time estimate is based on Toffolis having a T-depth of 3, and the $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
-Because Shor's algorithm is dominated by the cost of controlled additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $\approx 2.7$ and $1$ respectively.
+Because Shor's algorithm is dominated by the cost of controlled additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $\sim 2.7$ and $1$ respectively.
 This reduces the estimates to 9 hours and $8 \cdot 10^{11}$ distilled $|A\rangle$ states.
@@
 For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a total T-cost $4n + O(\text{poly}(\lg \frac{1}{\epsilon}) \lg n)$ as follows.
-First, iteratively apply our temporary adder to combine the target qubits into a single binary register storing their pop count (the number that are in the state $|1\rangle$).
+First, iteratively apply our temporary adder to combine the target qubits into a single binary register storing their Hamming weight.
 This takes $4n + O(\lg n)$ T gates to do.
-Then, for each position $k$ in the pop count register, synthesize and apply the operation $R_Z(\theta \cdot 2^k)$ to the register qubit at that position.
+Then, for each position $k$ in the Hamming weight register, synthesize and apply the operation $R_Z(\theta \cdot 2^k)$ to the register qubit at that position.
 This uses a number of T gates proportional to some polynomial of the desired bit precision, times the number of bits in the output register.
 The output register is exponentially smaller than $n$, so this cost is negligible for large $n$.
-Finally, uncompute the temporary pop count (using no T gates).
+Finally, uncompute the Hamming weight register (using no T gates).
 This applies the $n$ desired $R_Z(\theta)$ operations.
@@
 Normally this operation would be implemented by separately applying the operation $R_z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
-However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition using only $4n + O(1)$ T gates.
+However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{-2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition.
 Add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
@@
 Regardless of whether the true lower bound is 4, we expect that there are special cases with lower average T-counts.
-For example, consider a register being conditionally incremented several times.
-Incrementing is performed with paired Toffoli gates that can be replaced by temporary logical-ANDs.
-An alternative way to reduce the T-count is by cancelling computation work at the start of an increment against uncomputation work from the end of the previous increment.
-The alternative approach happens to have the same T-count in this case, but that may not always be the case.
+For example, if a task involves repeating an action several times (e.g.
+counting by repeatedly conditionally incrementing a register), T gates from Toffolis in uncomputation work at the end of one repetition may cancel against T gates from Toffolis in computation work at the start of the next repetition.
 
@@
 
-Regardless of whether there will be further T-count improvements for operations as basic as the temporary logical-AND, it is clear that many other T-count optimizations are waiting to be found.
+Regardless of whether there will be further T-count improvements for operations as basic as matched Toffoli gates, it is clear that many other T-count optimizations are waiting to be found.
 Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions, and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.

### 2017-09-12 paper.tex
Note that adder block is a variant of Cuccaro's, and improve the optimization condition figure
Applying more Dave advice - Noted the cost of initializing a phase gradient register (and generally rewrote that paragraph) - Standardized on 'Toffoli pair' instead of 'matched Toffoli' - Polished the conclusion

@@
 	Adder circuit building-block with a T-count of 4.
+	A variant of the Cuccaro adder building-block \citep{Cuccaro2004}.
   }
@@
 
-For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a total T-cost $4n + O(\text{poly}(\lg \frac{1}{\epsilon}) \lg n)$ as follows.
+For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a T-cost of $4n + O(\text{poly}(\lg \frac{1}{\epsilon}) \lg n)$ as follows.
 First, iteratively apply our temporary adder to combine the target qubits into a single binary register storing their Hamming weight.
@@
 
-Another rotation-based operation that can be implemented via an adder is the $n$-bit phase gradient operation $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
+Another quantum operation that can be implemented via an adder is the $n$-qubit phase gradient operation $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
 Normally this operation would be implemented by separately applying the operation $R_z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
@@
 Add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
-Ignoring the one-time cost of initializing the reusable phase gradient register, this uses $4n + O(1)$ T gates.
+This construction performs phase gradients using $4n + O(1)$ T gates, which is apparently independent of $\epsilon$.
+However, there are in fact three ways in which the construction's cost does depend on $\epsilon$.
+First, $\epsilon$ still bounds the minimum quality of the $|A\rangle$ states powering the T gates.
+Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\lg \lg \epsilon$.
+Third, initializing the reusable phase gradient register has a one-time cost of $O(n_{\text{max}} \cdot \text{poly}(\lg \frac{1}{\epsilon}))$ T gates.
 
@@
     \Qcircuit @R=0.7em @C=0.7em {
-      &&&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &&               &&&&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &&            &&\ustick{n}&\diagup\qw &\qw      &\multigate{3}{U}&\qw      &\qw     &\qw &\\
-      &&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&               &&&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&            &&          &\ctrl{1}   &\qw      &\ghost{U}       &\qw      &\ctrl{1}&\qw &\\
-      &&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&\stackrel{?}{=}&&&&          &\ctrl{2}   &\ghost{U}       &\ctrl{2}&\qw &&\Rightarrow &&          &\ctrl{2}   &\qw      &\ghost{U}       &\qw      &\ctrl{2}&\qw &\\
-      &&&          &\targ      &\ghost{U}       &\targ   &\qw &&               &&&&          &\targ      &\ghost{U}       &\targ   &\qw &&            &&          &\qw        &\targ    &\ghost{U}       &\targ    &\qw     &\qw &\\
-      &&&\qO       &\qw        &\qw             &\qw     &\qw &&               &&&&\qO       &\targ      &\qw             &\targ   &\qw &&            &&          &           &\ctrl{-1}&\qw             &\ctrl{-1}&\qw     &    &\\
+      &\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &&               &&&&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &          &&            &&\ustick{n}&\diagup\qw &\qw      &\multigate{3}{U}&\qw      &\qw     &\qw &\\
+      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&               &&&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &          &&            &&          &\ctrl{1}   &\qw      &\ghost{U}       &\qw      &\ctrl{1}&\qw &\\
+      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&\stackrel{?}{=}&&&&          &\ctrl{2}   &\ghost{U}       &\ctrl{2}&\qw &          &&\Rightarrow &&          &\ctrl{2}   &\qw      &\ghost{U}       &\qw      &\ctrl{2}&\qw &\\
+      &          &\targ      &\ghost{U}       &\targ   &\qw &&               &&&&          &\targ      &\ghost{U}       &\targ   &\qw &          &&            &&          &\qw        &\targ    &\ghost{U}       &\targ    &\qw     &\qw &\\
+      &          &           &                &        &    &&               &&&&\qO       &\targ      &\qw             &\targ   &\qw &|0\rangle &&            &&          &           &\ctrl{-1}&\qw             &\ctrl{-1}&\qw     &    &\\
     }
@@
 This classical reversible circuit is then used to compute an output bit, so that a Z gate can be applied to the output bit before the circuit is run in reverse to uncompute the output.
-Every Toffoli gate generated by this construction is paired (each Toffoli in the computation will match a Toffol in the uncomputation).
+Every Toffoli gate generated by this construction is part of a compute/uncompute pair (each Toffoli in the computation will match a Toffoli in the uncomputation).
 Using temporary logical-ANDs halves the T-count of this approach to constructing Grover oracles.
 
-Matched Toffolis also appear in quantum circuits rooted in physics instead of abstract mathematics.
+Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of abstract mathematics.
 For example, we used temporary logical-ANDs to cut the projected T-count of a step in a simulation algorithm under development by nearly 30\% \citep{RyanEmails2017}.
@@
 For over a decade, the T-count of addition has been $8n + O(1)$ \citep{Amy2013, Barenco1995, Cuccaro2004}.
-In this paper we showed how to halve the leading factor of this cost.
-Furthermore, we did so via a construction that applies to many other circuits: the temporary logical-AND gate.
+In this paper we showed how to halve the leading factor of this cost by replacing Toffolis in compute/uncompute pairs with temporary logical-ANDs.
+The temporary logical-AND is a basic circuit construction widget that can be used in many circuits, and a good example of measurement breaking the symmetry between computation and uncomputation (allowing one to be more efficient than the other).
+We covered a few examples of optimizing T-counts with our constructions but, for brevity, we leave further low hanging fruit as an exercise for the reader.
 
-Because the temporary logical-AND is so widely applicable, the T-count of many circuits is lower than previously thought.
-We covered a few examples in this paper but, for brevity, we leave further low hanging fruit as an exercise for the reader.
-
-It is interesting to consider if addition or temporary logical-AND gates can be done with still fewer T gates.
+It is interesting to consider if addition or temporary logical-ANDs can be done with still fewer T gates.
 \autoref{fig:lower-bound-logical-AND} presents a simple proof that the T-count of the temporary logical-AND is at least 2.
-Regardless of whether the true lower bound is 4, we expect that there are special cases with lower average T-counts.
+Although we suspect the true lower bound is 4, we also expect that there are special cases with lower T-counts.
 For example, if a task involves repeating an action several times (e.g.
@@
 
-Regardless of whether there will be further T-count improvements for operations as basic as matched Toffoli gates, it is clear that many other T-count optimizations are waiting to be found.
+Regardless of whether there will be further T-count improvements for operations as basic as Toffoli gates appearing in pairs, it is clear that many other T-count optimizations are waiting to be found.
 Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions, and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.

### 2017-09-12 paper.tex
Polishing pass before sending out to team for comments

@@
 
-Temporary logical-ANDs are a generally useful tool when optimizing T-count.
+Temporary logical-ANDs are a generally useful tool when optimizing T-counts.
 They can be applied to integer arithmetic, modular arithmetic, rotation synthesis, the quantum Fourier transform, Shor's algorithm, Grover oracles, and many other circuits.
@@
 Our paper is divided into four sections.
-We motivate the need to optimize T-counts and discuss existing work in \autoref{sec:introduction}.
+We discuss existing work and motivate the need to optimize T-counts in \autoref{sec:introduction}.
 \autoref{sec:construction} presents our $4n + O(1)$ T-count adder and explains its pieces, including the temporary logical-AND.
-In \autoref{sec:other-applications}, we justify the assertion that our constructions represent widely applicable optimizations by giving examples of quantum tasks and cost projections that are improved by efficient adders and temporary logical-ANDs.
+In \autoref{sec:other-applications} we justify the assertion that our constructions represent widely applicable optimizations by giving examples of quantum tasks and cost projections that are improved by efficient adders and temporary logical-ANDs.
 Finally, \autoref{sec:conclusion} summarizes our contributions and discusses future work.
@@
 This asymmetry is due to the fact that measurement is not reversible.
-The uncomputation uses measurement in a way that the computation can not.
+The uncomputation uses measurement in a way that the computation cannot.
 
@@
 Because Shor's algorithm is dominated by the cost of controlled additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $\sim 2.7$ and $1$ respectively.
-This reduces the estimates to 9 hours and $8 \cdot 10^{11}$ distilled $|A\rangle$ states.
+This reduces the time estimate to 9 hours and the count estimate to $8 \cdot 10^{11}$ distilled $|A\rangle$ states.
 
@@
 This takes $4n + O(\lg n)$ T gates to do.
-Then, for each position $k$ in the Hamming weight register, synthesize and apply the operation $R_Z(\theta \cdot 2^k)$ to the register qubit at that position.
-This uses a number of T gates proportional to some polynomial of the desired bit precision, times the number of bits in the output register.
-The output register is exponentially smaller than $n$, so this cost is negligible for large $n$.
+Then, for each position $p$ in the Hamming weight register, synthesize and apply the operation $R_Z(\theta \cdot 2^p)$ to the register qubit at that position.
+This uses a number of T gates proportional to some polynomial of the desired bit precision, times the number of bits in the Hamming weight register.
+The Hamming weight register is exponentially smaller than $n$, so this cost is negligible for large $n$.
 Finally, uncompute the Hamming weight register (using no T gates).
@@
 Another quantum operation that can be implemented via an adder is the $n$-qubit phase gradient operation $\text{Grad}_n = \sum_{k=0}^{2^n-1} e^{2 i \pi k / 2^n} |k\rangle \langle k|$.
-Normally this operation would be implemented by separately applying the operation $R_z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
+Normally this operation would be implemented by separately applying the operation $R_Z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
 However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{-2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition.
 Add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
-This construction performs phase gradients using $4n + O(1)$ T gates, which is apparently independent of $\epsilon$.
-However, there are in fact three ways in which the construction's cost does depend on $\epsilon$.
+This construction performs phase gradients using $4n + O(1)$ T gates.
+This T-count is interesting because it appears to be independent of $\epsilon$ despite the phase gradient operation involving arbitrarily small rotations.
+However, there are three ways in which our phase gradient construction's cost does depend on $\epsilon$.
 First, $\epsilon$ still bounds the minimum quality of the $|A\rangle$ states powering the T gates.
@@
 
-For example, a simple way to construct Grover oracles is to translate a classical predicate into a reversible circuit made out of Toffoli gates.
-This classical reversible circuit is then used to compute an output bit, so that a Z gate can be applied to the output bit before the circuit is run in reverse to uncompute the output.
+For example, a simple way to construct Grover oracles starts by translating a classical predicate into a classical reversible circuit made with Toffoli gates.
+This reversible circuit is then used on a quantum computer to compute an output qubit, and a Z gate is applied to the output qubit before the circuit is run in reverse to uncompute the output qubit.
 Every Toffoli gate generated by this construction is part of a compute/uncompute pair (each Toffoli in the computation will match a Toffoli in the uncomputation).
-Using temporary logical-ANDs halves the T-count of this approach to constructing Grover oracles.
+Using temporary logical-ANDs, instead of individually translating each Toffoli gate into T gates, halves the T-count of this approach to constructing Grover oracles.
 
@@
 
-Regardless of whether there will be further T-count improvements for operations as basic as Toffoli gates appearing in pairs, it is clear that many other T-count optimizations are waiting to be found.
+Regardless of whether there will be further T-count improvements for operations as basic as Toffoli gates appearing in pairs, it is clear that many other kinds of T-count optimizations are waiting to be found.
 Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions, and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
@@
 
-We thank Austin Fowler for sampling the opinions of other researchers, assistance locating references, and comments that greatly improved our paper.
+We thank Austin Fowler for sampling the opinions of other researchers, assistance locating references, and comments that significantly improved this paper.
+We thank Dave Bacon for comments that significantly improved this paper.
 

### 2017-09-12 paper.tex
Changes based on Matt's comments

@@
 
-Our adder uses temporary logical-AND operations, which we draw as wires emerging out of a pair of controls then later merging into an equivalent pair of controls.
+Our adder uses temporary logical-AND operations, which we draw as wires emerging out of a pair of controls then later merging into the same pair of controls.
 \autoref{fig:temporary-logical-AND} shows how we compute the logical-AND of the two controls (which we draw as a wire emerging from two controls then heading rightward), and also the corresponding uncomputation (which we draw as a wire turning upward into two controls).
@@
 The temporary logical-AND is a basic circuit construction widget that can be used in many circuits, and a good example of measurement breaking the symmetry between computation and uncomputation (allowing one to be more efficient than the other).
-We covered a few examples of optimizing T-counts with our constructions but, for brevity, we leave further low hanging fruit as an exercise for the reader.
+We demonstrated how to optimize the T-count of a few tasks using our adder and the temporary logical-AND, but there are many other low hanging applications.
 

### 2017-09-19 paper.tex
Polishing for submission to arxiv

@@
   \caption{
-	Adder circuit building-block with a T-count of 4.
+	Our adder circuit building-block, with a T-count of 4.
 	A variant of the Cuccaro adder building-block \citep{Cuccaro2004}.
@@
 First, $\epsilon$ still bounds the minimum quality of the $|A\rangle$ states powering the T gates.
-Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\lg \lg \epsilon$.
+Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\lg \lg \frac{1}{\epsilon}$.
 Third, initializing the reusable phase gradient register has a one-time cost of $O(n_{\text{max}} \cdot \text{poly}(\lg \frac{1}{\epsilon}))$ T gates.
@@
 The temporary logical-AND is useful for optimizing an even wider variety of circuits than addition is.
-Whenever Toffoli gates appear in compute/uncompute pairs, and intermediate operations are not sensitive to phase errors on the controls of the Toffoli gate (e.g.
+Whenever Toffoli gates appear in compute/uncompute pairs, and intermediate operations are not sensitive to phase errors on the controls of the Toffoli gate (i.e.
 the condition in \autoref{fig:pair-definition} is satisfied), it is possible to save 4 T gates by replacing the pair of Toffoli gates with a temporary logical-AND.
@@
 
-We thank Austin Fowler for sampling the opinions of other researchers, assistance locating references, and comments that significantly improved this paper.
+We thank Austin Fowler for sampling the opinions of other researchers, for assistance locating references, and for comments that significantly improved this paper.
 We thank Dave Bacon for comments that significantly improved this paper.

### 2017-09-26 paper.tex
Incorporating comments from Dmitri Maslov - Dmitri objected to the email thread with Austin being construed as stating a conjecture. Removed. - Added n-control NOT as an example.

@@
 
-The leading factor of 8 in the T-count of an $n$-bit adder has stood for over a decade \citep{Barenco1995, Cuccaro2004}, and has been conjectured to be optimal \cite{AustinDiscussionsAndEmails2017}.
-We disprove this conjecture with a construction based on the temporary logical-AND, which combines ideas from both the matched-phase-error construction and the ancilla-and-fix-up construction.
+The leading factor of 8 in the T-count of an $n$-bit adder has stood for over a decade \citep{Barenco1995, Cuccaro2004, AustinDiscussionsAndEmails2017}.
+We improve the leading factor with a construction based on the temporary logical-AND, which combines ideas from both the matched-phase-error construction and the ancilla-and-fix-up construction.
 The temporary logical-AND halves the number of T gates needed to perform Toffoli gates that appear in compute/uncompute pairs, and thereby halves the T-count of the Cuccaro adder from $8n + O(1)$ to $4n + O(1)$.
@@
 
-Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of abstract mathematics.
+As another example, note that the temporary logical-AND can perform NOT gates with many controls by iteratively combining the controls down to a single representative ancilla.
+The representative ancilla is then used to control a CNOT onto the target qubit.
+Overall this takes $4n-4$ T gates, equivalent to previous work \cite{Jones2013}.
+As when replacing Toffoli gates in compute/uncompute pairs with a temporary logical-AND, the ancilla representing the $n$ controls should be kept and used multiple times whenever possible (instead of being uncomputed and recomputed).
+
+Finally, we note that Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of abstract mathematics.
 For example, we used temporary logical-ANDs to cut the projected T-count of a step in a simulation algorithm under development by nearly 30\% \citep{RyanEmails2017}.
@@
 We thank Austin Fowler for sampling the opinions of other researchers, for assistance locating references, and for comments that significantly improved this paper.
-We thank Dave Bacon for comments that significantly improved this paper.
+We thank Dave Bacon and Dmitri Maslov for comments that significantly improved this paper.
 

### 2017-09-26 paper.tex
First pass of modifications to meet nature quantum information - Separated latex used to generate figures into separate files - Changed section names - Added required sections

@@
 We do so via a ``temporary logical-AND" construction, which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
-
 Temporary logical-ANDs are a generally useful tool when optimizing T-counts.
@@
 Because T gates dominate the cost of quantum computation based on the surface code, and temporary logical-ANDs are widely applicable, our constructions represent a significant reduction in projected costs of quantum computation.
-
 We also present an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until it is later uncomputed without using T gates, and discuss some other constructions whose T-count is improved by the temporary logical-AND.
@@
 
-\section{Introduction}
-\label{sec:introduction}
+\section*{Introduction}
 
@@
 \begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &i_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\ctrl{1}&\qw &i_0     &&&&&\\
-      &t_0 &&\ctrl{1} &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw &&&(t+i)_0 &&&\\
-      &    &&         &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\ctrl{1}&\qw     &        &    &&&        &&&\\
-      &i_1 &&\qw      &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\ctrl{1}&\qw &i_1     &&&&&\\
-      &t_1 &&\qw      &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\targ   &\qw &&&(t+i)_1 &&&\\
-      &    &&         &        &        &\targ   &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw     &\ctrl{1}&\targ   &\qw     &        &        &        &    &&&        &&&\\
-      &i_2 &&\qw      &\qw     &\qw     &\qw     &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_2     &&&&&\\
-      &t_2 &&\qw      &\qw     &\qw     &\qw     &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_2 &&&\\
-      &    &&         &        &        &        &        &        &\targ   &\ctrl{2}&\qw     &\ctrl{3}&\qw     &\ctrl{3}&\qw     &\ctrl{1}&\targ   &\qw     &        &        &        &        &        &        &    &&&        &&&\\
-      &i_3 &&\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\ctrl{1}&\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_3     &&&&&\\
-      &t_3 &&\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\ctrl{1}&\qw     &\qw     &\qw     &\ctrl{1}&\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_3 &&&\\
-      &    &&         &        &        &        &        &        &        &        &        &\targ   &\ctrl{2}&\targ   &\qw     &        &        &        &        &        &        &        &        &        &    &&&        &&&\\
-      &i_4 &&\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\ctrl{1}&\qw &i_4     &&&&&\\
-      &t_4 &&\qw      &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\qw     &\targ   &\qw &&&(t+i)_4 &&&\\
-    }
-  }
+  \input{figures/adder.qcircuit}
   \caption{
@@
 
-Our paper is divided into four sections.
-We discuss existing work and motivate the need to optimize T-counts in \autoref{sec:introduction}.
-\autoref{sec:construction} presents our $4n + O(1)$ T-count adder and explains its pieces, including the temporary logical-AND.
-In \autoref{sec:other-applications} we justify the assertion that our constructions represent widely applicable optimizations by giving examples of quantum tasks and cost projections that are improved by efficient adders and temporary logical-ANDs.
-Finally, \autoref{sec:conclusion} summarizes our contributions and discusses future work.
-
-\section{Adder Construction}
-\label{sec:construction}
+\section*{Results}
 
@@
 \begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &c_k &&\ctrl{2}&\qw     &\ctrl{3}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{3}&\qw     &\ctrl{1}&\qw     &\qw &c_k     &&&&& \\
-      &i_k &&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\targ   &\ctrl{1}&\qw &i_k     &&&&& \\
-      &t_k &&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\qw     &\targ   &\qw &&&(t+i)_k &&& \\
-      &    &&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &    &&&        &&& \\
-    }
-  }
+  \input{figures/adder-building-block.qcircuit}
   \caption{
@@
 
-\section{Other Applications}
-\label{sec:other-applications}
-
 \autoref{fig:variations-on-full-adder-block} shows the building-blocks for two variations on our adder: a controlled adder and a temporary adder.
@@
 \begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &x &&\ctrl{1} &\qw &x  && &&          &&\ctrl{2}&\qw     &\targ    &\qTi &\targ     &\qw &\qw &\qw &\\
-      &y &&\ctrl{1} &\qw &y  &&=&&          &&\qw     &\ctrl{1}&\targ    &\qTi &\targ     &\qw &\qw &\qw &\\
-      &  &&         &\qw &xy && &&|A\rangle &&\targ   &\targ   &\ctrl{-2}&\qT  &\ctrl{-2} &\qH &\qS &\qw &\\
-    }
-  }
-  \vspace*{0.25cm} \\
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &x  &&\ctrl{1} &\qw &x && &&&\qw &\qw    &\ctrl{1}         &\qw &&&&&&&&&&\\
-      &y  &&\ctrl{1} &\qw &y &&=&&&\qw &\qw    &\gate{Z}         &\qw &&&&&&&&&&\\
-      &xy &&\qw      &    &  && &&&\qH &\meter &\cw \cwx \bullet &    &&&&&&&&&&\\
-    }
-  }
+  \input{figures/temporary-logical-and.qcircuit}
   \caption{
@@
 \begin{figure}
-  (a) \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &&\text{control} &&&&\qw     &\qw     &\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\qw     &\ctrl{2}&\qw     &\ctrl{2}&\qw     &\qw &&&\text{control}               &&&&&&\\
-      &&&c_k            &&&\ctrl{3}&\qw     &\ctrl{4}&\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\ctrl{4}&\qw     &\qw     &\qw     &\qw     &\ctrl{3}&\qw &c_k                          &&&&&&&&\\
-      &&&i_k            &&&\targ   &\ctrl{2}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{2}&\ctrl{1}&\qw     &\ctrl{1}&\targ   &\qw &i_k                          &&&&&&&&\\
-      &&               &&&&        &        &        &    &    &       &    &    &    &       &    &    &    &       &    &    &        &        &        &\ctrl{1}&\qw     &        &    &    &&&&&&&&\\
-      &&&t_k            &&&\targ   &\ctrl{1}&\qw     &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw &\qw    &\qw &\qw &\qw     &\ctrl{1}&\qw     &\targ   &\qw     &\targ   &\qw &&&&&(t+i \cdot \text{control})_k &&&&&\\
-      &&               &&&&        &        &\targ   &\qw &    &c_{k+1}&    &    &    &\ldots &    &    &    &c_{k+1}&    &    &\targ   &\qw     &        &        &        &        &    &    &&&&&&&&\\
-    }
-  }
-  \vspace*{0.25cm} \\
-  (b) \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &a &&\ctrl{2}&\qw     &\qw     &\ctrl{3}&\qw &a          &&&&&&&&&&&&&&&&&&&&&&\\
-      &b &&\targ   &\ctrl{1}&\ctrl{1}&\targ   &\qw &b          &&&&&&&&&&&&&&&&&&&&&&\\
-      &c &&\targ   &\ctrl{1}&\targ   &\qw     &\qw &&&&(a+b+c)_0  &&&&&&&&&&&&&&&&&&&\\
-      &  &&        &        &\qw     &\targ   &\qw &&&&(a+b+c)_1  &&&&&&&&&&&&&&&&&&&\\
-    }
-  }
+  \input{figures/other-adder-building-blocks.qcircuit}
   \caption{
@@
 \begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &&               &&&&\ustick{n}&\diagup\qw &\multigate{3}{U}&\qw     &\qw &          &&            &&\ustick{n}&\diagup\qw &\qw      &\multigate{3}{U}&\qw      &\qw     &\qw &\\
-      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&               &&&&          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &          &&            &&          &\ctrl{1}   &\qw      &\ghost{U}       &\qw      &\ctrl{1}&\qw &\\
-      &          &\ctrl{1}   &\ghost{U}       &\ctrl{1}&\qw &&\stackrel{?}{=}&&&&          &\ctrl{2}   &\ghost{U}       &\ctrl{2}&\qw &          &&\Rightarrow &&          &\ctrl{2}   &\qw      &\ghost{U}       &\qw      &\ctrl{2}&\qw &\\
-      &          &\targ      &\ghost{U}       &\targ   &\qw &&               &&&&          &\targ      &\ghost{U}       &\targ   &\qw &          &&            &&          &\qw        &\targ    &\ghost{U}       &\targ    &\qw     &\qw &\\
-      &          &           &                &        &    &&               &&&&\qO       &\targ      &\qw             &\targ   &\qw &|0\rangle &&            &&          &           &\ctrl{-1}&\qw             &\ctrl{-1}&\qw     &    &\\
-    }
-  }
+  \input{figures/paired-toffoli-replacement-rule.qcircuit}
   \caption{
@@
 
-\section{Conclusion}
-\label{sec:conclusion}
+\section*{Discussion}
 
@@
 \begin{figure}
-  \resizebox{\linewidth}{!}{
-    \Qcircuit @R=0.7em @C=0.7em {
-      &\qT &\qw && &&&\ctrl{2}&\qw     &\ctrl{3}&\qw     &\qw &\qw     &\ctrl{3}&\qw     &\ctrl{2}&\qw &\\
-      &\qT &\qw &&=&&&\targ   &\ctrl{1}&\qw     &\ctrl{1}&\qw &\ctrl{1}&\qw     &\ctrl{1}&\targ   &\qw &\\
-      &\qT &\qw && &&&\targ   &\ctrl{1}&\targ   &\targ   &\qT &\targ   &\targ   &\ctrl{1}&\targ   &\qw &\\
-      &    &    && &&&        &        &\targ   &\qw     &\qS &\qw     &\targ   &\qw     &        &    &\\
-    }
-  }
+  \input{figures/synthesize-t-gates-from-temporary-logical-and.qcircuit}
   \caption{
@@
 
-\section{Acknowledgements}
+\section*{Methods}
+
+The described circuits can be verified with a quantum circuit simulator, such as ProjectQ or LIQUID or Quirk.
+
+\section*{Acknowledgements}
 
@@
 We thank Dave Bacon and Dmitri Maslov for comments that significantly improved this paper.
+
+\section*{Competing Interests}
+
+No competing interests.
+
+\section*{Contributions}
+
+Craig Gidney discovered and refined the temporary logical-AND construction, applied it to adders and the other discussed circuits, and is the sole author of the paper.
+
+\section*{Funding}
+
+Work was done while employed at Google.
 

### 2017-10-04 paper.tex
Generating figures as separate PDFs

@@
 \hypersetup{colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue}
-
-\input{Qcircuit}
-\newcommand{\qH}{\gate{H}}
-\newcommand{\qT}{\gate{T}}
-\newcommand{\qTi}{\gate{T^\dagger}}
-\newcommand{\qS}{\gate{S}}
-\newcommand{\qSi}{\gate{S^\dagger}}
-\newcommand{\qA}{\lstick{|A\rangle}}
-\newcommand{\qO}{\lstick{|0\rangle}}
 
@@
 \begin{figure}
-  \input{figures/adder.qcircuit}
+  \includegraphics[width=\linewidth]{figures/adder.pdf}
   \caption{
@@
 	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
-	See \autoref{fig:full-adder-block} for the adder building-block and \autoref{fig:temporary-logical-AND} for the logical-AND computation and uncomputation circuits.
+	See \autoref{fig:adder-building-block} for the adder building-block and \autoref{fig:temporary-logical-AND} for the logical-AND computation and uncomputation circuits.
   }
-  \label{fig:multi-bit-adder-example}
+  \label{fig:adder}
 \end{figure}
@@
 
-In \autoref{fig:multi-bit-adder-example}, we present a 5-bit adder with a T-count of 16.
+In \autoref{fig:adder}, we present a 5-bit adder with a T-count of 16.
 It performs 4 temporary logical-ANDs, each with a T-count of 4.
@@
 
-The building block of our adder is shown in \autoref{fig:full-adder-block}.
+The building block of our adder is shown in \autoref{fig:adder-building-block}.
 We construct $n$-bit adders by nesting $n$ copies of the building block inside of each other.
@@
 \begin{figure}
-  \input{figures/adder-building-block.qcircuit}
+  \includegraphics[width=\linewidth]{figures/adder-building-block.pdf}
   \caption{
@@
   }
-  \label{fig:full-adder-block}
+  \label{fig:adder-building-block}
 \end{figure}
@@
 
-\autoref{fig:variations-on-full-adder-block} shows the building-blocks for two variations on our adder: a controlled adder and a temporary adder.
+\autoref{fig:other-adder-building-blocks} shows the building-blocks for two variations on our adder: a controlled adder and a temporary adder.
 
@@
 \begin{figure}
-  \input{figures/temporary-logical-and.qcircuit}
+  \includegraphics[width=\linewidth]{figures/temporary-logical-and.pdf}
   \caption{
@@
 \begin{figure}
-  \input{figures/other-adder-building-blocks.qcircuit}
+  \includegraphics[width=\linewidth]{figures/other-adder-building-blocks.pdf}
   \caption{
@@
   }
-  \label{fig:variations-on-full-adder-block}
+  \label{fig:other-adder-building-blocks}
 \end{figure}
@@
 Whenever Toffoli gates appear in compute/uncompute pairs, and intermediate operations are not sensitive to phase errors on the controls of the Toffoli gate (i.e.
-the condition in \autoref{fig:pair-definition} is satisfied), it is possible to save 4 T gates by replacing the pair of Toffoli gates with a temporary logical-AND.
+the condition in \autoref{fig:paired-toffoli-replacement-rule} is satisfied), it is possible to save 4 T gates by replacing the pair of Toffoli gates with a temporary logical-AND.
 
 \begin{figure}
-  \input{figures/paired-toffoli-replacement-rule.qcircuit}
+  \includegraphics[width=\linewidth]{figures/paired-toffoli-replacement-rule.pdf}
   \caption{
@@
   }
-  \label{fig:pair-definition}
+  \label{fig:paired-toffoli-replacement-rule}
 \end{figure}
@@
 It is interesting to consider if addition or temporary logical-ANDs can be done with still fewer T gates.
-\autoref{fig:lower-bound-logical-AND} presents a simple proof that the T-count of the temporary logical-AND is at least 2.
+\autoref{fig:synthesize-t-gates-from-temporary-logical-and} presents a simple proof that the T-count of the temporary logical-AND is at least 2.
 Although we suspect the true lower bound is 4, we also expect that there are special cases with lower T-counts.
@@
 \begin{figure}
-  \input{figures/synthesize-t-gates-from-temporary-logical-and.qcircuit}
+  \includegraphics[width=\linewidth]{figures/synthesize-t-gates-from-temporary-logical-and.pdf}
   \caption{
@@
   }
-  \label{fig:lower-bound-logical-AND}
+  \label{fig:synthesize-t-gates-from-temporary-logical-and}
 \end{figure}
@@
 Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions, and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
-
-\section*{Methods}
-
-The described circuits can be verified with a quantum circuit simulator, such as ProjectQ or LIQUID or Quirk.
 
@@
 
-\section*{Competing Interests}
-
-No competing interests.
-
-\section*{Contributions}
-
-Craig Gidney discovered and refined the temporary logical-AND construction, applied it to adders and the other discussed circuits, and is the sole author of the paper.
-
-\section*{Funding}
-
-Work was done while employed at Google.
-
 \bibliographystyle{plainnat}

### 2017-10-05 paper.tex
Move figures to end of paper in own section
Abbreviated bibliography and dropped figure files to be inline

@@
 \usepackage{graphicx}
+\usepackage{float}
 \usepackage[pdfpagelabels,pdftex,bookmarks,breaklinks]{hyperref}
@@
 
+% [figure 1 "adder" was here]
+
 The Cuccaro adder uses $2n + O(1)$ Toffoli gates \citep{Cuccaro2004}.
@@
 The temporary logical-AND halves the number of T gates needed to perform Toffoli gates that appear in compute/uncompute pairs, and thereby halves the T-count of the Cuccaro adder from $8n + O(1)$ to $4n + O(1)$.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/adder.pdf}
-  \caption{
-	A 5-bit adder with T-count of 16.
-	Uses Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures requiring no T gates.
-	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
-	See \autoref{fig:adder-building-block} for the adder building-block and \autoref{fig:temporary-logical-AND} for the logical-AND computation and uncomputation circuits.
-  }
-  \label{fig:adder}
-\end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/adder-building-block.pdf}
-  \caption{
-	Our adder circuit building-block, with a T-count of 4.
-	A variant of the Cuccaro adder building-block \citep{Cuccaro2004}.
-  }
-  \label{fig:adder-building-block}
-\end{figure}
+% [figure 2 "adder-building-block" was here]
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/temporary-logical-and.pdf}
-  \caption{
-	How to compute and uncompute the logical-AND of two qubits.
-	The computation circuit (top) has a T-count of 4 and a T-depth of 1.
-	Note that the $|A\rangle$ state input contributes to the T-count, because $|A\rangle$ states are the resource used to perform T gates.
-	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero.
-    \\
-    An alternative uncomputation construction is to simply do the reverse of the computation circuit.
-    This alternative approach has a net T-count of 2 (because an $|A\rangle$ state is recovered).
-    The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.
-  }
-  \label{fig:temporary-logical-AND}
-\end{figure}
+% [figure 3 "temporary-logical-and" was here]
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/other-adder-building-blocks.pdf}
-  \caption{
-	Variations on our adder construction.
-	(a) is a controlled-adder building-block with a T-count of 8.
-	The addition will only occur when the control is on.
-	(b) is a temporary-adder building-block with a T-count of 4.
-	Has the same T-count as a normal addition, but makes the result available before any uncomputation steps are needed.
-  }
-  \label{fig:other-adder-building-blocks}
-\end{figure}
+% [figure 4 "other-adder-building-blocks" was here]
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/paired-toffoli-replacement-rule.pdf}
-  \caption{
-	A sufficient condition for replacing a pair of Toffoli gates with a temporary logical-AND, saving 4 T gates.\\
-	1) The later Toffoli gate must be uncomputing the earlier Toffoli gate.\\
-	2) Intermediate operations must not be sensitive to the presence of the entangled ancilla.
-  }
-  \label{fig:paired-toffoli-replacement-rule}
-\end{figure}
+% [figure 5 "paired-toffoli-replacement-rule" was here]
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/synthesize-t-gates-from-temporary-logical-and.pdf}
-  \caption{
-	Three T gates can be synthesized by performing a temporary logical-AND, one T gate, and various Clifford operations.
-	Clifford operations can't produce T gates, so the extra T gates must be coming from the temporary logical-AND.
-	This proves that performing the temporary logical-AND requires at least two T gates.
-  }
-  \label{fig:synthesize-t-gates-from-temporary-logical-and}
-\end{figure}
+% [figure 6 "synthesize-t-gates-from-temporary-logical-and" was here]
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{npj-quantum-information}
 \bibliography{citations}
 
+\section*{Figures}
+
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{adder.pdf}
+  \caption{
+	A 5-bit adder with T-count of 16.
+	Uses Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures requiring no T gates.
+	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
+	See \autoref{fig:adder-building-block} for the adder building-block and \autoref{fig:temporary-logical-AND} for the logical-AND computation and uncomputation circuits.
+  }
+  \label{fig:adder}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{adder-building-block.pdf}
+  \caption{
+	Our adder circuit building-block, with a T-count of 4.
+	A variant of the Cuccaro adder building-block \citep{Cuccaro2004}.
+  }
+  \label{fig:adder-building-block}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{temporary-logical-and.pdf}
+  \caption{
+	How to compute and uncompute the logical-AND of two qubits.
+	The computation circuit (top) has a T-count of 4 and a T-depth of 1.
+	Note that the $|A\rangle$ state input contributes to the T-count, because $|A\rangle$ states are the resource used to perform T gates.
+	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero.
+    \\
+    An alternative uncomputation construction is to simply do the reverse of the computation circuit.
+    This alternative approach has a net T-count of 2 (because an $|A\rangle$ state is recovered).
+    The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.
+  }
+  \label{fig:temporary-logical-AND}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{other-adder-building-blocks.pdf}
+  \caption{
+	Variations on our adder construction.
+	(a) is a controlled-adder building-block with a T-count of 8.
+	The addition will only occur when the control is on.
+	(b) is a temporary-adder building-block with a T-count of 4.
+	Has the same T-count as a normal addition, but makes the result available before any uncomputation steps are needed.
+  }
+  \label{fig:other-adder-building-blocks}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{paired-toffoli-replacement-rule.pdf}
+  \caption{
+	A sufficient condition for replacing a pair of Toffoli gates with a temporary logical-AND, saving 4 T gates.\\
+	1) The later Toffoli gate must be uncomputing the earlier Toffoli gate.\\
+	2) Intermediate operations must not be sensitive to the presence of the entangled ancilla.
+  }
+  \label{fig:paired-toffoli-replacement-rule}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{synthesize-t-gates-from-temporary-logical-and.pdf}
+  \caption{
+	Three T gates can be synthesized by performing a temporary logical-AND, one T gate, and various Clifford operations.
+	Clifford operations can't produce T gates, so the extra T gates must be coming from the temporary logical-AND.
+	This proves that performing the temporary logical-AND requires at least two T gates.
+  }
+  \label{fig:synthesize-t-gates-from-temporary-logical-and}
+\end{figure}
+
 \end{document}

### 2017-11-21 paper.tex
Restoring trivial sections.
Respond to reviewer comments from the npj rejection for resubmit.

@@
 We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \citep{Amy2013, Cuccaro2004, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
-We do so via a ``temporary logical-AND" construction, which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
+We do so via a ``temporary logical-AND" construction (a reframing of a result by Jones \cite{Jones2013}) which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
 Temporary logical-ANDs are a generally useful tool when optimizing T-counts.
@@
 Instead, $T$ gates are performed by distilling and consuming $|A\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle)$ states.
-Consuming an $|A\rangle$ state to perform a T gate is simple, but distilling $|A\rangle$ states has significant cost.
-Because T gates are so expensive for the surface code, and the surface code is a likely component of future quantum computers, it is important to consider and optimize the number of T gates used by quantum circuits.
+The T gate is performed via a controlled-NOT from the target qubit into the $|A\rangle$ ancilla, followed by a classical feedback step that measures the ancilla to determine if an $S$ gate is applied to the target.
+Unfortunately, distilling the $|A\rangle$ state that will ultimately be passed into this process has significant cost.
+The cost of distillation is high enough that error corrected quantum computations are expected to bottleneck waiting for $|A\rangle$ states to be produced, so that their runtime is dominated by the number of T gates (as opposed to more traditional measures of circuit depth).
+
+Because the surface code is a likely component of future quantum computers, and the runtime of computations within the surface code will be dominated in practice by the number of T gates, it is important to optimize the number of T gates used by quantum circuits.
 Optimizing the T-count of basic elements of quantum circuits, such as the construction of adders and Toffoli gates, is particularly important because any improvement is widely applicable.
@@
 The leading factor of 8 in the T-count of an $n$-bit adder has stood for over a decade \citep{Barenco1995, Cuccaro2004, AustinDiscussionsAndEmails2017}.
-We improve the leading factor with a construction based on the temporary logical-AND, which combines ideas from both the matched-phase-error construction and the ancilla-and-fix-up construction.
-The temporary logical-AND halves the number of T gates needed to perform Toffoli gates that appear in compute/uncompute pairs, and thereby halves the T-count of the Cuccaro adder from $8n + O(1)$ to $4n + O(1)$.
+We improve the leading factor with a construction based on the temporary logical-AND, combining ideas from both the matched-phase-error construction and the ancilla-and-fix-up construction.
+We thereby halve the number of T gates needed to perform Toffoli gates that appear in compute/uncompute pairs and thus halve the T-count of Cuccaro-style adders from $8n + O(1)$ to $4n + O(1)$.
+
+Although the title of this paper focuses on addition, we consider our main {\em conceptual} contribution to be reframing the ancilla-and-fix-up construction from \cite{Jones2013} into the temporary logical-AND operation.
+This conceptualization makes it clear that the construction is more widely applicable than previously realized.
+Anytime operations share controls, even if those operations are far apart, a temporary logical-AND allows the controls to be combined once instead of once per operation (as long as intermediate operations are not sensitive to phase error in the controls).
+Addition is just one example of an operation where this is beneficial.
+We also discuss other examples.
 
@@
 Our adder uses temporary logical-AND operations, which we draw as wires emerging out of a pair of controls then later merging into the same pair of controls.
-\autoref{fig:temporary-logical-AND} shows how we compute the logical-AND of the two controls (which we draw as a wire emerging from two controls then heading rightward), and also the corresponding uncomputation (which we draw as a wire turning upward into two controls).
+\autoref{fig:temporary-logical-AND} shows how we compute the logical-AND of the two controls, and also the corresponding uncomputation.
 
 Computing the temporary logical-AND has a T-count of 4, but uncomputing it has a T-count of 0.
-This asymmetry is due to the fact that measurement is not reversible.
+This surprising asymmetry is due to the fact that measurement is not reversible.
 The uncomputation uses measurement in a way that the computation cannot.
@@
 First, $\epsilon$ still bounds the minimum quality of the $|A\rangle$ states powering the T gates.
-Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\lg \lg \frac{1}{\epsilon}$.
+Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\lg \frac{1}{\epsilon}$.
 Third, initializing the reusable phase gradient register has a one-time cost of $O(n_{\text{max}} \cdot \text{poly}(\lg \frac{1}{\epsilon}))$ T gates.
@@
 The representative ancilla is then used to control a CNOT onto the target qubit.
-Overall this takes $4n-4$ T gates, equivalent to previous work \cite{Jones2013}.
+This construction takes $4n-4$ T gates to perform a NOT with $n$ controls, but is simply a reframing of the nesting construction from \cite{Jones2013}.
 As when replacing Toffoli gates in compute/uncompute pairs with a temporary logical-AND, the ancilla representing the $n$ controls should be kept and used multiple times whenever possible (instead of being uncomputed and recomputed).
 
-Finally, we note that Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of abstract mathematics.
+Yet another example of a kind of circuit that benefits from the temporary logical-AND is {\em low-depth} adder circuits.
+For example, the Toffoli gates produced by the $P$ and $P^{-1}$ rounds in Draper et al's logarithmic-depth adder \cite{Draper2004} form compute/uncompute pairs that can be replaced by temporary logical-ANDs.
+Although using temporary logical-ANDs improves the T-count of logarithmic-depth adders, the resulting T-count is still worse than the T-count of linear-depth adders.
+This means that in practice, given our expectation that error corrected quantum computation will bottleneck waiting on $|A\rangle$ states to perform T gates, the linear-depth adders will often execute in less time than the logarithmic-depth adders.
+
+Finally, we note that Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of mathematics.
 For example, we used temporary logical-ANDs to cut the projected T-count of a step in a simulation algorithm under development by nearly 30\% \citep{RyanEmails2017}.
@@
 The temporary logical-AND is a basic circuit construction widget that can be used in many circuits, and a good example of measurement breaking the symmetry between computation and uncomputation (allowing one to be more efficient than the other).
-We demonstrated how to optimize the T-count of a few tasks using our adder and the temporary logical-AND, but there are many other low hanging applications.
+We demonstrated how to optimize the T-count of a few tasks using our adder and the temporary logical-AND, but there are many other low hanging applications (e.g.
+converting between binary and unary, temporary sorting, applying an operation to a qubit indexed by a register in superposition, parameterized bit-rotation of a qubit register, computing the greatest common divisor of two values, etc).
 
@@
 
+\section*{Methods}
+
+Not applicable. The circuits can be verified with any quantum circuit simulator.
+
 \section*{Acknowledgements}
@@
 
+\section*{Competing Interests}
+
+None.
+
+\section*{Contributions}
+
+Craig Gidney: sole author.
+
+\section*{Funding}
+
+None. Work was done while employed at Google.
+
 \bibliographystyle{npj-quantum-information}
 \bibliography{citations}
+
+\section*{Figure legends}
+
+None.
+
+\section*{Tables}
+
+None.
 
@@
 
-\begin{figure}
+\begin{figure}[H]
   \includegraphics[width=\linewidth]{other-adder-building-blocks.pdf}
@@
 
-\begin{figure}
+\begin{figure}[H]
   \includegraphics[width=\linewidth]{paired-toffoli-replacement-rule.pdf}
@@
 
-\begin{figure}
+\vspace{5cm}
+
+\begin{figure}[H]
   \includegraphics[width=\linewidth]{synthesize-t-gates-from-temporary-logical-and.pdf}

### 2018-01-19 paper.tex
Polish the abstract

@@
 We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \citep{Amy2013, Cuccaro2004, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
-We do so via a ``temporary logical-AND" construction (a reframing of a result by Jones \cite{Jones2013}) which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
-Temporary logical-ANDs are a generally useful tool when optimizing T-counts.
-They can be applied to integer arithmetic, modular arithmetic, rotation synthesis, the quantum Fourier transform, Shor's algorithm, Grover oracles, and many other circuits.
-Because T gates dominate the cost of quantum computation based on the surface code, and temporary logical-ANDs are widely applicable, our constructions represent a significant reduction in projected costs of quantum computation.
-We also present an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until it is later uncomputed without using T gates, and discuss some other constructions whose T-count is improved by the temporary logical-AND.
+We do so via a ``temporary logical-AND" construction which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
+This construction is equivalent to one by Jones \cite{Jones2013}, except that our framing makes it clear that the technique is far more widely applicable than previously realized.
+Temporary logical-ANDs can be applied to integer arithmetic, modular arithmetic, rotation synthesis, the quantum Fourier transform, Shor's algorithm, Grover oracles, and many other circuits.
+Because T gates dominate the cost of quantum computation based on the surface code, and temporary logical-ANDs are widely applicable, this represents a significant reduction in projected costs of quantum computation.
+In addition to our $n$-bit adder, we present an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until it is later uncomputed without using T gates, and discuss some other constructions whose T-count is improved by the temporary logical-AND.
 \end{abstract}

### 2018-01-19 paper.tex
|A> state -> |T> state, and mention Brent-Kung for 12N log depth

@@
 One of the downsides of the surface code is that it has no cheap mechanism to apply non-Clifford operations such as $T$ gates.
-Instead, $T$ gates are performed by distilling and consuming $|A\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle)$ states.
-The T gate is performed via a controlled-NOT from the target qubit into the $|A\rangle$ ancilla, followed by a classical feedback step that measures the ancilla to determine if an $S$ gate is applied to the target.
-Unfortunately, distilling the $|A\rangle$ state that will ultimately be passed into this process has significant cost.
-The cost of distillation is high enough that error corrected quantum computations are expected to bottleneck waiting for $|A\rangle$ states to be produced, so that their runtime is dominated by the number of T gates (as opposed to more traditional measures of circuit depth).
+Instead, $T$ gates are performed by distilling and consuming $|T\rangle = \frac{1}{\sqrt{2}} (|0\rangle + e^{i \pi/4} |1\rangle)$ states.
+The T gate is performed via a controlled-NOT from the target qubit into the $|T\rangle$ ancilla, followed by a classical feedback step that measures the ancilla to determine if an $S$ gate is applied to the target.
+Unfortunately, distilling the $|T\rangle$ state that will ultimately be passed into this process has significant cost.
+The cost of distillation is high enough that error corrected quantum computations are expected to bottleneck waiting for $|T\rangle$ states to be produced, so that their runtime is dominated by the number of T gates (as opposed to more traditional measures of circuit depth).
 
@@
 Decreasing the T-count of addition reduces the T-count of any construction based on addition.
-For example, in \citep{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|A\rangle$ states.
-The time estimate is based on Toffolis having a T-depth of 3, and the $|A\rangle$ count estimate is based on Toffolis having a T-count of 7.
+For example, in \citep{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|T\rangle$ states.
+The time estimate is based on Toffolis having a T-depth of 3, and the $|T\rangle$ count estimate is based on Toffolis having a T-count of 7.
 Because Shor's algorithm is dominated by the cost of controlled additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $\sim 2.7$ and $1$ respectively.
-This reduces the time estimate to 9 hours and the count estimate to $8 \cdot 10^{11}$ distilled $|A\rangle$ states.
+This reduces the time estimate to 9 hours and the count estimate to $8 \cdot 10^{11}$ distilled $|T\rangle$ states.
 
@@
 However, there are three ways in which our phase gradient construction's cost does depend on $\epsilon$.
-First, $\epsilon$ still bounds the minimum quality of the $|A\rangle$ states powering the T gates.
+First, $\epsilon$ still bounds the minimum quality of the $|T\rangle$ states powering the T gates.
 Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\lg \frac{1}{\epsilon}$.
@@
 The representative ancilla is then used to control a CNOT onto the target qubit.
-This construction takes $4n-4$ T gates to perform a NOT with $n$ controls, but is simply a reframing of the nesting construction from \cite{Jones2013}.
+This construction takes $4n-4$ T gates to perform a NOT with $n$ controls (and is equivalent to a nesting construction from \cite{Jones2013}).
 As when replacing Toffoli gates in compute/uncompute pairs with a temporary logical-AND, the ancilla representing the $n$ controls should be kept and used multiple times whenever possible (instead of being uncomputed and recomputed).
@@
 Although using temporary logical-ANDs improves the T-count of logarithmic-depth adders, the resulting T-count is still worse than the T-count of linear-depth adders.
-This means that in practice, given our expectation that error corrected quantum computation will bottleneck waiting on $|A\rangle$ states to perform T gates, the linear-depth adders will often execute in less time than the logarithmic-depth adders.
+This means that in practice, given our expectation that error corrected quantum computation will bottleneck waiting on $|T\rangle$ states to perform T gates, the linear-depth adders will often execute in less time than the logarithmic-depth adders.
+In fact, by using temporary logical-ANDs to implement the classical Brent-Kung adder \cite{brent1982} in a reversible fashion, two $N$-bit numbers can be added in $O(\lg N)$ depth with a T-count of only $12N$.
 
 Finally, we note that Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of mathematics.
-For example, we used temporary logical-ANDs to cut the projected T-count of a step in a simulation algorithm under development by nearly 30\% \citep{RyanEmails2017}.
+For example, we used temporary logical-ANDs to cut the projected T-count of a step in a chemistry algorithm under development by nearly 30\% \citep{RyanEmails2017}.
 
@@
 	The computation circuit (top) has a T-count of 4 and a T-depth of 1.
-	Note that the $|A\rangle$ state input contributes to the T-count, because $|A\rangle$ states are the resource used to perform T gates.
+	Note that the $|T\rangle$ state input contributes to the T-count, because $|T\rangle$ states are the resource used to perform T gates.
 	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero.
@@
     An alternative uncomputation construction is to simply do the reverse of the computation circuit.
-    This alternative approach has a net T-count of 2 (because an $|A\rangle$ state is recovered).
+    This alternative approach has a net T-count of 2 (because an $|T\rangle$ state is recovered).
     The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.

### 2018-01-19 paper.tex
Re-insert figures into main text flow
Quantum style

@@
-\documentclass[twocolumn]{revtex4-1}
+\documentclass[twocolumn]{quantumarticle-customized}
 \usepackage{amssymb}
@@
 \begin{abstract}
-We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \citep{Amy2013, Cuccaro2004, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
+We improve the number of T gates needed to perform an $n$-bit adder from $8n + O(1)$ \cite{Amy2013, Cuccaro2004, AustinDiscussionsAndEmails2017} to $4n + O(1)$.
 We do so via a ``temporary logical-AND" construction which uses four T gates to store the logical-AND of two qubits into an ancilla and zero T gates to later erase the ancilla.
@@
 
-The surface code \citep{Brav98,Denn02,Raus07,Raus07d,Fowler2012} is a quantum error correcting code that works on a 2D nearest-neighbour array of qubits and achieves a threshold error rate of approximately 1\%.
-This makes the surface code a likely component in the architecture of future error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies \citep{Schl11,Bare13,Gamb17,Leik17,Laht17} and other well understood error correcting codes either have lower thresholds or require stronger connectivity.
+The surface code \cite{Brav98,Denn02,Raus07,Raus07d,Fowler2012} is a quantum error correcting code that works on a 2D nearest-neighbour array of qubits and achieves a threshold error rate of approximately 1\%.
+This makes the surface code a likely component in the architecture of future error corrected quantum computers, because 2D arrays of qubits with nearest-neighbor connections are possible with many qubit technologies \cite{Schl11,Bare13,Gamb17,Leik17,Laht17} and other well understood error correcting codes either have lower thresholds or require stronger connectivity.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{adder.pdf}
+  \caption{
+	A 5-bit adder with T-count of 16.
+	Uses Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures requiring no T gates.
+	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
+	See \autoref{fig:adder-building-block} for the adder building-block and \autoref{fig:temporary-logical-AND} for the logical-AND computation and uncomputation circuits.
+  }
+  \label{fig:adder}
+\end{figure}
 
@@
 
-The textbook construction of a Toffoli gate uses seven T gates \citep{Nielsen2009}.
+The textbook construction of a Toffoli gate uses seven T gates \cite{Nielsen2009}.
 When Toffoli operations are paired, i.e.
 when an initial Toffoli operation is later uncomputed by a second Toffoli operation, each Toffoli in the pair can omit three of the T gates from the textbook construction.
-This introduces phase errors but, assuming intermediate operations aren't sensitive to the phase errors, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \citep{Barenco1995, Nielsen2009}.
-It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and a classically conditioned fixup operation \citep{Jones2013}.
+This introduces phase errors but, assuming intermediate operations aren't sensitive to the phase errors, the second Toffoli gate can uncompute the phase errors while uncomputing the state permutation \cite{Barenco1995, Nielsen2009}.
+It is also possible to reduce the T-count of an unpaired Toffoli gate to 4 by using an ancilla qubit and a classically conditioned fixup operation \cite{Jones2013}.
 
-% [figure 1 "adder" was here]
+The Cuccaro adder uses $2n + O(1)$ Toffoli gates \cite{Cuccaro2004}.
+Existing constructions implement the Cuccaro adder's Toffolis using $8n + O(1)$ T gates \cite{Barenco1995, Cuccaro2004, Amy2013}.
+This T-count can be achieved either by applying the ancilla-and-fixup construction from \cite{Jones2013} to each individual Toffoli gate, or by noting that all but one of the adder's Toffoli gates appear in compute/uncompute pairs and then applying the matched-phase-error construction from \cite{Barenco1995, Nielsen2009} to each pair.
 
-The Cuccaro adder uses $2n + O(1)$ Toffoli gates \citep{Cuccaro2004}.
-Existing constructions implement the Cuccaro adder's Toffolis using $8n + O(1)$ T gates \citep{Barenco1995, Cuccaro2004, Amy2013}.
-This T-count can be achieved either by applying the ancilla-and-fixup construction from \citep{Jones2013} to each individual Toffoli gate, or by noting that all but one of the adder's Toffoli gates appear in compute/uncompute pairs and then applying the matched-phase-error construction from \citep{Barenco1995, Nielsen2009} to each pair.
+\begin{figure}
+  \includegraphics[width=\linewidth]{adder-building-block.pdf}
+  \caption{
+	Our adder circuit building-block, with a T-count of 4.
+	A variant of the Cuccaro adder building-block \cite{Cuccaro2004}.
+  }
+  \label{fig:adder-building-block}
+\end{figure}
 
-The leading factor of 8 in the T-count of an $n$-bit adder has stood for over a decade \citep{Barenco1995, Cuccaro2004, AustinDiscussionsAndEmails2017}.
+The leading factor of 8 in the T-count of an $n$-bit adder has stood for over a decade \cite{Barenco1995, Cuccaro2004, AustinDiscussionsAndEmails2017}.
 We improve the leading factor with a construction based on the temporary logical-AND, combining ideas from both the matched-phase-error construction and the ancilla-and-fix-up construction.
@@
 
-% [figure 2 "adder-building-block" was here]
-
 Our adder uses temporary logical-AND operations, which we draw as wires emerging out of a pair of controls then later merging into the same pair of controls.
@@
 Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
-Our controlled adder reduces the cost of these additions from $21n + O(1)$ \citep{Coreas2017} to $8n + O(1)$.
+Our controlled adder reduces the cost of these additions from $21n + O(1)$ \cite{Coreas2017} to $8n + O(1)$.
 
@@
 
-% [figure 3 "temporary-logical-and" was here]
+\begin{figure}
+  \includegraphics[width=\linewidth]{temporary-logical-and.pdf}
+  \caption{
+	How to compute and uncompute the logical-AND of two qubits.
+	The computation circuit (top) has a T-count of 4 and a T-depth of 1.
+	Note that the $|T\rangle$ state input contributes to the T-count, because $|T\rangle$ states are the resource used to perform T gates.
+	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero.
+    \\
+    An alternative uncomputation construction is to simply do the reverse of the computation circuit.
+    This alternative approach has a net T-count of 2 (because an $|T\rangle$ state is recovered).
+    The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.
+  }
+  \label{fig:temporary-logical-AND}
+\end{figure}
 
 Decreasing the T-count of addition reduces the T-count of any construction based on addition.
-For example, in \citep{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|T\rangle$ states.
+For example, in \cite{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|T\rangle$ states.
 The time estimate is based on Toffolis having a T-depth of 3, and the $|T\rangle$ count estimate is based on Toffolis having a T-count of 7.
@@
 
-% [figure 4 "other-adder-building-blocks" was here]
+\begin{figure}
+  \includegraphics[width=\linewidth]{other-adder-building-blocks.pdf}
+  \caption{
+	Variations on our adder construction.
+	(a) is a controlled-adder building-block with a T-count of 8.
+	The addition will only occur when the control is on.
+	(b) is a temporary-adder building-block with a T-count of 4.
+	Has the same T-count as a normal addition, but makes the result available before any uncomputation steps are needed.
+  }
+  \label{fig:other-adder-building-blocks}
+\end{figure}
 
@@
 
-% [figure 5 "paired-toffoli-replacement-rule" was here]
+\begin{figure}
+  \includegraphics[width=\linewidth]{paired-toffoli-replacement-rule.pdf}
+  \caption{
+	A sufficient condition for replacing a pair of Toffoli gates with a temporary logical-AND, saving 4 T gates.\\
+	1) The later Toffoli gate must be uncomputing the earlier Toffoli gate.\\
+	2) Intermediate operations must not be sensitive to the presence of the entangled ancilla.
+  }
+  \label{fig:paired-toffoli-replacement-rule}
+\end{figure}
 
@@
 For example, the Toffoli gates produced by the $P$ and $P^{-1}$ rounds in Draper et al's logarithmic-depth adder \cite{Draper2004} form compute/uncompute pairs that can be replaced by temporary logical-ANDs.
-Although using temporary logical-ANDs improves the T-count of logarithmic-depth adders, the resulting T-count is still worse than the T-count of linear-depth adders.
-This means that in practice, given our expectation that error corrected quantum computation will bottleneck waiting on $|T\rangle$ states to perform T gates, the linear-depth adders will often execute in less time than the logarithmic-depth adders.
-In fact, by using temporary logical-ANDs to implement the classical Brent-Kung adder \cite{brent1982} in a reversible fashion, two $N$-bit numbers can be added in $O(\lg N)$ depth with a T-count of only $12N$.
+Alternatively, by using temporary logical-ANDs to implement the classical Brent-Kung adder \cite{brent1982} in a reversible fashion, two $N$-bit numbers can be added in $O(\lg N)$ depth with a T-count of only $12N$.
 
 Finally, we note that Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of mathematics.
-For example, we used temporary logical-ANDs to cut the projected T-count of a step in a chemistry algorithm under development by nearly 30\% \citep{RyanEmails2017}.
+For example, we used temporary logical-ANDs to cut the projected T-count of a step in a chemistry algorithm under development by nearly 30\% \cite{RyanEmails2017}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{synthesize-t-gates-from-temporary-logical-and.pdf}
+  \caption{
+	Three T gates can be synthesized by performing a temporary logical-AND, one T gate, and various Clifford operations.
+	Clifford operations can't produce T gates, so the extra T gates must be coming from the temporary logical-AND.
+	This proves that performing the temporary logical-AND requires at least two T gates.
+  }
+  \label{fig:synthesize-t-gates-from-temporary-logical-and}
+\end{figure}
 
@@
 
-For over a decade, the T-count of addition has been $8n + O(1)$ \citep{Amy2013, Barenco1995, Cuccaro2004}.
+For over a decade, the T-count of addition has been $8n + O(1)$ \cite{Amy2013, Barenco1995, Cuccaro2004}.
 In this paper we showed how to halve the leading factor of this cost by replacing Toffolis in compute/uncompute pairs with temporary logical-ANDs.
@@
 
-% [figure 6 "synthesize-t-gates-from-temporary-logical-and" was here]
-
 Regardless of whether there will be further T-count improvements for operations as basic as Toffoli gates appearing in pairs, it is clear that many other kinds of T-count optimizations are waiting to be found.
 Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions, and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
-
-\section*{Methods}
-
-Not applicable. The circuits can be verified with any quantum circuit simulator.
 
@@
 
-\section*{Competing Interests}
-
-None.
-
-\section*{Contributions}
-
-Craig Gidney: sole author.
-
-\section*{Funding}
-
-None. Work was done while employed at Google.
-
 \bibliographystyle{npj-quantum-information}
@@
 
-\section*{Figure legends}
-
-None.
-
-\section*{Tables}
-
-None.
-
-\section*{Figures}
-
-\begin{figure}[H]
-  \includegraphics[width=\linewidth]{adder.pdf}
-  \caption{
-	A 5-bit adder with T-count of 16.
-	Uses Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures requiring no T gates.
-	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
-	See \autoref{fig:adder-building-block} for the adder building-block and \autoref{fig:temporary-logical-AND} for the logical-AND computation and uncomputation circuits.
-  }
-  \label{fig:adder}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\linewidth]{adder-building-block.pdf}
-  \caption{
-	Our adder circuit building-block, with a T-count of 4.
-	A variant of the Cuccaro adder building-block \citep{Cuccaro2004}.
-  }
-  \label{fig:adder-building-block}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\linewidth]{temporary-logical-and.pdf}
-  \caption{
-	How to compute and uncompute the logical-AND of two qubits.
-	The computation circuit (top) has a T-count of 4 and a T-depth of 1.
-	Note that the $|T\rangle$ state input contributes to the T-count, because $|T\rangle$ states are the resource used to perform T gates.
-	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero.
-    \\
-    An alternative uncomputation construction is to simply do the reverse of the computation circuit.
-    This alternative approach has a net T-count of 2 (because an $|T\rangle$ state is recovered).
-    The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.
-  }
-  \label{fig:temporary-logical-AND}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\linewidth]{other-adder-building-blocks.pdf}
-  \caption{
-	Variations on our adder construction.
-	(a) is a controlled-adder building-block with a T-count of 8.
-	The addition will only occur when the control is on.
-	(b) is a temporary-adder building-block with a T-count of 4.
-	Has the same T-count as a normal addition, but makes the result available before any uncomputation steps are needed.
-  }
-  \label{fig:other-adder-building-blocks}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\linewidth]{paired-toffoli-replacement-rule.pdf}
-  \caption{
-	A sufficient condition for replacing a pair of Toffoli gates with a temporary logical-AND, saving 4 T gates.\\
-	1) The later Toffoli gate must be uncomputing the earlier Toffoli gate.\\
-	2) Intermediate operations must not be sensitive to the presence of the entangled ancilla.
-  }
-  \label{fig:paired-toffoli-replacement-rule}
-\end{figure}
-
-\vspace{5cm}
-
-\begin{figure}[H]
-  \includegraphics[width=\linewidth]{synthesize-t-gates-from-temporary-logical-and.pdf}
-  \caption{
-	Three T gates can be synthesized by performing a temporary logical-AND, one T gate, and various Clifford operations.
-	Clifford operations can't produce T gates, so the extra T gates must be coming from the temporary logical-AND.
-	This proves that performing the temporary logical-AND requires at least two T gates.
-  }
-  \label{fig:synthesize-t-gates-from-temporary-logical-and}
-\end{figure}
-
 \end{document}

### 2018-04-04 paper.tex
Switch biblio style to plain
Proof reading before updating arxiv version
Initial cleanup and updates based on review feedback

@@
 
+\section*{Resource Tradeoffs}
+
+Before we move on to our results, it is necessary to discuss the tradeoffs between various resources in a quantum computation.
+We will address two questions.
+First: why the focus on T-count?
+Why not circuit depth, or total gate complexity?
+Second: what is the opportunity cost of using ancillae qubits?
+
+In error corrected quantum computation based on the surface code, there are four main classes of gates: Pauli operations such as the X and Z gates; unitary Clifford operations such as the CNOT, H, and S gates; Measurement; and non-Clifford operations such as the T gate.
+
+Pauli operations are applied via classical post-processing, and so are considered to be free.
+They don't require anything extra to be done to the qubits.
+
+Clifford operations are applied by introducing topological braids or defects throughout the computation.
+Braids and defects are interesting because they do not fundamentally distinguish between spacelike and timelike directions.
+Long sequences of Clifford operations can be performed instantaneously by laying them out spacewise, or even backwards in time, instead of forward along the time axis.
+Multiple Clifford operations can't be scheduled to happen in the same place at the same time, but otherwise there is extreme flexibility in how Clifford operations are placed.
+With this flexibility in mind, the cost of a Clifford operation is estimated by how much ``spacetime volume" is used when laying out the associated braid/defect and comparing this against the available spacetime volume during each cycle on a quantum computer of a given size.
+
+Measurements also technically don't have to be laid out in time-respecting directions.
+However, if the result of measurement $A$ depends on an operation conditioned on measurement $B$, then measurement $A$ must happen after measurement $B$ in real time.
+So measurements have a spacetime volume cost, but they also have an unavoidable time cost.
+If a quantum computation has a measurement-depth of 100000 (i.e.
+has a sequence of 100000 measurements where each depends on the previous), and the round trip time between starting a measurement, reacting to it, and starting the next measurement is 10us, then the computation will take at least one second to complete.
+
+To perform non-Clifford operations, the surface code uses magic state distillation \cite{Fowler2012}.
+Distilling a $|T\rangle$ state uses roughly as much spacetime volume as 50 minimum-size \textsc{CNOT} operations.
+The T-factory used to achieve this volume has a footprint covering roughly 100 logical qubits worth of area.
+On top of the distillation costs, consuming the $|T\rangle$ state to perform a T gate involves a measurement with a conditional fixup operation,
+Meaning T gates contribute to the measurement-depth of a circuit.
+Other non-Clifford operations are approximated by decomposing them into the Clifford+T gate set.
+
+T gates use over an order of magnitude more spacetime volume than other operations, contain measurements that impose an ordering on the computation, and require huge distillation factories covering hundreds of thousands of physical qubits.
+This is why we typically focus on optimizing the T-count of quantum circuits, and treat other operations as having negligible cost.
+
+For truly huge error corrected quantum computers, with enough T factories to produce $|T\rangle$ states faster than they can be consumed, the relevant cost metric does switch from T-count to measurement-depth.
+However, anecdotally, measurement-depth is much easier to optimize than T-count.
+For example, there are adders with logarithmic measurement-depth [[[cite]]] but there are no adders with logarithmic T-counts.
+As another example, Shor's algorithm has a naive measurement-depth of $O(n^3)$ where $n$ is the number of bits in the number to factor.
+But, by parallelizing the carries, additions, and multiplications, the measurement-depth can be made polylogarithmic in $n$ [[[cite]]].
+For the foreseeable future, anytime one finds that a given circuit for Shor's algorithm is bottlenecked by measurement-depth, one can simply reduce the measurement depth by parallelizing more operations until the T-count becomes the relevant bottleneck again.
+
+We do not claim that the T-count is a perfect representation of the cost of a quantum computation, but it is a reasonable proxy to use when optimizing.
+
+We now move on to addressing the second question: what is the opportunity cost of using ancillae qubits?
+Our adder construction consumes $4n$ fewer $|T\rangle$ states than the Cuccaro adder, but in doing so it holds $n$ ancillae through a measurement-depth of $2n$.
+Is this tradeoff worth making?
+The answer to this complicates the neat and tidy ``just use the T-count" conclusion we stated just last paragraph.
+
+Obviously, if performing an addition on a small quantum computer with no space to spare for additional ancillae, it would be necessary to use an inplace adder instead of our adder.
+However, lack of space is not the only reason ancillae can be problematic.
+In particular, note that any surface area dedicated to ancillae qubits is surface area {\em not dedicated to T factories}.
+
+Suppose for the sake of argument that a T factory has a footprint covering the equivalent of 100 logical ancillae qubits, and requires a braiding depth of 5 to produce one $|T\rangle$ state (i.e.
+a $|T\rangle$ state comes out every $5d$ surface code cycles, where $d$ is the code distance).
+Then, in an aggregate sense, we may approximate the opportunity cost of holding an ancilla qubit as $\frac{1}{500} |T\rangle$ states per braid depth.
+
+If the opportunity cost of ancillae not used by the Cucarro adder is included in the T-count of our adder, the result is an ``effective" T-count of (roughly) $4n + \frac{1}{500} n^2$.
+This means that, for a sufficiently large addition, there must be a crossover point where ancillae are kept so long that the opportunity cost is larger than the $4 |T\rangle$ state savings.
+At this crossover, it is better to propagate carries inplace (as in the Cuccarro adder) instead of via ancillae.
+(It would still be beneficial to use ancillae for the subset of bits below the cutoff.)
+Given the very rough numbers that have already been mentioned, this would happen somewhere around the 2000'th bit of an addition.
+However, we caution that this cutoff is a very rough estimate strongly affected by other optimizations and future research and overall system design.
+For example, when a logical qubit is kept idle for a long time, it can be temporarily transformed from its ``braidable" form into a ``memory" form that uses significantly less area [[[cite]]].
+Idle ancillae, like the ones in our adder, have a lower opportunity cost than active ancillae.
+Taking advantage of this would extend the (very rough) cutoff point from being around 2000 to being around 10000.
+
+An interesting consequence of the T-count opportunity cost of ancillae, besides making inplace circuits desirable, is that it makes circuit depth relevant to the T count.
+Logarithmic-depth adders may use a constant factor more $|T\rangle$ states and ancillae than our adder, but they use exponentially less {\em ancilla-depth}.
+The T-count opportunity cost of logarithmic-depth adders grows like $\Theta(n \lg n)$ instead of like $\Theta(n^2)$.
+So, for sufficiently large $n$, their effective T-count is actually lower than the effective T-count of our ripple-carry adder (despite their raw T-count being larger).
+
+Whenever ancillae are used to save T gates, it is important to consider that there may be alternative uses of those ancillae that net more $|T\rangle$ states.
+However, these tradeoffs depend strongly on overall system design.
+System optimization is important, but it is not the subject of this paper.
+We limit ourselves to introducing basic tools (i.e.
+the concept of a temporary logical-AND, as well as an adder that uses $4$ T gates and one ancillae per bit instead of $8$ T gates per bit) that future system architects can combine with other techniques on a case by case basis.
+
 \section*{Results}
@@
 
+\begin{figure}
+  \includegraphics[width=\linewidth]{temporary-logical-and.pdf}
+  \caption{
+	How to compute and uncompute the logical-AND of two qubits.
+	The computation circuit (top) has a T-count of 4 and a measurement-depth of 1.
+	Note that the $|T\rangle$ state input contributes to the T-count, because $|T\rangle$ states are the resource used to perform T gates.
+	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero and a measurement-depth of 1.
+    \\
+    An alternative uncomputation construction is to simply do the reverse of the computation circuit.
+    This alternative approach has a net T-count of 2 (because a $|T\rangle$ state is recovered).
+    The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.
+  }
+  \label{fig:temporary-logical-AND}
+\end{figure}
+
 \autoref{fig:other-adder-building-blocks} shows the building-blocks for two variations on our adder: a controlled adder and a temporary adder.
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{temporary-logical-and.pdf}
-  \caption{
-	How to compute and uncompute the logical-AND of two qubits.
-	The computation circuit (top) has a T-count of 4 and a T-depth of 1.
-	Note that the $|T\rangle$ state input contributes to the T-count, because $|T\rangle$ states are the resource used to perform T gates.
-	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero.
-    \\
-    An alternative uncomputation construction is to simply do the reverse of the computation circuit.
-    This alternative approach has a net T-count of 2 (because an $|T\rangle$ state is recovered).
-    The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.
-  }
-  \label{fig:temporary-logical-AND}
-\end{figure}
-
 Decreasing the T-count of addition reduces the T-count of any construction based on addition.
-For example, in \cite{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|T\rangle$ states.
-The time estimate is based on Toffolis having a T-depth of 3, and the $|T\rangle$ count estimate is based on Toffolis having a T-count of 7.
-Because Shor's algorithm is dominated by the cost of controlled additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $\sim 2.7$ and $1$ respectively.
-This reduces the time estimate to 9 hours and the count estimate to $8 \cdot 10^{11}$ distilled $|T\rangle$ states.
+For example, in \cite{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take $2 \cdot 10^{12}$ distilled $|T\rangle$ states and have a measurement-depth spanning 27 hours (though the actual computation would likely be bottlenecked on distillation rather than measurement).
+The time estimate assumes Toffolis have a measurement-depth of 3, and the T-count estimate assumes Toffolis have a T-count of 7.
+Shor's algorithm is dominated by the cost of controlled additions, so our techniques (and previous work) reduce the average T-count and measurement-depth of the relevant Toffolis to $\sim 2.7$ and $1$ respectively.
+This reduces the measurement-depth estimate to 9 hours and the T-count estimate to $8 \cdot 10^{11}$ distilled $|T\rangle$ states.
 
@@
 
-For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a T-cost of $4n + O(\text{poly}(\lg \frac{1}{\epsilon}) \lg n)$ as follows.
+For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a T-cost of $4n + O(\lg n \lg \frac{1}{\epsilon})$ as follows.
 First, iteratively apply our temporary adder to combine the target qubits into a single binary register storing their Hamming weight.
@@
 Then, for each position $p$ in the Hamming weight register, synthesize and apply the operation $R_Z(\theta \cdot 2^p)$ to the register qubit at that position.
-This uses a number of T gates proportional to some polynomial of the desired bit precision, times the number of bits in the Hamming weight register.
-The Hamming weight register is exponentially smaller than $n$, so this cost is negligible for large $n$.
+This uses $O(\lg n \lg \frac{1}{\epsilon}$ T gates, which is negligible in comparison to $4n$ for large $n$.
 Finally, uncompute the Hamming weight register (using no T gates).
-This applies the $n$ desired $R_Z(\theta)$ operations.
+This completes the application of the $n$ desired $R_Z(\theta)$ operations.
 
@@
 Normally this operation would be implemented by separately applying the operation $R_Z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
-However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{-2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition.
+However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{-2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition \cite{Kitaev2002}.
 Add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
-This construction performs phase gradients using $4n + O(1)$ T gates.
-This T-count is interesting because it appears to be independent of $\epsilon$ despite the phase gradient operation involving arbitrarily small rotations.
-However, there are three ways in which our phase gradient construction's cost does depend on $\epsilon$.
+With our adder, this construction performs phase gradients using $4n + O(1)$ T gates.
+Using phase gradients implemented via our adder, approximate quantum Fourier transform was implemented with 3 to 10 times fewer T gates than previous work \cite{Nam2018}.
+
+The phase gradient T-count of $4n + O(1)$ is interesting because it appears to be independent of $\epsilon$ despite the phase gradient operation involving arbitrarily small rotations.
+(Anecdotally, this is a somewhat common occurrence when using constructions that use Toffoli gates instead of rotation synthesis.)
+However, there are three ways in which this phase gradient construction's cost does depend on $\epsilon$.
 First, $\epsilon$ still bounds the minimum quality of the $|T\rangle$ states powering the T gates.
-Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\lg \frac{1}{\epsilon}$.
-Third, initializing the reusable phase gradient register has a one-time cost of $O(n_{\text{max}} \cdot \text{poly}(\lg \frac{1}{\epsilon}))$ T gates.
+Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\Theta(\lg \frac{1}{\epsilon})$.
+Third, initializing the reusable phase gradient register has a one-time cost of $O(n_{\text{max}} \lg \frac{1}{\epsilon})$ T gates.
 
@@
 	Has the same T-count as a normal addition, but makes the result available before any uncomputation steps are needed.
+	Equivalent to an out-of-place adder, where the result is stored in a separate register.
+	Because uncomputing the temporary-adder uses 0 T gates, this implies that uncomputing a register storing $a+b$, when registers storing both $a$ and $b$ are available, requires 0 T gates.
   }
@@
 
-\bibliographystyle{npj-quantum-information}
+\bibliographystyle{plain}
 \bibliography{citations}

### 2018-04-04 paper.tex
Polishing based on review feedback

--- 2018-01-19 paper.tex
@@
 Because T gates dominate the cost of quantum computation based on the surface code, and temporary logical-ANDs are widely applicable, this represents a significant reduction in projected costs of quantum computation.
-In addition to our $n$-bit adder, we present an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, a temporary adder that can be computed for the same cost as the normal adder but whose result can be kept until it is later uncomputed without using T gates, and discuss some other constructions whose T-count is improved by the temporary logical-AND.
+In addition to our $n$-bit adder, we present an $n$-bit controlled adder circuit with T-count of $8n + O(1)$, an out-of-place adder that can be uncomputed without using T gates, and discuss some other constructions whose T-count is improved by the temporary logical-AND.
 \end{abstract}
@@
 	Uses Clifford operations, four logical-AND computations each with a T-count of 4, and four logical-AND erasures requiring no T gates.
-	Generalizes to an $n$-bit adder with a T-count of $4n - 4$.
+	Generalizes to an $n$-bit adder with a T-count of $4n - 4$ and a measurement depth of $2n - 2$.
 	See \autoref{fig:adder-building-block} for the adder building-block and \autoref{fig:temporary-logical-AND} for the logical-AND computation and uncomputation circuits.
@@
 Unfortunately, distilling the $|T\rangle$ state that will ultimately be passed into this process has significant cost.
-The cost of distillation is high enough that error corrected quantum computations are expected to bottleneck waiting for $|T\rangle$ states to be produced, so that their runtime is dominated by the number of T gates (as opposed to more traditional measures of circuit depth).
+The cost of distillation is high enough that error corrected quantum computations are expected to bottleneck waiting for $|T\rangle$ states to be produced, so that their runtime is dominated by the number of T gates (as opposed to being dominated by gate count, circuit depth, or even measurement depth).
+For example, although our adder has the same measurement depth as the Cuccaro adder (i.e.
+$2n + O(1)$), we still expect our adder to execute more quickly in practice on early error corrected machines (assuming there is room for the ancillae).
 
@@
   \caption{
-	Our adder circuit building-block, with a T-count of 4.
+	Our adder circuit building-block, with a T-count of 4 and a measurement-depth of 2.
 	A variant of the Cuccaro adder building-block \cite{Cuccaro2004}.
@@
 We also discuss other examples.
+
+\section*{Opportunity Cost of Ancillae}
+
+Our adder construction consumes $4n$ fewer $|T\rangle$ states than the Cuccaro adder, but in doing so it holds $n$ ancillae through a measurement-depth of $2n$.
+Before we discuss our adder construction, we must first discuss the question: is that tradeoff worth making?
+
+Obviously, if performing an addition on a small quantum computer with no space to spare for additional ancillae, it would be not be possible to use our adder whereas Cuccaro's would still be viable.
+However, lack of space is not the only reason ancillae can be problematic.
+In particular, note that any qubits being used as ancillae are qubits {\em not being used in T factories}.
+
+Roughly speaking, state-of-the-art T factory designs that have been rigorously validated use 1000 units of ``spacetime volume" to produce one $|T\rangle$ state {\color{red} CITE CITE CITE}.
+On the other hand, if an ancilla qubit is stored as a double-defect and lasts through $k$ serial measurements, this consumes at least $2k$ units of spacetime volume.
+Putting these two facts together, and ignoring the fact that T factories are discrete objects with minimum sizes and packing constraints, we can say that the opportunity cost of holding an ancilla is one $|T\rangle$ state per 500 serial measurements.
+
+With the above in mind, the ``effective" T-count of our adder, including the opportunity cost of ancillae, is $\frac{1}{500} n^2 + 4n + O(1)$ whereas the effective T-count of the Cuccaro adder is $8n + O(1)$.
+For a sufficiently large addition, there is a cutoff point where ancillae are kept so long that the opportunity cost is larger than the four $|T\rangle$ states saved by each ancilla.
+Beyond this cutoff, it is better to propagate carries inplace (as in the Cuccarro adder) instead of via ancillae.
+Given the rough numbers we've given, the cutoff is approximately $n=2000$.
+(Note that it is not necessary for the entire adder to use the same carry propagation strategy.
+Parts below the cutoff can propagate carries via ancillae while parts above the cutoff propagate carries inplace.)
+
+We caution that the cutoff we just estimated is not a fixed constant.
+It is affected by other optimizations, by future research, and by overall system design.
+As the spacetime volume of T factories is improved, the cutoff moves downward.
+If the spacetime volume of qubits is improved, the cutoff moves upward.
+For example, qubits that are idle can be compressed into an inactive ``memory" form that covers six times fewer physical qubits than the usual double-defect form {\color{red} CITE CITE CITE}.
+This means that {\em idle} ancillae, like the ones in our adder, have a lower opportunity cost than active ancillae.
+This optimization extends the cutoff from $n=2000$ to $n=12000$, beyond the size of adders that would be used to break modern RSA keys with Shor's algorithm.
+
+An interesting consequence of ancillae having an T-count opportunity cost, besides making circuits that execute inplace desirable, is that it makes circuit depth relevant to the T-count.
+Logarithmic-depth adders may use a constant factor more $|T\rangle$ states and ancillae than our adder, but they use exponentially less {\em ancilla-depth}.
+The T-count opportunity cost of logarithmic-depth adders grows like $\Theta(n \lg n)$ instead of like $\Theta(n^2)$.
+So, despite their ``raw" T-count being larger, for sufficiently large $n$ their effective T-count must become lower than the effective T-count of our ripple-carry adder.
+
+We stress that the previous paragraph only applies in the regime where there are enough physical qubits to support enough T factories to properly feed a logarithmic-depth adder.
+In the early error corrected regime, when only a few T factories are available, the constant-factor penalty on the T-count of logarithmic-depth adders will result in a runtime that is longer than the runtime of ripple-carry adders (because they both have to wait for $|T\rangle$ states).
+Eventually, as more and more T factories become available, the parallelism inherent to logarithmic-depth adders will become accessible and overcome the constant-factor T-count penalty, resulting in a transition away from rippe-carry adders.
+
+Whenever ancillae are used to save T gates, it is important to consider that there may be alternative uses of those ancillae that net even more $|T\rangle$ states.
+These tradeoffs depend on overall system design.
+Overall system optimization is important, but it is not the subject of this paper and so we will limit ourselves to introducing basic tools (e.g.
+the concept of a temporary logical-AND, as well as an adder that uses $4$ T gates and one ancillae per bit instead of $8$ T gates per bit) that future system architects can combine with other techniques on a case by case basis.
 
@@
 
-\autoref{fig:other-adder-building-blocks} shows the building-blocks for two variations on our adder: a controlled adder and a temporary adder.
-
-Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
-Our controlled adder reduces the cost of these additions from $21n + O(1)$ \cite{Coreas2017} to $8n + O(1)$.
-
-Temporary adders are useful when a circuit is going to compute an addition, then later uncompute it.
-Instead of paying $4n + O(1)$ T gates to compute an addition, and then $4n + O(1)$ more T gates to uncompute the addition with a subtraction, our temporary adder stops halfway through the normal addition (after all the logical-ANDs have been computed, but before they've been uncomputed) and applies a few cleanup operations so that the sum is available for use.
-When the sum is no longer needed, the partially-performed addition is uncomputed instead of completed.
-
 \begin{figure}
@@
 	How to compute and uncompute the logical-AND of two qubits.
-	The computation circuit (top) has a T-count of 4 and a T-depth of 1.
+	The computation circuit (top) has a T-count of 4 and a measurement-depth of 1.
 	Note that the $|T\rangle$ state input contributes to the T-count, because $|T\rangle$ states are the resource used to perform T gates.
-	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero.
+	The uncomputation circuit (bottom) uses a measure-and-fixup approach \cite{Jones2013} that requires only Clifford gates, and so has a T-count of zero and a measurement-depth of 1.
     \\
     An alternative uncomputation construction is to simply do the reverse of the computation circuit.
-    This alternative approach has a net T-count of 2 (because an $|T\rangle$ state is recovered).
+    This alternative approach has a net T-count of 2 (because a $|T\rangle$ state is recovered).
     The resulting temporary logical-AND with a T-count of 6 would still be an improvement on existing work, but would be inferior to the measure-and-fixup approach shown above.
@@
 
+\autoref{fig:other-adder-building-blocks} shows the building-blocks for two variations on our adder: a controlled adder and an out-of-place adder.
+
+Some additions, such as the ones performed by the multiplications within the modular exponentiation in Shor's algorithm, are conditioned on a control qubit.
+Our controlled adder reduces the cost of these additions from $21n + O(1)$ \cite{Coreas2017} to $8n + O(1)$.
+
+Out-of-place adders are useful when a circuit is going to compute an addition, use it for awhile, then uncompute it.
+Our out-of-place adder does not improve on the cost of computing an out-of-place addition.
+However, because our out-of-place adder is based on computating a temporary logical-AND, its inverse does not use any T gates.
+This makes it possible to uncompute an out-of-place addition while consuming no T gates.
+
 Decreasing the T-count of addition reduces the T-count of any construction based on addition.
-For example, in \cite{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take 27 hours and $2 \cdot 10^{12}$ distilled $|T\rangle$ states.
-The time estimate is based on Toffolis having a T-depth of 3, and the $|T\rangle$ count estimate is based on Toffolis having a T-count of 7.
-Because Shor's algorithm is dominated by the cost of controlled additions, our techniques (and previous work) reduce the average T-count and T-depth of its Toffolis to $\sim 2.7$ and $1$ respectively.
-This reduces the time estimate to 9 hours and the count estimate to $8 \cdot 10^{11}$ distilled $|T\rangle$ states.
+For example, in \cite{Fowler2012} it is estimated that factoring a 2048-bit number on a surface-code-based quantum computer would take $2 \cdot 10^{12}$ distilled $|T\rangle$ states and have a measurement-depth spanning 27 hours (though the actual computation would likely be bottlenecked on distillation rather than measurement).
+The time estimate assumes Toffolis have a measurement-depth of 3, and the T-count estimate assumes Toffolis have a T-count of 7.
+Shor's algorithm is dominated by the cost of controlled additions.
+With our techniques, the average T-count and measurement-depth of the relevant Toffolis is $\sim 2.7$ and $1$ respectively.
+With these numbers, the measurement-depth estimate is reduced to 9 hours and the T-count estimate is reduced to $8 \cdot 10^{11}$ distilled $|T\rangle$ states.
 
@@
 
-For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a T-cost of $4n + O(\text{poly}(\lg \frac{1}{\epsilon}) \lg n)$ as follows.
-First, iteratively apply our temporary adder to combine the target qubits into a single binary register storing their Hamming weight.
-This takes $4n + O(\lg n)$ T gates to do.
-Then, for each position $p$ in the Hamming weight register, synthesize and apply the operation $R_Z(\theta \cdot 2^p)$ to the register qubit at that position.
-This uses a number of T gates proportional to some polynomial of the desired bit precision, times the number of bits in the Hamming weight register.
-The Hamming weight register is exponentially smaller than $n$, so this cost is negligible for large $n$.
+For example, our improved adder allows the operation $R_Z(\theta)$ to be applied to $n$ qubits with a T-cost of $4n + O(\lg n \lg \frac{1}{\epsilon})$ as follows.
+First, we must reduce the $n$ target qubits into a binary ``Hamming weight register", which indicates how many of the target qubits are \textsc{on}.
+We start by assigning each qubit a ``weight" of 1.
+We then start applying our out-of-place adder building-block to triplets of qubits of the same weight.
+Each application of the adder will turn 3 relevant qubits of weight $w$ into 1 relevant qubit of weight $w$ and 1 relevant qubit of weight $2w$.
+Note that this means the number of relevant qubits goes down by 1 per adder application, except when only 2 qubits of weight $w$ remain (which, as long as lower-weight qubits are always added first, occurs at most $\lg n$ times).
+Since we start with $n$ qubits, and the Hamming weight register will have $\lg n$ qubits, and the two-remaining-qubits-of-same-weight case occurs at most $\lg n$ times, it will take at most $n - \lg n + \lg n = n$ adder applications to reduce the $n$ initial qubits into the $\lg n$ qubits forming the Hamming weight register.
+Therefore, computing the Hamming weight register has a T-count of at most $4n$ (whereas uncomputing it will be free).
+Now, for each position $p$ in the Hamming weight register, synthesize and apply the operation $R_Z(\theta \cdot 2^p)$ to the register qubit at that position.
+This uses $O(\lg n \lg \frac{1}{\epsilon})$ T gates, which is negligible in comparison to $4n$ for large $n$.
 Finally, uncompute the Hamming weight register (using no T gates).
-This applies the $n$ desired $R_Z(\theta)$ operations.
+This completes the application of the $n$ desired $R_Z(\theta)$ operations.
 
@@
 Normally this operation would be implemented by separately applying the operation $R_Z(\pi 2^{-p})$ to each qubit of the target register, where $p$ is the qubit's index in the register and the number of T gates needed for each rotation depends on the maximum per-gate error $\epsilon$.
-However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{-2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition.
+However, assuming a ``phase gradient register" prepared in the state $2^{-b/2} \sum_{k=0}^{2^b-1} e^{-2 i \pi k / 2^b} |k\rangle$ is available, the phase gradient operation can be performed via addition \cite{Kitaev2002}.
 Add the target register into the phase gradient register, and phase kickback will apply the $\text{Grad}_n$ operation to the target.
-This construction performs phase gradients using $4n + O(1)$ T gates.
+With our adder, this construction performs phase gradients using $4n + O(1)$ T gates.
 This T-count is interesting because it appears to be independent of $\epsilon$ despite the phase gradient operation involving arbitrarily small rotations.
-However, there are three ways in which our phase gradient construction's cost does depend on $\epsilon$.
+However, there are three ways in which this phase gradient construction's cost does depend on $\epsilon$.
 First, $\epsilon$ still bounds the minimum quality of the $|T\rangle$ states powering the T gates.
-Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\lg \frac{1}{\epsilon}$.
-Third, initializing the reusable phase gradient register has a one-time cost of $O(n_{\text{max}} \cdot \text{poly}(\lg \frac{1}{\epsilon}))$ T gates.
+Second, large phase gradients can be truncated down to a size $n_{\text{max}}$ asymptotically equal to $\Theta(\lg \frac{1}{\epsilon})$.
+Third, initializing the reusable phase gradient register has a one-time cost of $O(n_{\text{max}} \lg \frac{1}{\epsilon})$ T gates.
 
@@
 	The addition will only occur when the control is on.
-	(b) is a temporary-adder building-block with a T-count of 4.
-	Has the same T-count as a normal addition, but makes the result available before any uncomputation steps are needed.
+	(b) is a standard out-of-place adder building-block, modified to use a temporary logical-AND.
+	Because of this modification, the inverse of this building-block does not use any T gates.
+	This means that uncomputing a register storing $a+b$, when registers storing both $a$ and $b$ are available, requires no T gates.
   }
@@
 Using temporary logical-ANDs, instead of individually translating each Toffoli gate into T gates, halves the T-count of this approach to constructing Grover oracles.
+(However, note that this naive optimization tends to require an unreasonable number of ancillae and will not generalize to all methods for producing Grover oracles.)
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{synthesize-t-gates-from-temporary-logical-and.pdf}
-  \caption{
-	Three T gates can be synthesized by performing a temporary logical-AND, one T gate, and various Clifford operations.
-	Clifford operations can't produce T gates, so the extra T gates must be coming from the temporary logical-AND.
-	This proves that performing the temporary logical-AND requires at least two T gates.
-  }
-  \label{fig:synthesize-t-gates-from-temporary-logical-and}
-\end{figure}
-
 \section*{Discussion}
@@
 It is interesting to consider if addition or temporary logical-ANDs can be done with still fewer T gates.
-\autoref{fig:synthesize-t-gates-from-temporary-logical-and} presents a simple proof that the T-count of the temporary logical-AND is at least 2.
-Although we suspect the true lower bound is 4, we also expect that there are special cases with lower T-counts.
-For example, if a task involves repeating an action several times (e.g.
-counting by repeatedly conditionally incrementing a register), T gates from Toffolis in uncomputation work at the end of one repetition may cancel against T gates from Toffolis in computation work at the start of the next repetition.
+In \cite{Howard2017} it is proven that a Toffoli gate requires at least 4 T gates, though the specific nature of the proof does not eliminate the possibility that $n$ Toffolis could be implemented with fewer than $4n$ T gates for some large $n$.
+Although we suspect the true lower bound really is 4, we also expect that there are regardless many opportunities to optimize how many T gates go into performing Toffolis.
+For example, if a task involves repeating an action several times, T gates inside Toffolis at the end of one repetition could cancel out T gates in work being done at the start of the next repetition.
 
-Regardless of whether there will be further T-count improvements for operations as basic as Toffoli gates appearing in pairs, it is clear that many other kinds of T-count optimizations are waiting to be found.
+Regardless of whether there will be further T-count improvements for operations as basic as paired Toffoli gates, it is clear that many other kinds of T-count optimizations are waiting to be found.
 Not just in the construction of basic low-level operations, but in medium-level constructions, high-level constructions, lower-than-circuit level constructions, and generally across the whole technology stack that will be needed to perform error corrected quantum computation with the surface code for the first time.
@@
 
-We thank Austin Fowler for sampling the opinions of other researchers, for assistance locating references, and for comments that significantly improved this paper.
+We thank Austin Fowler for sampling the opinions of other researchers, for assistance locating references, for checking our ancilla opportunity cost estimates, and for comments that significantly improved this paper.
 We thank Dave Bacon and Dmitri Maslov for comments that significantly improved this paper.
 
-\bibliographystyle{npj-quantum-information}
+\bibliographystyle{plain}
 \bibliography{citations}

### 2018-04-04 paper.tex
Typo fixes and polish in opportunity cost section. Still one citation missing.

@@
 For example, although our adder has the same measurement depth as the Cuccaro adder (i.e.
-$2n + O(1)$), we still expect our adder to execute more quickly in practice on early error corrected machines (assuming there is room for the ancillae).
+$2n + O(1)$), we still expect our adder to execute more quickly in practice on early error corrected machines (assuming there is room for the $n$ ancillae our adder requires).
 
@@
 
-Roughly speaking, state-of-the-art T factory designs that have been rigorously validated use 1000 units of ``spacetime volume" to produce one $|T\rangle$ state {\color{red} CITE CITE CITE}.
-On the other hand, if an ancilla qubit is stored as a double-defect and lasts through $k$ serial measurements, this consumes at least $2k$ units of spacetime volume.
-Putting these two facts together, and ignoring the fact that T factories are discrete objects with minimum sizes and packing constraints, we can say that the opportunity cost of holding an ancilla is one $|T\rangle$ state per 500 serial measurements.
+It is possible to produce a high-quality $|T\rangle$ state in 960 units of ``spacetime volume" \cite{} (though this is certainly not a lower bound!).
+For comparison, an ancilla qubit stored as a double-defect through $k$ serial measurements will cover at least $2k$ units of spacetime volume.
+Putting these two facts together, and ignoring the fact that T factories are discrete objects with packing constraints, we will approximate the opportunity cost of holding an ancilla as $\frac{1}{480} |T\rangle$ states per measurement-depth.
 
-With the above in mind, the ``effective" T-count of our adder, including the opportunity cost of ancillae, is $\frac{1}{500} n^2 + 4n + O(1)$ whereas the effective T-count of the Cuccaro adder is $8n + O(1)$.
-For a sufficiently large addition, there is a cutoff point where ancillae are kept so long that the opportunity cost is larger than the four $|T\rangle$ states saved by each ancilla.
-Beyond this cutoff, it is better to propagate carries inplace (as in the Cuccarro adder) instead of via ancillae.
-Given the rough numbers we've given, the cutoff is approximately $n=2000$.
-(Note that it is not necessary for the entire adder to use the same carry propagation strategy.
-Parts below the cutoff can propagate carries via ancillae while parts above the cutoff propagate carries inplace.)
+With the above in mind, the ``effective" T-count of our adder, including the opportunity cost of ancillae (and accounting for the fact that some are kept for longer than others), is $\frac{1}{480} n^2 + 4n$ whereas the effective T-count of the Cuccaro adder is $8n$.
+This implies that our adder is worse than the Cucarro adder when $n > 1920$.
+However, switching adders at $n=1920$ ignores the possibility of using a hybrid of both adders where bit positions below a cutoff propagate carries via ancillae as in our adder and bit positions above the cutoff propagate carries inline as in the Cuccaro adder.
+In a hybrid adder of this type, the cutoff actually occurs at $n=960$.
 
-We caution that the cutoff we just estimated is not a fixed constant.
+We caution that the cutoff of 960 that we just estimated is not a fixed constant.
 It is affected by other optimizations, by future research, and by overall system design.
-As the spacetime volume of T factories is improved, the cutoff moves downward.
-If the spacetime volume of qubits is improved, the cutoff moves upward.
-For example, qubits that are idle can be compressed into an inactive ``memory" form that covers six times fewer physical qubits than the usual double-defect form {\color{red} CITE CITE CITE}.
+As the spacetime volume of T factories is improved, the cutoff {\em will} move downward.
+Conversely, if the spacetime volume of ancillae is improved, the cutoff moves upward.
+For example, qubits that are idle can be compressed into an inactive ``memory" form that covers six times fewer physical qubits than the double-defect form \cite{horsman2012}.
 This means that {\em idle} ancillae, like the ones in our adder, have a lower opportunity cost than active ancillae.
-This optimization extends the cutoff from $n=2000$ to $n=12000$, beyond the size of adders that would be used to break modern RSA keys with Shor's algorithm.
+This optimization increases the cutoff to $n \approx 5760$, which is beyond the size of adders that would be used to break modern RSA keys with Shor's algorithm.
 
-An interesting consequence of ancillae having an T-count opportunity cost, besides making circuits that execute inplace desirable, is that it makes circuit depth relevant to the T-count.
+An interesting consequence of ancillae having a T-count opportunity cost, besides making circuits that execute inplace desirable, is that it makes circuit depth relevant to the T-count.
 Logarithmic-depth adders may use a constant factor more $|T\rangle$ states and ancillae than our adder, but they use exponentially less {\em ancilla-depth}.
@@
 
-We stress that the previous paragraph only applies in the regime where there are enough physical qubits to support enough T factories to properly feed a logarithmic-depth adder.
+Beware that the previous paragraph only applies in the regime where there are enough physical qubits to support enough T factories to properly feed a logarithmic-depth adder.
 In the early error corrected regime, when only a few T factories are available, the constant-factor penalty on the T-count of logarithmic-depth adders will result in a runtime that is longer than the runtime of ripple-carry adders (because they both have to wait for $|T\rangle$ states).
-Eventually, as more and more T factories become available, the parallelism inherent to logarithmic-depth adders will become accessible and overcome the constant-factor T-count penalty, resulting in a transition away from rippe-carry adders.
+It is only as more and more T factories become available that the parallelism inherent to logarithmic-depth adders becomes accessible, overcomes the constant-factor T-count penalty, and triggers a gradual transition away from ripple-carry adders.
 
-Whenever ancillae are used to save T gates, it is important to consider that there may be alternative uses of those ancillae that net even more $|T\rangle$ states.
+Whenever ancillae are used to save T gates, it is important to consider that there may be alternative uses of those ancillae that net even more $|T\rangle$ states (or other desirable resources).
 These tradeoffs depend on overall system design.
@@
 Our out-of-place adder does not improve on the cost of computing an out-of-place addition.
-However, because our out-of-place adder is based on computating a temporary logical-AND, its inverse does not use any T gates.
+However, because our out-of-place adder is based on computing a temporary logical-AND, its inverse does not use any T gates.
 This makes it possible to uncompute an out-of-place addition while consuming no T gates.
@@
 For example, the Toffoli gates produced by the $P$ and $P^{-1}$ rounds in Draper et al's logarithmic-depth adder \cite{Draper2004} form compute/uncompute pairs that can be replaced by temporary logical-ANDs.
-Alternatively, by using temporary logical-ANDs to implement the classical Brent-Kung adder \cite{brent1982} in a reversible fashion, two $N$-bit numbers can be added in $O(\lg N)$ depth with a T-count of only $12N$.
+Alternatively, by using temporary logical-ANDs to implement the classical Brent-Kung adder \cite{brent1982} in a reversible fashion, two $n$-bit numbers can be added in $O(\lg n)$ depth with a T-count of only $12n$.
 

### 2018-05-25 paper.tex

@@
 
-It is possible to produce a high-quality $|T\rangle$ state in 960 units of ``spacetime volume" \cite{} (though this is certainly not a lower bound!).
+It is possible to produce a high-quality $|T\rangle$ state in 960 units of ``spacetime volume" \cite{Babbush2018} (though this is certainly not a lower bound!).
 For comparison, an ancilla qubit stored as a double-defect through $k$ serial measurements will cover at least $2k$ units of spacetime volume.
@@
 This means that {\em idle} ancillae, like the ones in our adder, have a lower opportunity cost than active ancillae.
-This optimization increases the cutoff to $n \approx 5760$, which is beyond the size of adders that would be used to break modern RSA keys with Shor's algorithm.
+This optimization increases the cutoff to $n \approx 5760$, which is beyond the size of adders that would be used to break 4096-bit RSA keys with Shor's algorithm.
 
@@
 Finally, we note that Toffolis also appear in compute/uncompute pairs in quantum circuits rooted in physics instead of mathematics.
-For example, we used temporary logical-ANDs to cut the projected T-count of a step in a chemistry algorithm under development by nearly 30\% \cite{RyanEmails2017}.
+For example, we used temporary logical-ANDs to cut the T-count of a chemistry algorithm in preparation by nearly 50\% \cite{Babbush2018}.
 

### 2018-05-25 paper.tex
Second set of revisions

@@
 	The computation circuit (top) has a T-count of 4 and a measurement-depth of 1.
+	For systems where $|T\rangle$ states cannot be used as an input without performing a measurement, it is still possible to achieve a measurement depth of 1 by rearranging the circuit and using a temporary ancilla (e.g.
+as in Figure 1 of \cite{Jones2013}).
 	Note that the $|T\rangle$ state input contributes to the T-count, because $|T\rangle$ states are the resource used to perform T gates.
@@
 Using temporary logical-ANDs, instead of individually translating each Toffoli gate into T gates, halves the T-count of this approach to constructing Grover oracles.
-(However, note that this naive optimization tends to require an unreasonable number of ancillae and will not generalize to all methods for producing Grover oracles.)
+However, note that this naive optimization tends to require an unreasonable number of ancillae and will not generalize to all methods for producing Grover oracles.
 
